### 2013-01-16 Article/Article.tex
Added all files.

\documentclass{ba-kecs}
\usepackage{graphicx,, url}
\usepackage[numbers]{natbib}

\title{Turtlebot SLAM }

\author{Maurice Hermans, Lukas Kang, Michael Norget, Thomas Nowicki, Oliver Trinnes}

\begin{document}

\maketitle

\begin{abstract}
This article accompanies the research project ``TurtleBot SLAM'' from the Master ``Artificial Intelligence'' of Maastricht University.
\end{abstract}

\section{Introduction}
Nowadays a lot of tasks are performed by mobile robots, for example transportation, search and rescue and automated vacuum cleaning. To efficiently perform these tasks the robot needs a map of its environment. Sometimes the data about the environment is already available and using for example GPS the robot can locate itself, making the solving of the task at hand easier. But for other tasks the robot should only rely on internal sensors and make its own map of the environment.

The acquisition of these maps has been a major research focus in the robotics community the last decades. Building or learning these maps under the uncertainty of position and the accuracy of measurements is often referred to as the Simultaneous Localization And Mapping (SLAM) problem. A large variety of solutions is available to the SLAM problem, these approaches can be classified as either filtering or smoothing.

Filtering approaches model the problem as an on-line state estimation where the state of the system consists in the current robot position and the map. New measurements augment and refine the estimate as they come available. Due to their incremental nature these filtering approaches are usually referred to as on-line SLAM methods. The other type of approach estimates the full trajectory from the full set of measurements. These approaches address the so called full SLAM problem and they typically rely on least-square error minimization techniques.

An intuitive way to formulate the SLAM problem is to cast it as a pose graph optimization problem. Where the nodes represent robot poses or landmarks and in which the edges encode sensor measurements that constraint the connected poses. Once such a graph is constructed the method will try to find a configuration of the nodes that is maximally consistent with the measurements.

The rest of this paper is structured as follow. Section 2 will provide the reader with the necessary background information on this topic.

\begin{figure}[htbp]
	\centering
		\includegraphics[width=0.50\textwidth]{vacuum_cleaner.jpg}
	\caption{Automated vacuum cleaner}
	\label{fig:vacuum_cleaner}
\end{figure}

\subsection{Problem Statement}
The focus of this paper is SLAM, more specifically GraphSLAM and how it is used to construct a map of an unknown environment. GraphSLAM returns as a result the optimal graph representing the path the robot has explored. There is also the need for exploration to map the still unseen areas of the space where the robot is located. Since the goal of the robot in this paper is to map the entire area where it has been placed. The data returned by the GraphSLAM algorithm is turned into a occupancy grid map and then used for the autonomous exploration. Each time the robot takes another sensor measurement of its surroundings a node is added to the data in the GraphSLAM algorithm, all the obstacles detected in a measurement are as reference points to optimize the graph. The algorithm is tested in a robot simulator named STAGE with a provided map.

\section{State of the art}
\subsection{ROS}
ROS (Robot Operating System) \cite{Quigley}, is the framework used in this paper to program the robots and it provides an environment to run simulations for the implemented algorithm. As described by Quigley et al. \cite{Quigley}, writing software for robots is difficult since there are so many possible hardware setups which makes code reuse nontrivial. Adding to this problem of hardware setups is the amount of code needed as it must contain a deep stack starting from driver-level software going up through perception, abstract reasoning and beyond. Since the required expertise is well beyond the capabilities of a single researcher, robotics software architectures must also support large scale software integration efforts.

To deal with the above mentioned problems many robotics researchers have created a wide variety of frameworks to deal with this complexity and facilitate them with the needs for their own research. This resulted in the many robotic software systems currently used by academia and industry \cite{Kramer}. Each of these frameworks was designed for a specific purpose, perhaps the available frameworks did not meet the requirements this researcher had or the emphasis was on the wrong aspects according to the researcher.

The ROS framework used in this paper is also the product of trade-offs and prioritizations. The emphasis of this framework is on large-scale integrative robotics research, making it useful in a wide variety of situations.

\subsection{TurtleBot}
The robots on which the algorithm from this paper will be tested are called TurtleBots, see figure \ref{fig:turtlebot}, the simulator will also mimic these robots. The TurtleBot is a low-cost, customizable, personal robot kit with open-source software. As a base iRobot Create is used (A) which holds a battery pack and 150 degrees/second Single Axis Gyro. The model shown in figure  \ref{fig:turtlebot} uses a Microsoft Kinect as sensor (B), the one used in this paper uses a laser range sensor, and a laptop (C) to run all the processes. The robot can also be customized using the mounting hardware (D). The robot runs on an open-source SDK based on ROS which integrates all the software needed to get the TurtleBot running and already includes some advanced capabilities like mapping and navigation.

\begin{figure}[htbp]
	\centering
		\includegraphics[width=0.50\textwidth]{turtlebot.png}
	\caption{A TurtleBot}
	\label{fig:turtlebot}
\end{figure}

\subsubsection{TurtleBot simulator}
To help with development a simulator is used to simulate the TurtleBots, the one used in this paper is called Stage. In stage it is possible to load layouts of rooms, called worlds, to simulate the robot being placed there. Once a world is loaded into stage one can drive around using manual or autonomous navigation. In stage it is possible to visualize the data the robot uses, for example the laser scans it is measuring or the path it has traveled. Then there is another tool to visualize the output of the algorithm the robot is performing, the 3D visualization tool rviz. This tool helps to evaluate and validate the output returned when performing the GraphSLAM algorithm.

\subsection{Exploration}
Many robots can navigate using maps provided to them, usually the mapping is done by humans before the robot is send to perform its task, this however limits the robot in its ability to navigate in unknown environments. The robots discussed in this paper however do not know anything of their environment at the start and need to build their own map. Thats why robots should be able to explore their environment, which we will define as the act of moving through an unknown environment while building a map that will be used for subsequent navigation. A good exploration strategy is one that generates a (nearly) complete map in the shortest amount of time.

Since the robot does not know anything about its environment the central idea is that it should move to where it would gain as much information as possible. This target should be extracted from the information the robot has obtained so far. Initially the robot knowns nothing except what is measures with its sensors. The map the robot creates should be as complete as possible and build as quickly as possible.

To achieve this the robot makes use of the idea of frontier-based exploration, first introduced by Yamauchi \cite{Yamauchi}. Frontiers are regions on the boundaries between open explored and unexplored space. When a robot moves to a frontier it will take measurements of unexplored space and add this new information to its map. As a result the explored region on the map expands and new frontiers are added to the map. By constantly moving to frontiers the robot can keep expanding its knowledge about the environment. Note that when the robot detects an obstacle there will be no frontiers behind them since it is no region separating open explored from unexplored space.

The performance of the exploration is then not only depending on the strategy the robot uses to select frontiers to explore but also on how to incorporate the new information into the map. This is a challenge since all the new information obtained is effected by noise, not only on the sensor measurements but also on the odometry data used to incorporate newly obtained information.

\subsection{SLAM}
To navigate through an unknown environment a robot needs to be able to build a map of the environment and simultaneously be able to localize itself within this map. This is called Simultaneous Localization And Mapping (SLAM). It basically means that the robot should be able to learn the map under pose uncertainty.

A large variety of solutions is available to the SLAM problem. These approaches can be classified into two categories, namely filtering and smoothing.

The filtering approaches make a state estimation where the state of the system are stored in the current position and the map. New measurements are incorporated as they become available. In this category fall popular techniques like Kalman filters, particle filters, or information filters.

Conversely, the smoothing approaches keep using all measurements to estimate the entire trajectory the robot has traveled.

\subsubsection{GraphSLAM}
An intuitive way to address the SLAM problem is to pose it as a graph optimization problem. A graph will be constructed where the nodes represent robot poses or landmarks and where edges correspond to measurements that constraint the connecting nodes. Note that constraints can be contradictory because of the noise on the measurements. Once a graph is constructed the algorithm will try to find a configuration of the nodes that is maximally consistent with the measurements.

\section{Implementation}
\subsection{Wavefront frontier detection}
\subsection{GraphSLAM}
\subsubsection{Scan matching}
\subsection{Mapping}
Figure \ref{fig:random_map} shows a partial map created by the ROS gmapping package.
\begin{figure}[htbp]
	\centering
		\includegraphics[width=0.50\textwidth]{random_map.png}
	\caption{A partial map}
	\label{fig:random_map}
\end{figure}

\section{Experiments and Results}
\subsubsection{Benchmarks}

\section{Discussion}
\subsubsection{Future research}

\section{Conclusion}

\bibliography{references}
\nocite{*}
\onecolumn
\appendix

\end{document}

### 2013-01-17 trunk/turtlebot_graph_slam/src/exploration.cpp
Navigation update

@@ -46,6 +46,7 @@ public:
 
 //Tell the navigation stack where to run to
 void run(double x, double y){
+    ROS_ERROR("Run is called in turtlebot_graph_slam");
 	//@TODO make a turn
 
           //tell the action client that we want to spin a thread by default
@@ -68,7 +69,7 @@ void run(double x, double y){
   ROS_INFO_STREAM("Moving to position x:"<<goal.target_pose.pose.position.x<<" y:"<<goal.target_pose.pose.position.y);
   ac.sendGoal(goal);
 
- //ac.waitForResult();
+  ac.waitForResult();
 
   if(ac.getState() == actionlib::SimpleClientGoalState::SUCCEEDED)
     ROS_INFO_STREAM("Robot moved to position x:"<< goal.target_pose.pose.position.x<<" y:"<<goal.target_pose.pose.position.y);

### 2013-01-17 trunk/turtlebot_graph_slam/src/test.cpp

@@ -326,6 +326,7 @@ void timer_callback(const ros::TimerEvent&) {
         occupancyGrid = updateOccupancyGridAll(occupancyGrid,scans,nodes);
         blocked = false;
     }
+    publishOccupancyGrid(occupancyGrid,occupub);
 }
 
 int main(int argc, char **argv) {

### 2013-01-17 turtlebot_slam/src/exploration.cpp
Navigation update

@@ -7,6 +7,8 @@
 #include <geometry_msgs/PointStamped.h>
 #include <tf/transform_listener.h>
 #include <std_msgs/String.h>
+#include "geometry_msgs/Pose.h"
+#include "nav_msgs/Odometry.h"
 #include <tf/transform_listener.h>
 #include <tf/transform_broadcaster.h>
 #include <geometry_msgs/Vector3.h>
@@ -29,189 +31,222 @@ typedef actionlib::SimpleActionClient<move_base_msgs::MoveBaseAction> MoveBaseCl
 class Exploration {
 
 public:
-	Exploration(ros::NodeHandle& nh) {
-		// Initialize random time generator
-		srand(time(NULL));
-
-		commandPub = nh.advertise < geometry_msgs::Twist > ("cmd_vel", 1);
-
-		mapSub = nh.subscribe("/world", 1, &Exploration::occupancyGridCallback,
-				this);
-
-		frontier_publisher = nh.advertise < sensor_msgs::PointCloud> ("frontiers", 1);
-		next_frontier_publisher = nh.advertise < sensor_msgs::PointCloud> ("next_frontier", 1);
+    Exploration(ros::NodeHandle& nh) {
+        // Initialize random time generator
+        srand(time(NULL));
+        robotX = 0;
+        robotY = 0;
+        commandPub = nh.advertise < geometry_msgs::Twist > ("cmd_vel", 1);
+        onTheMove = false;
+        mapSub = nh.subscribe("/world", 1, &Exploration::occupancyGridCallback,
+                              this);
+        velSub = nh.subscribe("odom", 5, &Exploration::vel_callback, this);
+        frontier_publisher = nh.advertise < sensor_msgs::PointCloud> ("frontiers", 1);
+        next_frontier_publisher = nh.advertise < sensor_msgs::PointCloud> ("next_frontier", 1);
         next_frontier.header.frame_id = "/world";
-		frontier_cloud.header.frame_id = "/world";
-	}
+        frontier_cloud.header.frame_id = "/world";
+    }
 
-//Tell the navigation stack where to run to
-void run(double x, double y){
-	//@TODO make a turn
+    //Tell the navigation stack where to run to
+    void run(double x, double y){
+        onTheMove = true;
+        MoveBaseClient ac("move_base", true);
 
-          //tell the action client that we want to spin a thread by default
-  MoveBaseClient ac("move_base", true);
+        while(!ac.waitForServer(ros::Duration(5.0))){
[…]

### 2013-01-17 turtlebot_slam/src/wfd.cpp

@@ -14,7 +14,7 @@ const int FRONTIER_close_list = 4;
 
 // WFD implementation
 	std::vector<std::vector<int> > frontierDetection(const nav_msgs::OccupancyGrid& map, int x, int y) {
-		ROS_INFO("FRONTIER DETECTION START....");
+//		ROS_INFO("FRONTIER DETECTION START....");
 		std::vector<std::vector<int> > frontierList; 
 		std::vector<int> pose;
 		std::vector<std::vector<int> > marker;
@@ -29,7 +29,7 @@ const int FRONTIER_close_list = 4;
 		pose.push_back(y); // y-coordinate robot
 		//create neccessary lists
 		std::vector<std::vector<int> > queue_m;
-		ROS_INFO("queue_m first push pose[0] = %i , pose[1] = %i", pose[0], pose[1]);
+//		ROS_INFO("queue_m first push pose[0] = %i , pose[1] = %i", pose[0], pose[1]);
 		queue_m.push_back(pose);
 		int index1 = pose[0];
 		int index2 = pose[1];
@@ -141,7 +141,7 @@ const int FRONTIER_close_list = 4;
 			marker[index1][index2] = MAP_close_list;
 			
 		} // end of queue_m (loop)
-		ROS_INFO("WFD terminated");
+//		ROS_INFO("WFD terminated");
 
 		return frontierList;
 	}
@@ -150,7 +150,7 @@ const int FRONTIER_close_list = 4;
 	// needs to take care of the actual positions, corners and borders
 	// need to be handled carefully
 	bool isFrontier(const nav_msgs::OccupancyGrid& grid, int x, int y) {
-        std::vector<int> neighbors = getSurrounding(grid, x, y, 3);//*/getNeighbors(grid, x, y);
+        std::vector<int> neighbors = getNeighbors(grid, x, y);
         int numExplored = 0;
         int numUnknown = 0;
 		// checking neighbors

### 2013-01-17 turtlebot_graph_slam/src/robotpos.cpp
Updates on some parameters

@@ -11,7 +11,7 @@ void robotpospub::robotpos(double x, double y, double roll, double pitch, double
 	robotxx = x;
 	robotyy = y;
 	robotyyaw = yaw;
-    int32_t publish_rate_ = 10;
+    int32_t publish_rate_ = 75;
 	tf::TransformBroadcaster tf_br_;
 	tf::StampedTransform tf_map_to_baselink_;
 	tf::TransformBroadcaster tf_br_1;

### 2013-01-17 trunk/turtlebot_graph_slam/src/test.cpp
Decreased the inflation.

@@ -202,7 +202,7 @@ void vel_callback(const nav_msgs::Odometry& msg) {
         scans.push_back(currentScan);
 
         initialised = true;
-	occupancyGrid = updateOccupancyGrid(occupancyGrid,scans,nodes);
+        occupancyGrid = updateOccupancyGrid(occupancyGrid,scans,nodes);
         publishOccupancyGrid(occupancyGrid,occupub);
         // Initialize robotpublisher He will make a ros::spin at this point so the whole program will be runthrough again
         robopub.robotpos(0,0,0,0,0);

### 2013-01-17 turtlebot_slam/src/exploration.cpp
Implemented evaluation for the selection of the next goal to navigate to.

@@ -68,7 +68,7 @@ public:
         ROS_INFO_STREAM("Moving to position x:"<<goal.target_pose.pose.position.x<<" y:"<<goal.target_pose.pose.position.y);
         ac.sendGoal(goal);
 
-//        ac.waitForResult();
+        ac.waitForResult(ros::Duration(3));
 
         if(ac.getState() == actionlib::SimpleClientGoalState::SUCCEEDED)
             ROS_INFO_STREAM("Robot moved to position x:"<< goal.target_pose.pose.position.x<<" y:"<<goal.target_pose.pose.position.y);
@@ -86,6 +86,21 @@ public:
     }
 
     // get the grid from gmapping by listener
+    double prevGoalX;
+    double prevGoalY;
+    std::vector<std::vector<double> > blocked;
+    int sameCounter;
+
+    bool isBlocked(double x, double y)
+    {
+        for(int i = 0; i<blocked.size(); i++)
+        {
+            if(blocked[i][0] == x && blocked[i][1] == y)
+                return true;
+        }
+        return false;
+    }
+
     void occupancyGridCallback(const nav_msgs::OccupancyGrid occupancyGrid) {
         if(!onTheMove) {
             onTheMove = true;
@@ -131,44 +146,106 @@ public:
                 double minDist = 40000.0;
                 //represents the closest frontier
                 std::vector <int> closestFrontier;
+
+                int fewestObstacles = 999999;
+
+//                std::vector<std::vector<int> > blocked;
+
                 //fill frontier cloud for publishing and calculate frontier closest to robot base
                 for(unsigned int i = 0 ; i < frontiersList.size() ; i++ ) {
                     for(unsigned int j = 0 ; j < frontiersList[i].size() ; j++) {
                         double fX = (frontiersList[i][j] - (int)(occupancyGrid.info.width/2)) * occupancyGrid.info.resolution;
                         frontier_cloud.points[frontierIndex].x = fX;
                         double fY = (frontiersList[i][j+1] - (int)(occupancyGrid.info.height/2)) * occupancyGrid.info.resolution;
+
+
                         frontier_cloud.points[frontierIndex].y = fY;
-                        double distance = sqrt((fX-x)*(fX-x)+(fY-y)*(fY-y));
                         frontier_cloud.points[frontierIndex].z = 0;
                         frontierIndex++;
                         j++;
-                        if(distance < minDist && distance > 3.5) {
-                            closestFrontier = frontiersList[i];
-                            minDist = distance;
[…]

### 2013-01-18 Article/Article.tex
Moved figures to a seperate folder.

@@
 
-\begin{figure}[htbp]
+\begin{figure}[htp]
 	\centering
-		\includegraphics[width=0.50\textwidth]{vacuum_cleaner.jpg}
+		\includegraphics[width=0.50\textwidth]{figures/vacuum_cleaner.jpg}
 	\caption{Automated vacuum cleaner}
@@
 	\centering
-		\includegraphics[width=0.50\textwidth]{turtlebot.png}
+		\includegraphics[width=0.50\textwidth]{figures/turtlebot.png}
 	\caption{A TurtleBot}
@@
 \section{Implementation}
+
+The implementation used for this paper is written on a Linux operating system.
+Programming has been done using ROS and the object oriented programming language C++.
+The installation process for ROS, as done in this paper, can be reviewed on Maastricht's swarmlab homepage \cite{swarmlab}.
+ For the coding part there exist two main packages, one of which includes the exploration and the other one the mapping part of the SLAM problem.
+To enable the robot with exploration capabilities, the ROS internal navigation stack for the TurtleBot has been used, making movement possible by sending a designated goal to the robot.
+
 \subsection{Wavefront frontier detection}
+
+As a part of the assignment the robot has to drive autonomously through an unknown map. This feature is implemented using Wavefront frontier detection (WFD).
+
+WFD is using frontiers for its exploration purposes.
+Creation of frontiers requires the robot to know about the current environment in form of a map.
+In this example an Occupancy grid is used \ref{mapping}.
+The general algorithm is then used to acquire four different data structures that store map and frontier information.
+These four data structures are :
+
+1. Map-Open-List
+\newline
+2. Map-Close-List
+\newline
+3. Frontier-Open-List
+\newline
+4. Frontier-Close-List 
+\newline
+
+To understand the need of these four structures it is necessary to know that WFD is using a breadth first search (BFS), starting at the current robot position that has to be passed into the function.
+WFD also requires a current map representation, in this program acquired by the mapping package.
+ Upon launching the algorithm it first searches for unoccupied space around the current robot location, which is then saved into the Map-Open-List. 
+ Frontiers are found by making use of the occupancy grid's structure.
+A criterion for a point to be a frontier is that it has to have unknown space and known unoccupied space around it.
+If a frontier point is found a new BFS is performed, to extract all adjacent frontier points and connect them to one consistent frontier into the Frontier-Open-List.
+ The benefit of using WFD is that by maintaining these lists, rescans only need to incorporate points that have not been scanned yet, as the algorithm saves once detected open space and frontiers into its adjacent close lists.
+For upfollowing scans these points are then no longer considered and by that the algorithm is optimized.
+
+ Detecting frontiers and navigating towards them is not possible due to ROS's navigation stack.
+Instead of sending the robot the goal to a frontier point, an algorithm has to be performed to find a point next to the frontier cell with as much free space around it as possible.
 \subsection{GraphSLAM}
 \subsubsection{Scan matching}
-\subsection{Mapping}
+
+The scan matcher, used by this program, is necessary to acquire an optimized movement vector between two robot positions, which is needed by the GraphSLAM algorithm.
+The programming of the scan matcher used requires two steps.
+For the first step the two scans are matched together by making use of raw odometry data, retrieved by the robot's motion sensor.
+The improvement done by the scan matcher is then to run an iteration of the iterative closest point algorithm (ICP), to incorporate its perception with the movement.
+
+The general procedure of the algorithm is to first find corresponding neighbors.
+These are found by using the nearest neighbor search.
+Points in this implementation can have multiple correspondences to neighbors.
+To avoid incorporating outlying points, a fixed distance has been set.
+Outliers would otherwise impact further calculations negatively.
+The next step in the process is to calculate center points for both clouds, which are needed to normalize both clouds in respect ot each other.
+After performing the calculation the center points for each cloud are subtracted from every point belonging to that cloud.
+The calculated points are then added to two seperate matrizes, both representing one point cloud.
+The first cloud's matrix is then multiplied by the transposition of the second cloud matrix, so that a new Matrix is created for which we calculate the SVD values to retrieve a rotational matrix, responsible for rotating the second point cloud to the first one.
+\subsection{Mapping}\label{mapping}
+
+For Mapping purposes an Occupancy grid is created on which information about obstacles, free/- and unknown space is saved. Using ROS's visualization program, RVIZ, the map can be visualized entirely.
+
+The occupancy grid is a structure that can be represented by a single dimensional array.
+Its size is responsible for the size of the grid and is often problem specific, especially for the purpose of Slam.
+Making the grid too large can lead to performance problems and by that the grid's size gets to be a tunable parameter that has to be chosen carefully for each problem, although making the grid large prevents the robot from going ``out of space''.
+ This program converts the array into a two dimensional representation, so that navigating to either side gets more intuitive.
+Each value of the array is an integer and represents the current status of the cell.
+The value ranges for ROS's occupancy grid lie between $-1$ and $100$ \cite{occupancy}.
+Each value has its own meaning :
+
+\begin{description}
+\item{$-1$} : unknown space, not yet covered by robot's perception sensors
+\item{$ 0$} : known free space, robot can drive on it safely
+\item{$1-100$} : probability of being an obstacle, $1$ being a low probability and $100$ being a definite obstacle
+\end{description}
+
+\begin{figure}[htbp]
+	\centering
+		\includegraphics[width=0.50\textwidth,height = 44mm]{figures/Occup.png}
+	\caption{Occupancy Grid visualized}
+	\label{fig:Occupancy}
+\end{figure}
+Fig.\ref{fig:Occupancy} shows these three occurrences visualized, with a robot in the middle performing a sample scan of its environment.
+The occupancy grid is using laser scan data for this implementation.
+Each laser scan has information about whether an obstacle has been detected, or whether the maximum range of the scan has been reached.
+The scan is then transformed into a point cloud to extract its coordinate information and by that, obstacles can be found.
+
+Free space is being incorporated into the map by filling the occupancy grid between a detected obstacle, or detected maximum range point and the current robot position with $0$ values.
+This is achieved with the help of a linear function.
+If a point is not the scanner's maximum range away, the point is saved as an obstacle with value $100$.
+
+As time progresses and the robot explores more of its environment the map is constantly updated.
+This is programmed by saving laser scans for every pose of the robot, which, when traveling for a distance of $0.5$ meters is added and by that the occupancy grid is updated.
+Upon a loop closure the occupancy grid is optimized with its information about obstacles and free spaces.
 Figure \ref{fig:random_map} shows a partial map created by the ROS gmapping package.
@@
 	\centering
-		\includegraphics[width=0.50\textwidth]{random_map.png}
+		\includegraphics[width=0.50\textwidth]{figures/random_map.png}
 	\caption{A partial map}

### 2013-01-19 Article/Article.tex
Graph Slam implementation + discussion (Draft)

@@
 \usepackage[numbers]{natbib}
+\usepackage{amssymb,amsmath}
 
@@
 
-\begin{figure}[htbp]
+\begin{figure}[h]
 	\centering
@@
 \subsection{GraphSLAM}
+Two different approaches of graph-based SLAM are implemented for this project. They differ in complexity and thereby applicability to the problem. Both algorithms create an occupancy grid map.
+
+First, an algorithm as suggested by \citep{Grisetti} is implemented.
+Here the graph consists of nodes representing positions along the robot trajectory.
+Odometry and sensor measurements are incorporated into the constraints contained in the edges.
+Again, an edge connects two adjacent poses of the robot.
+This graph-based SLAM approach optimizes the pose graph according to the constraints.
+This ``optimized" trajectory then determines the positions of obstacles on the map by refining the laser scan data according to the poses they were obtained at.
+Each edge thus is labelled with the spatial difference between two poses \(x_i\) and \(x_j\).
+Let this error be denoted by \(e_{ij}(x_i, x_j)\).
+This graph-based SLAM approach calculates the most likely path as follows:
+First it calculates the Jacobians \(A\) and \(B\) of every two estimated positions connected by an edge.
+\begin{equation}
+\label{eq:jacobi_a}
+	A_{ij} = \frac{\partial e_{ij}(x)}{\partial x_i}\uplus
+\end{equation}
+\begin{equation}
+\label{eq:jacobi_b}
+	B_{ij} = \frac{\partial e_{ij}(x)}{\partial x_j}
+\end{equation}
+Note that in Equations \ref{eq:jacobi_a} and \ref{eq:jacobi_b} the x represents the current guess of the concerning robot position.
+These Jacobians are then added to an information matrix \(H\) (see Equation \ref{eq:H}).
+\begin{equation}
+\label{eq:H}
+\begin{aligned}
+	H_{[ii]} += A^{T}_{ij} \Omega_{ij} A_{ij} \\
+	H_{[ij]} += A^{T}_{ij} \Omega_{ij} B_{ij} \\
+	H_{[ji]} += B^{T}_{ij} \Omega_{ij} A_{ij} \\
+	H_{[jj]} += B^{T}_{ij} \Omega_{ij} B_{ij} \\
+\end{aligned}
+\end{equation}
+Furthermore a coefficient vector \(b\) is always updated as shown in Equation \ref{eq:b}. This vector in combination with the \(H\) matrix form a set of linear constraints.
+\begin{equation}
+\label{eq:b}
+\begin{aligned}
+	b_{[i]} += A^{T}_{ij} \Omega_{ij} e_{ij} \\
+	b_{[j]} += B^{T}_{ij} \Omega_{ij} e_{ij} 
+\end{aligned}
+\end{equation}
+Optimizing the vector \(x = (x_1, ..., x_n)^T \) according to the constraints specified in \(H\) and \(b\) is the core of this approach. This step is performed using sparse Cholesky factorization. 
+A new node is added to the graph once the robot moved 0.5 meters.
+The algorithm optimizes the trajectory like explained when a new node is added.
+In one special case the algorithm adds and edge connecting to non-adjacent nodes to the graph.
+This happens when the distance between the two connected robot positions is smaller than a margin.
+In fact, this special case enables the algorithm to create loops in the graph and more accurately estimate the robots trajectory and the map.
+This procedure is referred to as ``loop closing".
+\\
+\\
+The second implementation is based on \cite{Thrun}.
+It creates nodes for each odometry measurement of the robot as well as for each feature in the map spotted by the laser sensor.
+Each occupancy grid cell containing an obstacle is considered a map feature.
+Edges are labelled with spatial constraints of the two nodes they are connecting.
+The output of this implementation is a vector containing the positions of the land marks as well as the trajectory of the robot.
+Here the constraints labelling the edges ar created using Jacobians similar to the previous approach.
+The major difference lies in the node creation.
+Here nodes are added for landmarks while the previous implementation considered the robots' trajectory only.
+Since no obstacle in the final occupancy grid must be missed all obstacle points are treated as landmarks for this implementation.
 \subsubsection{Scan matching}
@@
 \section{Discussion}
+The Wavefront Frontier detection algorithm \citep{Keidar} succeeds in finding frontiers in a fast and accurate manner.
+
+Implementing the algorithm suggested in \citep{Grisetti} turned out to be well suited for attacking the problem statement of this project.
+The maps created by this algorithm look accurate.
+Specifically the margin between the output map and the real map is negligible (FIGURE???).
+Maps are created in real time and loop closures are added when needed.
+In fact, the algorithm keeps updating the map with no noticable delay while the robot drives around steering to new goals selected by the WFD algorithm and navigated by ROS' move\_base package.
+
+The second graph-based SLAM implementation \cite{Thrun} clearly is not suited for robots equipped with laser scanners only.
+Extensive feature extraction methods are required to create occupancy grid maps with the used TurtleBot.
+Furthermore the algorithm is expected to work best with a robot which is able to detect distinct landmarks as well as a simple environment.
+A number of unique landmarks can then serve as orientation points to build the map.
+Additionally the algorithm is able to assess map and trajectory more quickly if there are fewer nodes and constraints in the graph.
+
 \subsubsection{Future research}

### 2013-01-19 Article/Article.tex
Test to see if I set up the svn correctly in windows
Some small consistencies fixed

@@
 \begin{abstract}
-This article accompanies the research project ``TurtleBot SLAM'' from the Master ``Artificial Intelligence'' of Maastricht University.
+This article accompanies the research project "TurtleBot SLAM" from the Master "Artificial Intelligence" of Maastricht University.
 \end{abstract}
@@
 Creation of frontiers requires the robot to know about the current environment in form of a map.
-In this example an Occupancy grid is used \ref{mapping}.
+In this example an Occupancy grid is used \ref{sec:mapping}.
 The general algorithm is then used to acquire four different data structures that store map and frontier information.
@@
 The first cloud's matrix is then multiplied by the transposition of the second cloud matrix, so that a new Matrix is created for which we calculate the SVD values to retrieve a rotational matrix, responsible for rotating the second point cloud to the first one.
-\subsection{Mapping}\label{mapping}
+\subsection{Mapping}\label{sec:mapping}
 

### 2013-01-19 trunk/Article/Article.tex
Improvements of the introduction section

--- 2013-01-19 Article/Article.tex
@@
 \section{Introduction}
-Nowadays a lot of tasks are performed by mobile robots, for example transportation, search and rescue and automated vacuum cleaning.
+Nowadays a lot of tasks are performed by mobile robots, for example transportation, search and rescue and automated vacuum cleaning (figure \ref{fig:vacuum_cleaner}).
 To efficiently perform these tasks the robot needs a map of its environment.
-Sometimes the data about the environment is already available and using for example GPS the robot can locate itself, making the solving of the task at hand easier.
-But for other tasks the robot should only rely on internal sensors and make its own map of the environment.
+Sometimes the data about the environment is already available as a map.
+Using for example GPS the robot can then locate itself, making the planning and solving of the task at hand easier.
+But for tasks in unknown environments or where external referencing  like GPS is unavailable the robot should only rely on internal sensors and make its own map of the environment.
 
-The acquisition of these maps has been a major research focus in the robotics community the last decades.
-Building or learning these maps under the uncertainty of position and the accuracy of measurements is often referred to as the Simultaneous Localization And Mapping (SLAM) problem.
+The acquisition of these maps has been a major research focus in the robotics community the last decades \cite{Grisetti}.
+Building or learning these maps under position uncertainty with inaccurate sensor measurements is often referred to as the Simultaneous Localization And Mapping (SLAM) problem.
 A large variety of solutions is available to the SLAM problem, these approaches can be classified as either filtering or smoothing.
 
-Filtering approaches model the problem as an on-line state estimation where the state of the system consists in the current robot position and the map.
-New measurements augment and refine the estimate as they come available.
+Filtering approaches model the problem as an on-line state estimation.
+The state of the system consists in the current robot position and the map.
+New measurements augment and refine the estimate as they become available.
 Due to their incremental nature these filtering approaches are usually referred to as on-line SLAM methods.
@@
 
-The rest of this paper is structured as follow. Section 2 will provide the reader with the necessary background information on this topic.
+The rest of this paper is structured as follows.
+Section 2 will provide the reader with the necessary background information on this topic.
+Section 3 is about the implementation details of the methods used.
+Section 4 will explain the experiments ran to validate and benchmark the implemented methods and also evaluate the obtained results.
+Section 5 discusses the acquired results of the experiments and proposes further research based on those.
+And the last section wraps up the paper with the conclusions of this research.
 
@@
 The focus of this paper is SLAM, more specifically GraphSLAM and how it is used to construct a map of an unknown environment.
-GraphSLAM returns as a result the optimal graph representing the path the robot has explored.
-There is also the need for exploration to map the still unseen areas of the space where the robot is located.
-Since the goal of the robot in this paper is to map the entire area where it has been placed.
-The data returned by the GraphSLAM algorithm is turned into a occupancy grid map and then used for the autonomous exploration.
-Each time the robot takes another sensor measurement of its surroundings a node is added to the data in the GraphSLAM algorithm, all the obstacles detected in a measurement are as reference points to optimize the graph.
-The algorithm is tested in a robot simulator named STAGE with a provided map.
+GraphSLAM returns as a result the optimized graph representing the path the robot has explored.
+Because the goal of the robot in this paper is to map the entire area where it has been placed.
+There is also the need for an exploration strategy to map the still unseen areas of the space where the robot is located.
+The results returned by the GraphSLAM algorithm are turned into an occupancy grid map and then used for the autonomous exploration.
+The algorithm is tested in a robot simulator called STAGE which simulates a provided map.
 

### 2013-01-20 Article/Article.tex
Introduction update

@@
 \section{Introduction}
-Nowadays a lot of tasks are performed by mobile robots, for example transportation, search and rescue and automated vacuum cleaning.
-To efficiently perform these tasks the robot needs a map of its environment.
-Sometimes the data about the environment is already available and using for example GPS the robot can locate itself, making the solving of the task at hand easier.
-But for other tasks the robot should only rely on internal sensors and make its own map of the environment.
+Nowadays mobile robots are able to assist humans in a diverse set of problems.
+Transportation, search, rescue and automated vacuum cleaning are examples from different fields.
+Typically, mobile robots can move through an environment using legs or wheels.
+In addition they are able to gather information about their environment using sensors.
+Some even have devices similar to the human arm or hand which allows them to interact with the environment.
+The vaccuum cleaning robot presented in Figure \ref{fig:vacuum_cleaner} is an example of a robot which has whells to move around.
+Furthermore it is equipped with a set of sensors.
+These sensors provide data about the location of obstacles as well as, for example, gaps in the floor the robot is moving on.
+To efficiently perform this task robots needs a map of its environment.
+For some tasks map data is available for the robot from the beginning.
+Provided a map the robot can locate itself using the incoming sensor data.
+For tasks in unknown environments, however, the robot must rely on internal sensors and make its own map of the environment.
 
-The acquisition of these maps has been a major research focus in the robotics community the last decades.
-Building or learning these maps under the uncertainty of position and the accuracy of measurements is often referred to as the Simultaneous Localization And Mapping (SLAM) problem.
-A large variety of solutions is available to the SLAM problem, these approaches can be classified as either filtering or smoothing.
+The acquisition of these maps has been a major research focus in the robotics community \cite{Grisetti} [\textbf{MORE SOURCES}].
+Building or learning these maps under position uncertainty provided only sensor measurements is often referred to as the Simultaneous Localization And Mapping (SLAM) problem.
+A large variety of solutions is available to the SLAM problem, these approaches can be classified as either filtering or smoothing [\textbf{SOURCE}].
 
-Filtering approaches model the problem as an on-line state estimation where the state of the system consists in the current robot position and the map.
-New measurements augment and refine the estimate as they come available.
-Due to their incremental nature these filtering approaches are usually referred to as on-line SLAM methods.
+Filtering approaches model the problem as a so called ``on-line" state estimation problem.
+The state of such a system consists of the current robot position and the map.
+New measurements augment and refine the estimate as they become available.
+Due to their incremental nature these filtering approaches are commonly referred to as on-line SLAM methods.
 The other type of approach estimates the full trajectory from the full set of measurements.
-These approaches address the so called full SLAM problem and they typically rely on least-square error minimization techniques.
+These approaches address the so called full SLAM problem and they typically rely on least-square error minimization techniques [\textbf{SOURCE}].
 
-An intuitive way to formulate the SLAM problem is to cast it as a pose graph optimization problem.
-Where the nodes represent robot poses or landmarks and in which the edges encode sensor measurements that constraint the connected poses.
-Once such a graph is constructed the method will try to find a configuration of the nodes that is maximally consistent with the measurements.
+An intuitive way to formulate the SLAM problem is to consider it as a pose graph optimization problem.
+The nodes of the graph represent robot poses or landmarks.
+The edges encode sensor measurements that constraint the connected poses.
+Once such a graph is constructed the graph-based SLAM method tries to find a configuration of the nodes that is maximally consistent with the contraints contained in the edges.
+(\textbf{abstract graph picture here})
 
-The rest of this paper is structured as follow. Section 2 will provide the reader with the necessary background information on this topic.
+The rest of this paper is structured as follows.
+Section \ref{sec:sota} provides the reader with the necessary background information on the topic.
+Section \ref{sec:impl} is about the implementation details of the methods used.
+Section \ref{sec:exp} explains the experiments ran to validate and benchmark the implemented methods and also evaluate the obtained results.
+Section \ref{sec:disc} discusses the acquired results of the experiments and proposes further research based on those.
+And the last section wraps up the paper with the conclusions of this research.
 
@@
 The focus of this paper is SLAM, more specifically GraphSLAM and how it is used to construct a map of an unknown environment.
-GraphSLAM returns as a result the optimal graph representing the path the robot has explored.
-There is also the need for exploration to map the still unseen areas of the space where the robot is located.
-Since the goal of the robot in this paper is to map the entire area where it has been placed.
-The data returned by the GraphSLAM algorithm is turned into a occupancy grid map and then used for the autonomous exploration.
-Each time the robot takes another sensor measurement of its surroundings a node is added to the data in the GraphSLAM algorithm, all the obstacles detected in a measurement are as reference points to optimize the graph.
-The algorithm is tested in a robot simulator named STAGE with a provided map.
+GraphSLAM returns as a result the optimized graph representing the path the robot has explored.
+Because the goal of the robot in this paper is to map the entire area where it has been placed.
+There is also the need for an exploration strategy to map the still unseen areas of the space where the robot is located.
+The results returned by the GraphSLAM algorithm are turned into an occupancy grid map and then used for the autonomous exploration.
+The algorithm is tested in a robot simulator called STAGE which simulates a provided map.
 
 \section{State of the art}
+\label{sec:sota}
 \subsection{ROS}
@@
 \section{Implementation}
+\label{sec:impl}
 
@@
 \section{Experiments and Results}
+\label{sec:exp}
 \subsubsection{Benchmarks}
@@
 \section{Discussion}
+\label{sec:disc}
 The Wavefront Frontier detection algorithm \citep{Keidar} succeeds in finding frontiers in a fast and accurate manner.
@@
 \section{Conclusion}
-
+\label{sec:conc}
 \bibliography{references}

### 2013-01-20 Article/Article.tex
Deleted gmapping figure. Added the beginning for experiments and results.

@@
 Upon a loop closure the occupancy grid is optimized with its information about obstacles and free spaces.
-Figure \ref{fig:random_map} shows a partial map created by the ROS gmapping package.
+
+\section{Experiments and Results}
+This section demonstrates the results of the Turtlebot SLAM implementation as described above.
+Several outputs of the visualization tool RViz are presented and discussed.
+Comparisons are drawn between different variations of the implementation and the performance of the program is evaluated.
+
+\subsection{Visualization}
+When running the program via the launch file, RViz and Stage are automatically started.
+Stage shows the robot in its actual position on an accurate test map (see figure \ref{fig:stage_and_rviz} left).
+RViz shows the current belief of the robot about its environment, position and trajectory (see figure \ref{fig:stage_and_rviz} right).
+The nodes of the pose graph are visualized by green points and red arrows indicating the past orientation of the robot in this position.
+The constraints between the nodes are visualized by lines connecting the poses.
+The yellow points represent the frontier points that have been detected, the orange point being the one the robot is currently navigating to.
+The acquired information about the environment is displayed as an occupancy grid containing obstacles (black), free space (light gray) and unknown space (dark gray).
+
 \begin{figure}[htbp]
 	\centering
-		\includegraphics[width=0.50\textwidth]{figures/random_map.png}
-	\caption{A partial map}
-	\label{fig:random_map}
+		\includegraphics[width=0.50\textwidth]{figures/Stage_and_rviz.jpg}
+	\caption{Robot and environment visualization with Stage (left) and RViz (right)}
+	\label{fig:stage_and_rviz}
 \end{figure}
 
-\section{Experiments and Results}
+\subsection{Loop closing effect}
+
+\begin{figure}[htbp]
+	\centering
+		\includegraphics[width=0.50\textwidth]{figures/Loop_closing_compared.jpg}
+	\caption{Observed environment without (left) and with (right) loop closing of the pose graph}
+	\label{fig:loop_closing_comparison}
+\end{figure}
+
+\subsection{Navigation evaluation}
+
+\begin{figure}[htbp]
+	\centering
+		\includegraphics[width=0.50\textwidth]{figures/Navigation_comparison.jpg}
+	\caption{Resulting robot trajectory when applying different strategies for selecting the next navigation goal}
+	\label{fig:navigation_comparison}
+\end{figure}
+
+\subsection{Runtime}
+
 \label{sec:exp}
-\subsubsection{Benchmarks}
 

### 2013-01-20 trunk/Article/Article.tex
Visualization finished. Loop closing finished. Navigation evaluation finished.

@@
 \section{Introduction}
-Nowadays a lot of tasks are performed by mobile robots, for example transportation, search and rescue and automated vacuum cleaning (figure \ref{fig:vacuum_cleaner}).
-To efficiently perform these tasks the robot needs a map of its environment.
-Sometimes the data about the environment is already available as a map.
-Using for example GPS the robot can then locate itself, making the planning and solving of the task at hand easier.
-But for tasks in unknown environments or where external referencing  like GPS is unavailable the robot should only rely on internal sensors and make its own map of the environment.
+Nowadays mobile robots are able to assist humans in a diverse set of problems.
+Transportation, search, rescue and automated vacuum cleaning are examples from different fields.
+Typically, mobile robots can move through an environment using legs or wheels.
+In addition they are able to gather information about their environment using sensors.
+Some even have devices similar to the human arm or hand which allows them to interact with the environment.
+The vaccuum cleaning robot presented in Figure \ref{fig:vacuum_cleaner} is an example of a robot which has whells to move around.
+Furthermore it is equipped with a set of sensors.
+These sensors provide data about the location of obstacles as well as, for example, gaps in the floor the robot is moving on.
+To efficiently perform this task robots needs a map of its environment.
+For some tasks map data is available for the robot from the beginning.
+Provided a map the robot can locate itself using the incoming sensor data.
+For tasks in unknown environments, however, the robot must rely on internal sensors and make its own map of the environment.
 
-The acquisition of these maps has been a major research focus in the robotics community the last decades \cite{Grisetti}.
-Building or learning these maps under position uncertainty with inaccurate sensor measurements is often referred to as the Simultaneous Localization And Mapping (SLAM) problem.
-A large variety of solutions is available to the SLAM problem, these approaches can be classified as either filtering or smoothing.
+The acquisition of these maps has been a major research focus in the robotics community \cite{Grisetti} [\textbf{MORE SOURCES}].
+Building or learning these maps under position uncertainty provided only sensor measurements is often referred to as the Simultaneous Localization And Mapping (SLAM) problem.
+A large variety of solutions is available to the SLAM problem, these approaches can be classified as either filtering or smoothing [\textbf{SOURCE}].
 
-Filtering approaches model the problem as an on-line state estimation.
-The state of the system consists in the current robot position and the map.
+Filtering approaches model the problem as a so called ``on-line" state estimation problem.
+The state of such a system consists of the current robot position and the map.
 New measurements augment and refine the estimate as they become available.
-Due to their incremental nature these filtering approaches are usually referred to as on-line SLAM methods.
+Due to their incremental nature these filtering approaches are commonly referred to as on-line SLAM methods.
 The other type of approach estimates the full trajectory from the full set of measurements.
-These approaches address the so called full SLAM problem and they typically rely on least-square error minimization techniques.
+These approaches address the so called full SLAM problem and they typically rely on least-square error minimization techniques [\textbf{SOURCE}].
 
-An intuitive way to formulate the SLAM problem is to cast it as a pose graph optimization problem.
-Where the nodes represent robot poses or landmarks and in which the edges encode sensor measurements that constraint the connected poses.
-Once such a graph is constructed the method will try to find a configuration of the nodes that is maximally consistent with the measurements.
+An intuitive way to formulate the SLAM problem is to consider it as a pose graph optimization problem.
+The nodes of the graph represent robot poses or landmarks.
+The edges encode sensor measurements that constraint the connected poses.
+Once such a graph is constructed the graph-based SLAM method tries to find a configuration of the nodes that is maximally consistent with the contraints contained in the edges.
+(\textbf{abstract graph picture here})
 
 The rest of this paper is structured as follows.
-Section 2 will provide the reader with the necessary background information on this topic.
-Section 3 is about the implementation details of the methods used.
-Section 4 will explain the experiments ran to validate and benchmark the implemented methods and also evaluate the obtained results.
-Section 5 discusses the acquired results of the experiments and proposes further research based on those.
+Section \ref{sec:sota} provides the reader with the necessary background information on the topic.
+Section \ref{sec:impl} is about the implementation details of the methods used.
+Section \ref{sec:exp} explains the experiments ran to validate and benchmark the implemented methods and also evaluate the obtained results.
+Section \ref{sec:disc} discusses the acquired results of the experiments and proposes further research based on those.
 And the last section wraps up the paper with the conclusions of this research.
@@
 \section{State of the art}
+\label{sec:sota}
 \subsection{ROS}
@@
 Upon a loop closure the occupancy grid is optimized with its information about obstacles and free spaces.
-Figure \ref{fig:random_map} shows a partial map created by the ROS gmapping package.
+
+\section{Experiments and Results}
+\label{sec:exp}
+This section demonstrates the results of the Turtlebot SLAM implementation as described in section \ref{sec:impl}.
+Several outputs of the visualization tool RViz are presented and discussed.
+Comparisons are drawn between different variations of the implementation and the performance of the program is evaluated.
+
+\subsection{Visualization}
+When running the program via the launch file, RViz and Stage are automatically started.
+Stage shows the robot in its actual position on an accurate test map (see figure \ref{fig:stage_and_rviz} left).
+RViz shows the current belief of the robot about its environment, position and trajectory (see figure \ref{fig:stage_and_rviz} right).
+
+The components of the pose graph and the environment are highlighted by the following labels:
+
+\begin{description}
+\item[green dots: ]past positions of the robot, saved as nodes of the pose graph
+\item[red arrows: ]past orientations of the robot, included in the nodes of the pose graph
+\item[blue lines: ]measurement constraints between the connected poses
+\item[yellow dots: ]frontier points detected by the wavefront frontier detection algorithm
+\item[orange dot: ] next navigation goal
+\item[black area: ] detected obstacles
+\item[light gray area: ] detected free space
+\item[dark gray area: ] unexplored space
+\end{description}
+
 \begin{figure}[htbp]
 	\centering
-		\includegraphics[width=0.50\textwidth]{figures/random_map.png}
-	\caption{A partial map}
-	\label{fig:random_map}
+		\includegraphics[width=0.50\textwidth]{figures/Stage_and_rviz.jpg}
+	\caption{Robot and environment visualization with Stage (left) and RViz (right)}
+	\label{fig:stage_and_rviz}
 \end{figure}
 
-\section{Experiments and Results}
-\subsubsection{Benchmarks}
+\subsection{Loop closing effect}
+One of the core properties of the Graph SLAM algorithm is the correction of the pose graph and the map whenever a loop closing is detected.
+This means that when the robot has been exploring an unknown area and then reenters an explored area, the transformation between the current pose and the past re-detected pose gives essential information on the complete pose graph, correcting the current belief about the environment and position.
+
+The effect of loop closing detection is illustrated in figure \ref{fig:loop_closing_comparison}.
+The transparent red lines indicate the accurate map, while the underlying black areas indicate the belief of the robot about the obstacles of the environment.
+
+The left part of figure \ref{fig:loop_closing_comparison} shows the explored map when walking an eight-shaped path through the map without applying loop detection.
+It can be seen that the belief about the map differs from the actual map in many areas, especially in the lower half of the map because this half has been explored last, exposing it to most of the measurement uncertainties.
+
+The right part of figure \ref{fig:loop_closing_comparison} depicts the exploration when walking the same path as before but applying loop closing.
+This results in a pose graph with 7 more constrains, providing more information for correcting the poses and corresponding measurements.
+One can now see that the lower half of the belief about the map is much more accurate since the area in the center has been re-detected by the algorithm.
+
+\begin{figure}[htbp]
+	\centering
+		\includegraphics[width=0.50\textwidth]{figures/Loop_closing_compared.jpg}
+	\caption{Observed environment without (left) and with (right) loop closing of the pose graph}
+	\label{fig:loop_closing_comparison}
+\end{figure}
+
+\subsection{Navigation evaluation}
+An important aspect which has a high impact on the exploration behavior of the robot is the strategy by which the next navigation goal is determined.
+When facing the decision for the next goal, the given information are the detected frontiers, which can be spread all over the current map.
+The most obvious first implementation idea was to select the frontier which lies closest to the robot as the next navigation goal.
+This is definitely desirable because the robot can then explore this closest point very quickly and find out about the next frontiers.
+
+Figure \ref{fig:navigation_comparison}(a) shows the resulting exploration path of the robot and the explored map after 300 seconds when applying this simple strategy.
+10 nodes and 14 constraints have been created in the pose graph during this time which is very little information.
+Also the nodes are limited a small area of the map.
+This little degree of exploration is due to the fact that when only considering the closest frontier point there is a high risk that points are selected which are very close to obstacles.
+When the navigation algorithm tries to find a way to this point, it is inflating the obstacles first, making very close points unreachable.
+After a point has been discovered to be unreachable, the next closest point to the robot is chosen, which will likely be unreachable as well.
+
+So it can be concluded that the number of close obstacles have to be taken into account in order to select reachable points as next goals.
+In the upcoming strategy, the sum of obstacles within a radius of 50 grid cells is added to the distance to the robot for every frontier point.
+The point with the smallest sum is then selected as the next goal.
+The application of this strategy is shown in figure \ref{fig:navigation_comparison}(b).
+The robot managed to explore the map within 300 seconds, creating a pose graph with 104 nodes and 150 constraints.
+However, one can see that the robot did not take a very exploration-efficient route through the map.
+There are a lot of unreasonable turns in areas that are already explored, while unexplored areas get left out for the time being.
+
+Another aspect is to be added to the evaluation of every frontier point.
+That is the distance between the currently evaluated frontier point and the previous navigation goal.
+When this distance is minimized additionally to the distance to the robot and the number of close obstacles, the algorithm will select the next goals in a chain-like fashion, creating a rather intuitive exploration path through the map (figure \ref{fig:navigation_comparison}(c)).
+The robot was able to explore the map within only 150 seconds, creating a much smaller pose graph of 64 nodes and 66 constraints.
+This evaluation feature has been given a weight factor of 5 to increase the effect on the goal selection.
+
+\begin{figure}[htbp]
+	\centering
+		\includegraphics[width=0.50\textwidth]{figures/Navigation_comparison.jpg}
+	\caption{Resulting robot trajectory when applying different strategies for selecting the next navigation goal}
+	\label{fig:navigation_comparison}
+\end{figure}
+
+\subsection{Runtime}
 
 \section{Discussion}
+\label{sec:disc}
 The Wavefront Frontier detection algorithm \citep{Keidar} succeeds in finding frontiers in a fast and accurate manner.

### 2013-01-20 trunk/Article/Article.tex
state of the art almost ready

@@
 \section{Introduction}
+\label{sec:intro}
 Nowadays mobile robots are able to assist humans in a diverse set of problems.
@@
 Once such a graph is constructed the graph-based SLAM method tries to find a configuration of the nodes that is maximally consistent with the contraints contained in the edges.
-(\textbf{abstract graph picture here})
+(\textbf{ABSTRACT GRAPH PICTURE HERE})
 
@@
 \subsection{Problem Statement}
-The focus of this paper is SLAM, more specifically GraphSLAM and how it is used to construct a map of an unknown environment.
-GraphSLAM returns as a result the optimized graph representing the path the robot has explored.
-Because the goal of the robot in this paper is to map the entire area where it has been placed.
-There is also the need for an exploration strategy to map the still unseen areas of the space where the robot is located.
-The results returned by the GraphSLAM algorithm are turned into an occupancy grid map and then used for the autonomous exploration.
-The algorithm is tested in a robot simulator called STAGE which simulates a provided map.
+\label{sec:problem}
+This paper accompanies a university group project.
+The project assignment is to implement both a graph-based SLAM algorithm and an autonomous exploration strategy.
+The hardware to be used is a mobile robot called TurtleBot.
+Thus the final product should allow the robot to autonomously navigate through an unknown environment and simultaneously build a map of it using a graph-based SLAM algorithm.
+Information about SLAM methods, autonomous exploration and the TurtleBot are given in Section \ref{sec:sota}.
 
 \section{State of the art}
-\label{sec:sota}
+This section presents the state of the art of the hardware, software, and methods used for this project.
+Subsection \ref{subsec:ros} gives an outline about the operating system in which the product is implemented.
+Next, Subsection \ref{subsec:turtle} is about the TurtleBot.
+At last, autonomous exploration (Subsection \ref{subsec:sotaExplore}) and SLAM (Subsection \ref{subsec:sotaSlam})
+\label{sec:sota} are defined and explained.
 \subsection{ROS}
-ROS (Robot Operating System) \cite{Quigley}, is the framework used in this paper to program the robots and it provides an environment to run simulations for the implemented algorithm.
-As described by Quigley et al.
-\cite{Quigley}, writing software for robots is difficult since there are so many possible hardware setups which makes code reuse nontrivial.
-Adding to this problem of hardware setups is the amount of code needed as it must contain a deep stack starting from driver-level software going up through perception, abstract reasoning and beyond.
-Since the required expertise is well beyond the capabilities of a single researcher, robotics software architectures must also support large scale software integration efforts.
+\label{subsec:ros}
+The Robot Operating System (ROS) \cite{Quigley} is the environment used throughout the development of all product components for this project.
+As described by Quigley et al., writing software for robots is difficult since there is a huge variety of possible hardware setups.
+This also makes code reuse non-trivial.
+An additional problem is the amount of code needed to acttually ``run" something on a robot.
+A deep stack including driver-level software, perception modules and, for example, abstract reasoning systems is required before the actual problem can be attacked.
+Since the necessary expertise and effort is beyond capabilities of most single researchers, an environment that features this functionality is required to achieve anything similar to the problem specified in Subsection \ref{sec:problem}.
 
-To deal with the above mentioned problems many robotics researchers have created a wide variety of frameworks to deal with this complexity and facilitate them with the needs for their own research.
-This resulted in the many robotic software systems currently used by academia and industry \cite{Kramer}.
-Each of these frameworks was designed for a specific purpose, perhaps the available frameworks did not meet the requirements this researcher had or the emphasis was on the wrong aspects according to the researcher.
+To deal with the above mentioned problems robotics researchers have created a wide variety of frameworks.
+Many robotic software systems are currently used by researchers and industry \cite{Kramer}.
+Each of these frameworks is designed for one specific purpose.
+Generally, none of these systems is directly applicable to a problem it was not designed for in the first place.
 
-The ROS framework used in this paper is also the product of trade-offs and prioritizations.
-The emphasis of this framework is on large-scale integrative robotics research, making it useful in a wide variety of situations.
+The ROS framework used in this paper is the product of trade-offs and prioritizations.
+The emphasis of this framework lies on large-scale integrative robotics research, making it useful in a wide variety of situations.
+The difference between ROS and the above mentioned frameworks, is ROS' applicability to a wide range of hardware setups.
+It provides several so called stacks.
+Each is an assembly of code made to solve one specific problem.
+To run a stack or your own code on any robot it is necessary to configure the robot and the ROS framework in a way that they can communicate with each other.
+For most robots the configuration is straight forward thanks to the online documentation of ROS \citep{Roswiki}.
+An example for a stack is the multi-robot simulation environment ``Stage".
+It is explained in the following paragraph.
+
+\subsubsection{Simulation}
+During the development process of a system running on a robot one curcial part is testing.
+A ROS stack called Stage enables the developers to test code on a simulated robot.
+Stage allows the simulated robot to move around in virtual environments.
+It visualizes sensor data and other robot data when required.
+The user can load her own maps into stage and specify the kind of noise to be applied to measurements.
+A second important stack needed especially for validation is RViz.
+It visualizes all requested data.
+In Figure \ref{fig:stage_and_rviz} Stage and RViz are presented next to each other.
+While Stage (Left) shows the real map, RViz (Right) presents the map data as seen by the robot.
 
 \subsection{TurtleBot}
-The robots on which the algorithm from this paper will be tested are called TurtleBots, see figure \ref{fig:turtlebot}, the simulator will also mimic these robots.
-The TurtleBot is a low-cost, customizable, personal robot kit with open-source software.
-As a base iRobot Create is used (A) which holds a battery pack and 150 degrees/second Single Axis Gyro.
-The model shown in figure  \ref{fig:turtlebot} uses a Microsoft Kinect as sensor (B), the one used in this paper uses a laser range sensor, and a laptop (C) to run all the processes.
+ \label{subsec:turtle}
+
+The product of this project is designed to run on a robot called TurtleBot(Figure \ref{fig:turtlebot}).
+The TurtleBot is a low-cost, customizable, personal robot kit (\textbf{PARAPHRASE?}).The base of the TurtleBot is an iRobot Create (A).
+It holds a battery pack and 150 degrees/second Single Axis Gyro.
+The model shown in Figure \ref{fig:turtlebot} uses a Microsoft Kinect sensor (B), the one used in this paper uses a laser range sensor, and a laptop (C) to run all the processes.
 The robot can also be customized using the mounting hardware (D).
-The robot runs on an open-source SDK based on ROS which integrates all the software needed to get the TurtleBot running and already includes some advanced capabilities like mapping and navigation.
-
+The robot runs on an open-source SDK based on ROS which integrates all the software needed to get the TurtleBot running.
+Already included are some advanced capabilities like mapping and navigation.
 \begin{figure}[h]
@@
 
-\subsubsection{TurtleBot simulator}
-To help with development a simulator is used to simulate the TurtleBots, the one used in this paper is called Stage.
-In stage it is possible to load layouts of rooms, called worlds, to simulate the robot being placed there.
-Once a world is loaded into stage one can drive around using manual or autonomous navigation.
-In stage it is possible to visualize the data the robot uses, for example the laser scans it is measuring or the path it has traveled.
-Then there is another tool to visualize the output of the algorithm the robot is performing, the 3D visualization tool rviz.
-This tool helps to evaluate and validate the output returned when performing the GraphSLAM algorithm.
+\subsection{Exploration}
+\label{subsec:sotaExplore}
+Usually robots localize themselves and navigate through environments using a map which is provided by an external source.
+In this case navigating from one location to another is an easy problem.
+If, however, the robot does not have access to a map, navigating through an environment is problematic.
+The problem treated in this paper (Subsection \ref{sec:problem}) assumes an unknown environment and, hence, no map.
+The robot needs to be able to explore its environment without any knowledge about it.
+It needs to rely on the incoming laser and odometry data only.
+In addition to the creation of the map and the localization of itself (the SLAM problem) the robot needs to explore the environment as efficient as possible.
+We define a good exploration strategy as one that generates an accurate map in the shortest amount of time.
 
-\subsection{Exploration}
-Many robots can navigate using maps provided to them, usually the mapping is done by humans before the robot is send to perform its task, this however limits the robot in its ability to navigate in unknown environments.
-The robots discussed in this paper however do not know anything of their environment at the start and need to build their own map.
-Thats why robots should be able to explore their environment, which we will define as the act of moving through an unknown environment while building a map that will be used for subsequent navigation.
-A good exploration strategy is one that generates a (nearly) complete map in the shortest amount of time.
-
-Since the robot does not know anything about its environment the central idea is that it should move to where it would gain as much information as possible.
-This target should be extracted from the information the robot has obtained so far.
-Initially the robot knowns nothing except what is measures with its sensors.
-The map the robot creates should be as complete as possible and build as quickly as possible.
-
-To achieve this the robot makes use of the idea of frontier-based exploration, first introduced by Yamauchi \cite{Yamauchi}.
-Frontiers are regions on the boundaries between open explored and unexplored space.
-When a robot moves to a frontier it will take measurements of unexplored space and add this new information to its map.
-As a result the explored region on the map expands and new frontiers are added to the map.
-By constantly moving to frontiers the robot can keep expanding its knowledge about the environment.
-Note that when the robot detects an obstacle there will be no frontiers behind them since it is no region separating open explored from unexplored space.
-
-The performance of the exploration is then not only depending on the strategy the robot uses to select frontiers to explore but also on how to incorporate the new information into the map.
-This is a challenge since all the new information obtained is effected by noise, not only on the sensor measurements but also on the odometry data used to incorporate newly obtained information.
+Ideally the robot updates its internal map of the environment constantly while moving.
+One idea to efficiently explore the environment is to move to locations which maximize the robots information gain.
+This target location needs to be extracted from the limited information the robot has collected with its sensors.
+To achieve this the robot makes use of the idea of frontier-based exploration, first introduced by Yamauchi \cite{Yamauchi} in 1998 (\textbf{REALLY?}).
+Frontiers are defined to be the boundaries between free and unexplored space.
+Intuitively, when a robot moves to a frontier it is likely that it makes measurements in previously unexplored space.
+Assuming that the new data is immediately added to its map, the explored region grows once the robot reaches a frontier.
+Moreover an extended map results in new frontiers.
+By constantly moving to frontiers the robot expands its knowledge about the environment.
 
 \subsection{SLAM}
-To navigate through an unknown environment a robot needs to be able to build a map of the environment and simultaneously be able to localize itself within this map.
-This is called Simultaneous Localization And Mapping (SLAM).
-It basically means that the robot should be able to learn the map under pose uncertainty.
+\label{subsec:sotaSlam}
+As mentioned in Section \ref{sec:intro} Simultaneous Localization And Mapping (SLAM) is the process of building a map based on sensor data while simultaneously localizing yourself in an unknown environment.
 
-A large variety of solutions is available to the SLAM problem. These approaches can be classified into two categories, namely filtering and smoothing.
-
-The filtering approaches make a state estimation where the state of the system are stored in the current position and the map.
-New measurements are incorporated as they become available.
-In this category fall popular techniques like Kalman filters, particle filters, or information filters.
+SLAM is a problem because, generally, localization requires a map and mapping requires the loacation of a device or a robot that wants to create a map.
+There are quite a few approaches that successfully solve the SLAM problem.
+Mostly, filtering techniques like Kalman filters Particle filters and Information filters are used to estimate the state of the system.
+FastSlam \citep{Montemerlo02} is an example of a SLAM implementation using a particle filter to approximate the robots trajecotry and positions of landmarks on the map.
 
 Conversely, the smoothing approaches keep using all measurements to estimate the entire trajectory the robot has traveled.
-
+\textbf{TODO: CONNECTION}
 \subsubsection{GraphSLAM}
-An intuitive way to address the SLAM problem is to pose it as a graph optimization problem.
+An intuitive way to formulate the SLAM problem is to view it as a pose graph optimization problem.
 A graph will be constructed where the nodes represent robot poses or landmarks and where edges correspond to measurements that constraint the connecting nodes.

### 2013-01-20 trunk/Article/Article.tex

@@
  For the coding part there exist two main packages, one of which includes the exploration and the other one the mapping part of the SLAM problem.
-To enable the robot with exploration capabilities, the ROS internal navigation stack for the TurtleBot has been used, making movement possible by sending a designated goal to the robot.
+To enable the robot with exploration capabilities, the ROS internal navigation stack for the TurtleBot has been used, making movement possible, by sending a designated goal to the robot.
 
@@
 In this example an Occupancy grid is used \ref{sec:mapping}.
-The general algorithm is then used to acquire four different data structures that store map and frontier information.
+The general algorithm is then applied to acquire four different data structures that store map and frontier information.
 These four data structures are :
-
+\newline
 1. Map-Open-List
@@
 
-To understand the need of these four structures it is necessary to know that WFD is using a breadth first search (BFS), starting at the current robot position that has to be passed into the function.
-WFD also requires a current map representation, in this program acquired by the mapping package.
+To understand the need of these four structures it is necessary to know that WFD is using a breadth first search (BFS), starting at the current robot position that has to be passed into the algorithm.
+WFD also requires a current map representation, which is acquired by the mapping package in this program.
 Upon launching the algorithm it first searches for unoccupied space around the current robot location, which is then saved into the Map-Open-List. 
@@
 If a frontier point is found a new BFS is performed, to extract all adjacent frontier points and connect them to one consistent frontier into the Frontier-Open-List.
- The benefit of using WFD is that by maintaining these lists, rescans only need to incorporate points that have not been scanned yet, as the algorithm saves once detected open space and frontiers into its adjacent close lists.
-For upfollowing scans these points are then no longer considered and by that the algorithm is optimized.
+The benefit of using WFD is that by maintaining these lists, rescans only need to incorporate points that have not been scanned yet, as the algorithm saves detected open space and frontiers into its adjacent close lists, if a point is rescanned.
+For up following scans these points are no longer considered and by that the algorithm does not need to be rerun all over again.
 
- Detecting frontiers and navigating towards them is not possible due to ROS's navigation stack.
+ Detecting frontiers and navigating towards them directly is not possible due to ROS's navigation stack.
 Instead of sending the robot the goal to a frontier point, an algorithm has to be performed to find a point next to the frontier cell with as much free space around it as possible.
+For detection purposes an evaluation function for the frontiers has been written, which incorporates three features that are used to send a goal to the robot's navigation stack.
+The three features are :
+ \newline
+ 1. distance of frontier to the robot location
+  \newline
+ 2. number of obstacles around the designated frontier
+  \newline
+ 3. - ?
+  \newline
+
+  The features are added up and the sum is used for the final goal for the navigation stack.
+By applying these three features the robot can always access the next point and is by that able to drive throughout the whole map and perform the algorithm and exploration perfectly.
 \subsection{GraphSLAM}
@@
 The general procedure of the algorithm is to first find corresponding neighbors.
-These are found by using the nearest neighbor search.
+These are found by using the nearest neighbor search, through which adjacent points are found by calculating two points from different sets according to their euclidean distance to one another.
 Points in this implementation can have multiple correspondences to neighbors.
@@
 Outliers would otherwise impact further calculations negatively.
-The next step in the process is to calculate center points for both clouds, which are needed to normalize both clouds in respect ot each other.
+The next step in the process is to calculate center points for both clouds, which are needed to normalize them in respect ot each other.
 After performing the calculation the center points for each cloud are subtracted from every point belonging to that cloud.
-The calculated points are then added to two seperate matrizes, both representing one point cloud.
-The first cloud's matrix is then multiplied by the transposition of the second cloud matrix, so that a new Matrix is created for which we calculate the SVD values to retrieve a rotational matrix, responsible for rotating the second point cloud to the first one.
+The calculated points are then added to two separate matrices, both representing one point cloud as a $2xn$ Matrix, where n is the number of neighbor pairs.
+The first cloud's matrix is then multiplied by the transposition of the second cloud matrix, so that a new matrix is created for which we calculate the Singular Value Decomposition (SVD) values to retrieve a rotational matrix, responsible for rotating the second point cloud to the first one.
+The next step is to translate the second point cloud to the first one with the use of a translation vector.
+Both rotation matrix and translation vector are then passed to a matrix containing both of them as well as a third line with the values $0 0 1$, for symmetry purposes.
+That matrix is then used to adjust the second point cloud to the first one.
+It is also used to correct the robot position.
+
 \subsection{Mapping}\label{sec:mapping}
@@
 \end{figure}
+
 Fig.\ref{fig:Occupancy} shows these three occurrences visualized, with a robot in the middle performing a sample scan of its environment.

### 2013-01-20 Article/Article.tex
Used reference to implementation.
Some changes including itemize

@@
 In this example an Occupancy grid is used \ref{sec:mapping}.
-The general algorithm is then used to acquire four different data structures that store map and frontier information.
+The general algorithm is then applied to acquire four different data structures that store map and frontier information.
 These four data structures are :
-
-1. Map-Open-List
-\newline
-2. Map-Close-List
-\newline
-3. Frontier-Open-List
-\newline
-4. Frontier-Close-List 
-\newline
-
-To understand the need of these four structures it is necessary to know that WFD is using a breadth first search (BFS), starting at the current robot position that has to be passed into the function.
-WFD also requires a current map representation, in this program acquired by the mapping package.
+\begin{enumerate}
+\item{Map-Open-List}
+\item{Map-Close-List}
+\item{Frontier-Open-List}
+\item{Frontier-Close-List}
+\end{enumerate}
+To understand the need of these four structures it is necessary to know that WFD is using a breadth first search (BFS), starting at the current robot position that has to be passed into the algorithm.
+WFD also requires a current map representation, which is acquired by the mapping package in this program.
 Upon launching the algorithm it first searches for unoccupied space around the current robot location, which is then saved into the Map-Open-List. 
@@
 If a frontier point is found a new BFS is performed, to extract all adjacent frontier points and connect them to one consistent frontier into the Frontier-Open-List.
- The benefit of using WFD is that by maintaining these lists, rescans only need to incorporate points that have not been scanned yet, as the algorithm saves once detected open space and frontiers into its adjacent close lists.
-For upfollowing scans these points are then no longer considered and by that the algorithm is optimized.
+The benefit of using WFD is that by maintaining these lists, rescans only need to incorporate points that have not been scanned yet, as the algorithm saves detected open space and frontiers into its adjacent close lists, if a point is rescanned.
+For up following scans these points are no longer considered and by that the algorithm does not need to be rerun all over again.
 
- Detecting frontiers and navigating towards them is not possible due to ROS's navigation stack.
+ Detecting frontiers and navigating towards them directly is not possible due to ROS's navigation stack.
 Instead of sending the robot the goal to a frontier point, an algorithm has to be performed to find a point next to the frontier cell with as much free space around it as possible.
+For detection purposes an evaluation function for the frontiers has been written, which incorporates three features that are used to send a goal to the robot's navigation stack.
+The three features are :
+\begin{enumerate}
+\item{distance of frontier to the robot location}
+\item{number of obstacles around the designated} \item{distance to the previous goal}
+\end{enumerate}
+  The features are added up and the sum is used for the final goal for the navigation stack.
+By applying these three features the robot can always access the next point and is by that able to drive throughout the whole map and perform the algorithm and exploration perfectly.
 \subsection{GraphSLAM}
@@
 \section{Experiments and Results}
-This section demonstrates the results of the Turtlebot SLAM implementation as described above.
+\label{sec:exp}
+This section demonstrates the results of the Turtlebot SLAM implementation as described in section \ref{sec:impl}.
 Several outputs of the visualization tool RViz are presented and discussed.
@@
 RViz shows the current belief of the robot about its environment, position and trajectory (see figure \ref{fig:stage_and_rviz} right).
-The nodes of the pose graph are visualized by green points and red arrows indicating the past orientation of the robot in this position.
-The constraints between the nodes are visualized by lines connecting the poses.
-The yellow points represent the frontier points that have been detected, the orange point being the one the robot is currently navigating to.
-The acquired information about the environment is displayed as an occupancy grid containing obstacles (black), free space (light gray) and unknown space (dark gray).
+
+The components of the pose graph and the environment are highlighted by the following labels:
+
+\begin{description}
+\item[green dots: ]past positions of the robot, saved as nodes of the pose graph
+\item[red arrows: ]past orientations of the robot, included in the nodes of the pose graph
+\item[blue lines: ]measurement constraints between the connected poses
+\item[yellow dots: ]frontier points detected by the wavefront frontier detection algorithm
+\item[orange dot: ] next navigation goal
+\item[black area: ] detected obstacles
+\item[light gray area: ] detected free space
+\item[dark gray area: ] unexplored space
+\end{description}
 
@@
 \subsection{Loop closing effect}
+One of the core properties of the Graph SLAM algorithm is the correction of the pose graph and the map whenever a loop closing is detected.
+This means that when the robot has been exploring an unknown area and then reenters an explored area, the transformation between the current pose and the past re-detected pose gives essential information on the complete pose graph, correcting the current belief about the environment and position.
+
+The effect of loop closing detection is illustrated in figure \ref{fig:loop_closing_comparison}.
+The transparent red lines indicate the accurate map, while the underlying black areas indicate the belief of the robot about the obstacles of the environment.
+
+The left part of figure \ref{fig:loop_closing_comparison} shows the explored map when walking an eight-shaped path through the map without applying loop detection.
+It can be seen that the belief about the map differs from the actual map in many areas, especially in the lower half of the map because this half has been explored last, exposing it to most of the measurement uncertainties.
+
+The right part of figure \ref{fig:loop_closing_comparison} depicts the exploration when walking the same path as before but applying loop closing.
+This results in a pose graph with 7 more constrains, providing more information for correcting the poses and corresponding measurements.
+One can now see that the lower half of the belief about the map is much more accurate since the area in the center has been re-detected by the algorithm.
 
@@
 \subsection{Navigation evaluation}
+An important aspect which has a high impact on the exploration behavior of the robot is the strategy by which the next navigation goal is determined.
+When facing the decision for the next goal, the given information are the detected frontiers, which can be spread all over the current map.
+The most obvious first implementation idea was to select the frontier which lies closest to the robot as the next navigation goal.
+This is definitely desirable because the robot can then explore this closest point very quickly and find out about the next frontiers.
+
+Figure \ref{fig:navigation_comparison}(a) shows the resulting exploration path of the robot and the explored map after 300 seconds when applying this simple strategy.
+10 nodes and 14 constraints have been created in the pose graph during this time which is very little information.
+Also the nodes are limited a small area of the map.
+This little degree of exploration is due to the fact that when only considering the closest frontier point there is a high risk that points are selected which are very close to obstacles.
+When the navigation algorithm tries to find a way to this point, it is inflating the obstacles first, making very close points unreachable.
+After a point has been discovered to be unreachable, the next closest point to the robot is chosen, which will likely be unreachable as well.
+
+So it can be concluded that the number of close obstacles have to be taken into account in order to select reachable points as next goals.
+In the upcoming strategy, the sum of obstacles within a radius of 50 grid cells is added to the distance to the robot for every frontier point.
+The point with the smallest sum is then selected as the next goal.
+The application of this strategy is shown in figure \ref{fig:navigation_comparison}(b).
+The robot managed to explore the map within 300 seconds, creating a pose graph with 104 nodes and 150 constraints.
+However, one can see that the robot did not take a very exploration-efficient route through the map.
+There are a lot of unreasonable turns in areas that are already explored, while unexplored areas get left out for the time being.
+
+Another aspect is to be added to the evaluation of every frontier point.
+That is the distance between the currently evaluated frontier point and the previous navigation goal.
+When this distance is minimized additionally to the distance to the robot and the number of close obstacles, the algorithm will select the next goals in a chain-like fashion, creating a rather intuitive exploration path through the map (figure \ref{fig:navigation_comparison}(c)).
+The robot was able to explore the map within only 150 seconds, creating a much smaller pose graph of 64 nodes and 66 constraints.
+This evaluation feature has been given a weight factor of 5 to increase the effect on the goal selection.
 
@@
 \subsection{Runtime}
-
-\label{sec:exp}
 

### 2013-01-20 trunk/Article/Article.tex
Runtime figure added and used.

@@
 These four data structures are :
-\newline
-1. Map-Open-List
-\newline
-2. Map-Close-List
-\newline
-3. Frontier-Open-List
-\newline
-4. Frontier-Close-List 
-\newline
-
+\begin{enumerate}
+\item{Map-Open-List}
+\item{Map-Close-List}
+\item{Frontier-Open-List}
+\item{Frontier-Close-List}
+\end{enumerate}
 To understand the need of these four structures it is necessary to know that WFD is using a breadth first search (BFS), starting at the current robot position that has to be passed into the algorithm.
@@
 The three features are :
- \newline
- 1. distance of frontier to the robot location
-  \newline
- 2. number of obstacles around the designated frontier
-  \newline
- 3. - ?
-  \newline
-
+\begin{enumerate}
+\item{distance of frontier to the robot location}
+\item{number of obstacles around the designated} \item{distance to the previous goal}
+\end{enumerate}
   The features are added up and the sum is used for the final goal for the navigation stack.
@@
 \subsection{Runtime}
+\begin{figure}[htbp]
+	\centering
+		\includegraphics[width=0.50\textwidth]{figures/Runtime.png}
+	\caption{Runtime of the Graph SLAM update function in micro seconds}
+	\label{fig:runtime}
+\end{figure}
 

### 2013-01-20 trunk/Article/Article.tex
Runtime finished.

@@
 
-\subsection{Runtime}
+\subsection{Runtime analysis}
+The implemented Graph SLAM algorithm has been tested on its feasibility.
+Navigating the robot manually, the complete test map has been explored, building up a pose graph of 123 nodes and 150 constraints.
+While doing that, the runtime of the complete update function was measured in microseconds.
+The update function is executed whenever a new node in the graph is created.
+It contains the algorithms for scan matching and Graph SLAM correction, which have been measured separately.
+
+Figure \ref{fig:runtime} shows the outcome of this runtime measurement. Naturally the runtime increases with increasing graph size, especially due to the growing matrices in the Graph SLAM algorithm.
+
 \begin{figure}[htbp]
@@
 
+Even with a graph size of over 100 nodes, the runtime of the function stays in a very feasible range between 60 and 80 milliseconds.
+The robot can be moved without any delay.
+Taking only the Graph SLAM update into account, the runtime even stays below 20 milliseconds.
+In contrast the scan matching is executed outstandingly quickly since it is independent of the size of the pose graph.
+Most of the time is occupied by updating the occupancy grid because every single grid cell has to redrawn for every pose.
+In order to compromise this, the update of the complete occupancy grid is only executed every third second and only if a loop closing is detected.
 \section{Discussion}

### 2013-01-21 Article/Article.tex
Implementation edited

@@
 \label{sec:intro}
-Nowadays mobile robots are able to assist humans in a diverse set of problems.
+Nowadays mobile robots are able to assist industry and society in a diverse set of problems.
 Transportation, search, rescue and automated vacuum cleaning are examples from different fields.
@@
 
-The product of this project is designed to run on a robot called TurtleBot(Figure \ref{fig:turtlebot}).
+The product of this project is designed to run on a robot called TurtleBot (Figure \ref{fig:turtlebot}).
 The TurtleBot is a low-cost, customizable, personal robot kit (\textbf{PARAPHRASE?}).The base of the TurtleBot is an iRobot Create (A).
@@
 Ideally the robot updates its internal map of the environment constantly while moving.
-One idea to efficiently explore the environment is to move to locations which maximize the robots information gain.
+One idea to efficiently explore the world is to move to locations which maximize the robots information gain.
 This target location needs to be extracted from the limited information the robot has collected with its sensors.
-To achieve this the robot makes use of the idea of frontier-based exploration, first introduced by Yamauchi \cite{Yamauchi} in 1998 (\textbf{REALLY?}).
+To achieve this the robot makes use of the idea of frontier-based exploration, first introduced by Yamauchi \cite{Yamauchi} in 1998.
 Frontiers are defined to be the boundaries between free and unexplored space.
@@
 
-SLAM is a problem because, generally, localization requires a map and mapping requires the loacation of a device or a robot that wants to create a map.
+SLAM is a problem because, generally, localization requires a map and mapping requires the location of the device or robot that wants to create a map.
 There are quite a few approaches that successfully solve the SLAM problem.
-Mostly, filtering techniques like Kalman filters Particle filters and Information filters are used to estimate the state of the system.
-FastSlam \citep{Montemerlo02} is an example of a SLAM implementation using a particle filter to approximate the robots trajecotry and positions of landmarks on the map.
+In general there are two interpretations of SLAM.
+On-line SLAM methods estimate the current robot position along with the map.
+The core of a system solving the on-line SLAM problem is a filtering technique.
+FastSlam \citep{Montemerlo02} is an example of an on-line SLAM implementation using a Particle filter to approximate the robots position and the landmark positions on the map.
+Other methods have the goal to find the trajectory of the robot and the map.
+They solve the full SLAM problem.
+An example of such a an approach is graph-based SLAM of graphSLAM.
 
-Conversely, the smoothing approaches keep using all measurements to estimate the entire trajectory the robot has traveled.
-\textbf{TODO: CONNECTION}
 \subsubsection{GraphSLAM}
 An intuitive way to formulate the SLAM problem is to view it as a pose graph optimization problem.
-A graph will be constructed where the nodes represent robot poses or landmarks and where edges correspond to measurements that constraint the connecting nodes.
-Note that constraints can be contradictory because of the noise on the measurements.
-Once a graph is constructed the algorithm will try to find a configuration of the nodes that is maximally consistent with the measurements.
+Each position along the trajectory of the robot is represented by one node.
+Edges are labelled with relative spatial information between the connecting nodes.
+This graph is build incrementally.
+While the robot is in motion, new nodes are added to the graph.
+While the graph building is a straight forward procedure the essence of the graphSLAM method is the optimization step.
+In this step the algorithm looks for a configuration of the trajectory which is maximally consistent with a system of constraints obtained from the edges.
+This problem is commonly solved using least-square error minimization techniques \citep{Grisetti}.
+
+The next section presents the algorithms implemented for this research project.
 
@@
 
-The implementation used for this paper is written on a Linux operating system.
-Programming has been done using ROS and the object oriented programming language C++.
-The installation process for ROS, as done in this paper, can be reviewed on Maastricht's swarmlab homepage \cite{swarmlab}.
+The algorithms explained in this sections are implemented in the object-oriented programming language C++.
+The code is specifically designed for the ROS environment (see Subsection \ref{subsec:ros}).
+The installation process for ROS, as performed for this paper, can be reviewed on \cite{swarmlab}.\\
+(\textbf{TODO:})
 For the coding part there exist two main packages, one of which includes the exploration and the other one the mapping part of the SLAM problem.
@@
 
-As a part of the assignment the robot has to drive autonomously through an unknown map. This feature is implemented using Wavefront frontier detection (WFD).
+One part of the assignment is autonomous exploration. The robot has to drive autonomously through an unknown world. This feature is implemented using Wavefront frontier detection (WFD) \citep{Keidar}.
 
-WFD is using frontiers for its exploration purposes.
-Creation of frontiers requires the robot to know about the current environment in form of a map.
-In this example an Occupancy grid is used \ref{sec:mapping}.
-The general algorithm is then applied to acquire four different data structures that store map and frontier information.
-These four data structures are :
+WFD is a frontier-based exploration algorithm. The WFD algorithm uses four different data structures. These store map and frontier information during the search process. They are :
 \begin{enumerate}
@@
 \end{enumerate}
-To understand the need of these four structures it is necessary to know that WFD is using a breadth first search (BFS), starting at the current robot position that has to be passed into the algorithm.
-WFD also requires a current map representation, which is acquired by the mapping package in this program.
-Upon launching the algorithm it first searches for unoccupied space around the current robot location, which is then saved into the Map-Open-List. 
+To understand the need of these four structures it is necessary to know that WFD is using a breadth first search (BFS), starting at the current robot position.
+Moreover frontier detection requires the robot to have an internal representation of the current environment.
+In this implementation an Occupancy grid (Subsection \ref{sec:mapping}) is used.
+Upon launching the algorithm it first searches for unoccupied space around the current robot location. This grid cells are then saved in the Map-Open-List. 
 Frontiers are found by making use of the occupancy grid's structure.
-A criterion for a point to be a frontier is that it has to have unknown space and known unoccupied space around it.
-If a frontier point is found a new BFS is performed, to extract all adjacent frontier points and connect them to one consistent frontier into the Frontier-Open-List.
-The benefit of using WFD is that by maintaining these lists, rescans only need to incorporate points that have not been scanned yet, as the algorithm saves detected open space and frontiers into its adjacent close lists, if a point is rescanned.
-For up following scans these points are no longer considered and by that the algorithm does not need to be rerun all over again.
+The criterion for a map-point to be a frontier is the existence of both, unknown space and known unoccupied space, in adjacent map-points.
+When a frontier point is found a new BFS is performed.
+Starting at the current point it extracts all adjacent frontier points and connects them to one frontier.
+It is saved in the Frontier-Open-List.
+The benefit of using WFD is that by maintaining these lists, rescans only need to incorporate points that have not been iterated yet.
+In fact, the algorithm keeps track of detected open space and frontiers.
+Following iterations make use of this information because they do not need to consider previously known regions any more.
 
- Detecting frontiers and navigating towards them directly is not possible due to ROS's navigation stack.
-Instead of sending the robot the goal to a frontier point, an algorithm has to be performed to find a point next to the frontier cell with as much free space around it as possible.
-For detection purposes an evaluation function for the frontiers has been written, which incorporates three features that are used to send a goal to the robot's navigation stack.
+Exploration does not only consist of finding frontiers.
+Based on the frontiers a target location for the robot needs to be selected.
+Navigating to frontier points directly is the simplest frontier-based exploration strategy.
+For this project, however, a more sophisticated implementation is chosen.
+Instead of sending the robot directly to a frontier point, a small search is performed to find suitable destinations.
+Generally, a small region around the frontier cell is searched.
+The goal is to have a cell with as much free space around it as possible.
+For this purposes an evaluation function has been created.
+It incorporates three features that are used to send a goal to the robot's navigation stack.
 The three features are :
 \begin{enumerate}
-\item{distance of frontier to the robot location}
-\item{number of obstacles around the designated} \item{distance to the previous goal}
+\item{Distance of frontier to the robot location}
+\item{Number of obstacles around the designated} \item{Distance to the previous goal}
 \end{enumerate}
-  The features are added up and the sum is used for the final goal for the navigation stack.
-By applying these three features the robot can always access the next point and is by that able to drive throughout the whole map and perform the algorithm and exploration perfectly.
+ The features have equal weight.
+Their sum is used to find the final goal to be sent to the navigation stack.
+This evaluation function enables the algorithm to select a reachable destination for the robot in almost every case.
+Additionally the information gain is still next to optimal since the selected goal is preferred to be close to frontier points.
+At last the third feature makes sure that the robots path is smooth.
+Selecting a goal near the previous one makes the robot avoid to many loops in the emerging graph.
 \subsection{GraphSLAM}
+\label{sec:implSLAM}
 Two different approaches of graph-based SLAM are implemented for this project. They differ in complexity and thereby applicability to the problem. Both algorithms create an occupancy grid map. 
-
+\begin{figure}[h]
+	\centering
+		\includegraphics[width=0.50\textwidth]{figures/graph.png}
+	\caption{An example of a simple graph (based on \citep{Grisetti}).}
+	\label{fig:turtlebot}
+\end{figure}
 First, an algorithm as suggested by \citep{Grisetti} is implemented.
@@
 Odometry and sensor measurements are incorporated into the constraints contained in the edges.
-Again, an edge connects two adjacent poses of the robot.
-This graph-based SLAM approach optimizes the pose graph according to the constraints.
+Similar to the graphSLAM procedure explained in Section \ref{subsec:sotaSlam}, an edge connects two adjacent poses of the robot.
+This implementation optimizes the pose graph according to the constraints.
 This ``optimized" trajectory then determines the positions of obstacles on the map by refining the laser scan data according to the poses they were obtained at.
@@
 Let this error be denoted by \(e_{ij}(x_i, x_j)\).
+This information is obtained from a combination of odometry data and laser range sensor data.
+The odometry is corrected using a laser scan matcher (see Subsection \ref{sec:scan}).
+
 This graph-based SLAM approach calculates the most likely path as follows:
-First it calculates the Jacobians \(A\) and \(B\) of every two estimated positions connected by an edge.
+First it finds the Jacobians \(A\) and \(B\) of every two estimated positions connected by an edge.
 \begin{equation}
@@
 Note that in Equations \ref{eq:jacobi_a} and \ref{eq:jacobi_b} the x represents the current guess of the concerning robot position.
-These Jacobians are then added to an information matrix \(H\) (see Equation \ref{eq:H}).
+As explained above, \(e_{ij}(x_i, x_j)\) is the error between the two poses \(x_i\) and \(x_j\).
+The Jacobians \(A\) and \(B\) are then added to an information matrix \(H\) (Equation \ref{eq:H}).
 \begin{equation}
@@
 \end{equation}
+In this equation $\Omega_{ij}$ is the information matrix of a measurement between \(x_i\) and \(x_j\).
+\\
 Furthermore a coefficient vector \(b\) is always updated as shown in Equation \ref{eq:b}. This vector in combination with the \(H\) matrix form a set of linear constraints.
@@
 \end{equation}
-Optimizing the vector \(x = (x_1, ..., x_n)^T \) according to the constraints specified in \(H\) and \(b\) is the core of this approach. This step is performed using sparse Cholesky factorization. 
-A new node is added to the graph once the robot moved 0.5 meters.
-The algorithm optimizes the trajectory like explained when a new node is added.
-In one special case the algorithm adds and edge connecting to non-adjacent nodes to the graph.
+The optimization of the vector \(x = (x_1, ..., x_n)^T \) according to the constraints specified in \(H\) and \(b\) is the core of this approach.
+This step is performed using sparse Cholesky factorization.
+A new node is added to the graph once the robot moves 0.5 meters. The algorithm optimizes the trajectory like explained once a new node is added.\\
+In one special case the algorithm adds an extra edge, connecting to non-adjacent nodes, to the graph.
 This happens when the distance between the two connected robot positions is smaller than a margin.
@@
 Each occupancy grid cell containing an obstacle is considered a map feature.
+This is because the map features are all that is added to the occupancy grid in the end.
 Edges are labelled with spatial constraints of the two nodes they are connecting.
 The output of this implementation is a vector containing the positions of the land marks as well as the trajectory of the robot.
-Here the constraints labelling the edges ar created using Jacobians similar to the previous approach.
+Here the constraints labelling the edges are created using Jacobians similar to the previous approach.
 The major difference lies in the node creation.
 Here nodes are added for landmarks while the previous implementation considered the robots' trajectory only.
-Since no obstacle in the final occupancy grid must be missed all obstacle points are treated as landmarks for this implementation.
-\subsubsection{Scan matching}
+An additional difference lies in the fact that this implementation uses features to characterize the map.
+The algorithm needs to align each new observed feature with the set of features already seen.
+This matching step has to be performed and repeated along with the complete optimization procedure until convergence.
 
-The scan matcher, used by this program, is necessary to acquire an optimized movement vector between two robot positions, which is needed by the GraphSLAM algorithm.
-The programming of the scan matcher used requires two steps.
-For the first step the two scans are matched together by making use of raw odometry data, retrieved by the robot's motion sensor.
-The improvement done by the scan matcher is then to run an iteration of the iterative closest point algorithm (ICP), to incorporate its perception with the movement.
+\subsection{Scan matching}
+\label{sec:scan}
+The scan matcher, used by this program, is necessary to acquire an optimized movement vector between two robot positions.
+This is needed by the GraphSLAM algorithm (see Subsection \ref{sec:implSLAM}).
+The scan matcher realized for this project can be divided into two steps.
+First, the two scans are matched making use of raw odometry data, retrieved by the robot's motion sensor.
+This is improved upon by running an iteration of the iterative closest point algorithm (ICP).
+It tries to align the two scans minimizing the total distance of each point in the first scan to its nearest neighbour in the second.
 
-The general procedure of the algorithm is to first find corresponding neighbors.
-These are found by using the nearest neighbor search, through which adjacent points are found by calculating two points from different sets according to their euclidean distance to one another.
-Points in this implementation can have multiple correspondences to neighbors.
-To avoid incorporating outlying points, a fixed distance has been set.
-Outliers would otherwise impact further calculations negatively.
-The next step in the process is to calculate center points for both clouds, which are needed to normalize them in respect ot each other.
-After performing the calculation the center points for each cloud are subtracted from every point belonging to that cloud.
-The calculated points are then added to two separate matrices, both representing one point cloud as a $2xn$ Matrix, where n is the number of neighbor pairs.
-The first cloud's matrix is then multiplied by the transposition of the second cloud matrix, so that a new matrix is created for which we calculate the Singular Value Decomposition (SVD) values to retrieve a rotational matrix, responsible for rotating the second point cloud to the first one.
-The next step is to translate the second point cloud to the first one with the use of a translation vector.
-Both rotation matrix and translation vector are then passed to a matrix containing both of them as well as a third line with the values $0 0 1$, for symmetry purposes.
-That matrix is then used to adjust the second point cloud to the first one.
-It is also used to correct the robot position.
+The general procedure of the algorithm is to first find the nearest neighbouring point in the second scan for all point in the first. To achieve this nearest neighbour search is used.
+In this implementation multiple points can have the same corresponding neighbour-point.
+If there is no point from the other scan within a specified radius of one point, it is excluded.
+These outliers would influence further calculations negatively.
+The next step in the process is to calculate the centroid of both clouds.
+These are subtracted from every point in the concerning cloud.
+In that way the clouds are centred at the origin.
+The calculated points are then added to two separate $2xn$ matrices, where $n$ is the number of neighbour pairs.
+Next the rotation matrix which transforms one cloud to the other is needed.
+The Singular Value Decomposition (SVD) of the product of the two previously mentioned matrices returns three new matrices $U$, $\Sigma$ and $V$.
+The required rotation matrix R can be obtained as described in formula \ref{eq:RVUT}.
+\begin{equation}
+\label{eq:RVUT}
+ R = V * U^T
+\end{equation}
+The translation vector $t$ is the difference between the two centroids.
+$R$ and $t$ together form a rigid transformation.
+This transformation matches the two scans.
+Moreover it is used to correct the robot's odometry data.
 
@@
 
-For Mapping purposes an Occupancy grid is created on which information about obstacles, free/- and unknown space is saved. Using ROS's visualization program, RVIZ, the map can be visualized entirely.
+The map is internally represented as an occupancy grid.
+It consists of cells which hold information about obstacles, free/- and unknown space.
+Using ROS's visualization program, RViz, the map can be visualized on the screen.
 
-The occupancy grid is a structure that can be represented by a single dimensional array.
-Its size is responsible for the size of the grid and is often problem specific, especially for the purpose of Slam.
-Making the grid too large can lead to performance problems and by that the grid's size gets to be a tunable parameter that has to be chosen carefully for each problem, although making the grid large prevents the robot from going ``out of space''.
- This program converts the array into a two dimensional representation, so that navigating to either side gets more intuitive.
-Each value of the array is an integer and represents the current status of the cell.
-The value ranges for ROS's occupancy grid lie between $-1$ and $100$ \cite{occupancy}.
-Each value has its own meaning :
-
+The occupancy grid is a structure that is represented by a one-dimensional array.
+The size of an occupancy grid has to be tuned problem specific.
+A too large grid can lead to performance problems.
+If the grid is too small, however, the map will be inaccurate.
+Fitting a map on a smaller grid results in a low resolution.
+Internally the program represents the grid as a two dimensional array.
+This makes navigating to either side more intuitive.
+Each grid cell contains an integer value.
+Different values represent the states a cell can be in.
+The value-range for ROS's occupancy grid lies between $-1$ and $100$ \cite{occupancy}.
+Each value has a specific meaning:
 \begin{description}
-\item{$-1$} : unknown space, not yet covered by robot's perception sensors
-\item{$ 0$} : known free space, robot can drive on it safely
-\item{$1-100$} : probability of being an obstacle, $1$ being a low probability and $100$ being a definite obstacle
+\item{$-1$} : Unknown space, not yet covered by robot's perception sensors
+\item{$0$} : Known free space, robot can drive on it safely
+\item{$1-100$} : Probability of being an obstacle, $1$ being a low probability and $100$ being a definite obstacle.
 \end{description}
-
 \begin{figure}[htbp]
 	\centering
-		\includegraphics[width=0.50\textwidth,height = 44mm]{figures/Occup.png}
-	\caption{Occupancy Grid visualized}
+		\includegraphics[width=0.45\textwidth]{figures/Occup.png}
+	\caption{A visualized occupancy grid.}
 	\label{fig:Occupancy}
 \end{figure}
+Figure \ref{fig:Occupancy} shows an example occupancy grid.
+The robot is standing in the middle.
+It senses the state of all cells within a distance of one.
+The border cells are marked with a $-1$, which means that they are still unexplored.
+The remaining cells are explored being either an obstacle or free space.
+The occupancy grid is created using laser scan data.
+Each laser scan contains information about whether or not an obstacle has been detected at a certain bearing.
+Maximum range readings occur when the robot senses only free space in a certain direction.
+All laser scans are transformed into point clouds.
+The absolute locations of the points are dependent on the position at which the laser scan was recorded.
+In the occupancy grid detected obstacle points are indicated by value 100.
+Free space is being incorporated into the map by filling regions between detected points and the concerning robot position. This is achieved with the help of a linear function.
 
-Fig.\ref{fig:Occupancy} shows these three occurrences visualized, with a robot in the middle performing a sample scan of its environment.
-The occupancy grid is using laser scan data for this implementation.
-Each laser scan has information about whether an obstacle has been detected, or whether the maximum range of the scan has been reached.
-The scan is then transformed into a point cloud to extract its coordinate information and by that, obstacles can be found.
-
-Free space is being incorporated into the map by filling the occupancy grid between a detected obstacle, or detected maximum range point and the current robot position with $0$ values.
-This is achieved with the help of a linear function.
-If a point is not the scanner's maximum range away, the point is saved as an obstacle with value $100$.
-
-As time progresses and the robot explores more of its environment the map is constantly updated.
-This is programmed by saving laser scans for every pose of the robot, which, when traveling for a distance of $0.5$ meters is added and by that the occupancy grid is updated.
-Upon a loop closure the occupancy grid is optimized with its information about obstacles and free spaces.
+As time progresses, the robot explores more of its environment.
+As explained above, the locations of the observed obstacles need to be extracted from the laser-scan data.
+The program keeps track of the scan-data for every robot position saved in the trajectory \(x = (x_1, ..., x_n)^T \).
+Since the laser-scan data is relative to the position of recording, new information is added to the map once the robot's trajectory is adjusted by the graphSLAM module.
+When a new scan is made, it is added to the occupancy grid after the trajectory has been updated.
+When a loop closure is detected, the complete occupancy grid is updated.
+All previous scan points are refined relative to their corresponding recording positions and the grid is regenerated.
 
@@
 
-\subsection{Runtime}
+\subsection{Runtime analysis}
+The implemented Graph SLAM algorithm has been tested on its feasibility.
+Navigating the robot manually, the complete test map has been explored, building up a pose graph of 123 nodes and 150 constraints.
+While doing that, the runtime of the complete update function was measured in microseconds.
+The update function is executed whenever a new node in the graph is created.
+It contains the algorithms for scan matching and Graph SLAM correction, which have been measured separately.
 
+Figure \ref{fig:runtime} shows the outcome of this runtime measurement. Naturally the runtime increases with increasing graph size, especially due to the growing matrices in the Graph SLAM algorithm.
+
+\begin{figure}[htbp]
+	\centering
+		\includegraphics[width=0.50\textwidth]{figures/Runtime.png}
+	\caption{Runtime of the Graph SLAM update function in micro seconds}
+	\label{fig:runtime}
+\end{figure}
+
+Even with a graph size of over 100 nodes, the runtime of the function stays in a very feasible range between 60 and 80 milliseconds.
+The robot can be moved without any delay.
+Taking only the Graph SLAM update into account, the runtime even stays below 20 milliseconds.
+In contrast the scan matching is executed outstandingly quickly since it is independent of the size of the pose graph.
+Most of the time is occupied by updating the occupancy grid because every single grid cell has to redrawn for every pose.
+In order to compromise this, the update of the complete occupancy grid is only executed every third second and only if a loop closing is detected.
 \section{Discussion}

### 2013-01-21 wfd_exploration/include/wfd.hpp
Added everything for the exploration package.

@@ -1,0 +1,13 @@
+#include "ros/ros.h"
+#include <cstdlib> // Needed for rand()
+#include <ctime>
+#include <queue>
+#include "nav_msgs/OccupancyGrid.h"
+
+
+
+std::vector<std::vector<int> > frontierDetection(const nav_msgs::OccupancyGrid& grid, int x, int y);
+bool isFrontier(const nav_msgs::OccupancyGrid& grid, int x, int y);
+std::vector<int> getNeighbors(const nav_msgs::OccupancyGrid& grid, int x, int y);
+int getMapValue(const nav_msgs::OccupancyGrid& grid, int x, int y);
+std::vector<int> getSurrounding(const nav_msgs::OccupancyGrid& grid, int x, int y, int size);

### 2013-01-21 wfd_exploration/src/exploration.cpp

@@ -1,0 +1,334 @@
+#include "ros/ros.h"
+#include "ros/console.h"
+#include "geometry_msgs/Twist.h"
+#include "sensor_msgs/LaserScan.h"
+#include <cstdlib> // Needed for rand()
+#include <ctime> // Needed to seed random number generator with a time value
+#include <geometry_msgs/PointStamped.h>
+#include <tf/transform_listener.h>
+#include <std_msgs/String.h>
+#include "geometry_msgs/Pose.h"
+#include "nav_msgs/Odometry.h"
+#include <tf/transform_listener.h>
+#include <tf/transform_broadcaster.h>
+#include <geometry_msgs/Vector3.h>
+#include <std_msgs/Float64.h>
+#include <nav_msgs/OccupancyGrid.h>
+#include <std_msgs/Float64MultiArray.h>
+#include <list>
+#include <vector>
+#include <wfd.hpp>
+#include <math.h>
+// Include for Ocupancy Grid
+#include <nav_msgs/OccupancyGrid.h>
+
+//includes for navigation stack
+#include <move_base_msgs/MoveBaseAction.h>
+#include <actionlib/client/simple_action_client.h>
+#include <tf/transform_broadcaster.h>
+typedef actionlib::SimpleActionClient<move_base_msgs::MoveBaseAction> MoveBaseClient;
+
+class Exploration {
+
+public:
+    Exploration(ros::NodeHandle& nh) {
+        // Initialize random time generator
+        srand(time(NULL));
+        robotX = 0;
+        robotY = 0;
+        commandPub = nh.advertise < geometry_msgs::Twist > ("cmd_vel", 1);
+        onTheMove = false;
+        mapSub = nh.subscribe("/world", 1, &Exploration::occupancyGridCallback,
+                              this);
+        velSub = nh.subscribe("odom", 5, &Exploration::vel_callback, this);
+        frontier_publisher = nh.advertise < sensor_msgs::PointCloud> ("frontiers", 1);
+        next_frontier_publisher = nh.advertise < sensor_msgs::PointCloud> ("next_frontier", 1);
+        next_frontier.header.frame_id = "/world";
+        frontier_cloud.header.frame_id = "/world";
+    }
+
+    //Tell the navigation stack where to run to
+    void run(double x, double y){
+        onTheMove = true;
+        MoveBaseClient ac("move_base", true);
+
+        while(!ac.waitForServer(ros::Duration(5.0))){
[…]

### 2013-01-21 trunk/wfd_exploration/src/wfd.cpp
Added everything for the exploration package.

@@ -1,0 +1,249 @@
+#include "ros/ros.h"
+#include <cstdlib>
+#include <ctime>
+#include <queue>
+#include <map>
+#include "nav_msgs/OccupancyGrid.h"
+#include <wfd.hpp>
+#include <vector>
+
+const int MAP_open_list = 1;
+const int MAP_close_list = 2;
+const int FRONTIER_open_list = 3;
+const int FRONTIER_close_list = 4;
+
+// WFD implementation
+	std::vector<std::vector<int> > frontierDetection(const nav_msgs::OccupancyGrid& map, int x, int y) {
+//		ROS_INFO("FRONTIER DETECTION START....");
+		std::vector<std::vector<int> > frontierList; 
+		std::vector<int> pose;
+		std::vector<std::vector<int> > marker;
+		for(int i = 0 ; i < map.info.width ; i++) {
+			std::vector<int> row;
+			for(int j = 0 ; j < map.info.height ; j++) {
+				row.push_back(-1);
+			}
+			marker.push_back(row);
+		}
+		pose.push_back(x); // x-coordinate robot
+		pose.push_back(y); // y-coordinate robot
+		//create neccessary lists
+		std::vector<std::vector<int> > queue_m;
+//		ROS_INFO("queue_m first push pose[0] = %i , pose[1] = %i", pose[0], pose[1]);
+		queue_m.push_back(pose);
+		int index1 = pose[0];
+		int index2 = pose[1];
+		marker[index1][index2] = MAP_open_list; 
+		while (!queue_m.empty()) {
+			//ROS_INFO("Loop starting");
+			// dequeue
+			std::vector<int> p = queue_m.front();
+			queue_m.erase(queue_m.begin());
+			//ROS_INFO("if map close list contains p, continue");
+			//if map close list contains p, continue
+			index1 = p[0];
+			index2 = p[1];
+			if(marker[index1][index2] == MAP_close_list) {	
+				//ROS_INFO("CONTINUE!!!!! NO1");
+				continue;
+			}//*/
+			//ROS_INFO("if(isFrontier(p[0], p[1])){");
+			if (isFrontier(map, p[0], p[1])) {
+				std::vector<std::vector<int> > queue_f;
+				std::vector<int> newFrontier;
+				queue_f.push_back(p);
+				marker[index1][index2] = FRONTIER_open_list;	
+
+		
+				while (!queue_f.empty()) {
+					//ROS_INFO("Second queue starting");
[…]

### 2013-01-21 wfd_exploration/src/exploration.cpp
Some refactoring.

@@ -19,10 +19,7 @@
 #include <vector>
 #include <wfd.hpp>
 #include <math.h>
-// Include for Ocupancy Grid
 #include <nav_msgs/OccupancyGrid.h>
-
-//includes for navigation stack
 #include <move_base_msgs/MoveBaseAction.h>
 #include <actionlib/client/simple_action_client.h>
 #include <tf/transform_broadcaster.h>
@@ -82,7 +79,7 @@ public:
         // Get the current position of the robot
         robotX = msg.pose.pose.position.x;
         robotY = msg.pose.pose.position.y;
-//        ROS_INFO_STREAM("Robot position : ("<< robotX <<", "<<robotY<<").");
+        //        ROS_INFO_STREAM("Robot position : ("<< robotX <<", "<<robotY<<").");
     }
 
     // get the grid from gmapping by listener
@@ -93,7 +90,7 @@ public:
 
     bool isBlocked(double x, double y)
     {
-        for(int i = 0; i<blocked.size(); i++)
+        for(unsigned int i = 0; i<blocked.size(); i++)
         {
             if(blocked[i][0] == x && blocked[i][1] == y)
                 return true;
@@ -117,13 +114,9 @@ public:
                 double x = transform.getOrigin().x();
                 double y = transform.getOrigin().y();
 
-//                ROS_INFO_STREAM("STARTPOSITION X: " <<x<<" STARTPOSITION Y: " <<y);
                 robot_pos[0] = (int) ((x / resolution) + (int)(occupancyGrid.info.width/2));
                 robot_pos[1] = (int) ((y / resolution) + (int)(occupancyGrid.info.height/2));
-//                ROS_INFO("X = %i , Y = %i", robot_pos[0], robot_pos[1]);
-//                ROS_INFO("LETS GET THE FRONTIERS");
                 std::vector<std::vector<int> > frontiersList = frontierDetection(occupancyGrid, robot_pos[0], robot_pos[1]);
-//                ROS_INFO("FRONIERS SERVED!");
 
                 x = robotX;
                 y = robotY;
@@ -136,86 +129,66 @@ public:
                     }
                 }
 
-//                ROS_INFO("%i frontier cells found",num_frontier_cells);
                 double goalX = 0;
                 double goalY = 0;
                 frontier_cloud.points.resize(0);
                 frontier_cloud.points.resize(num_frontier_cells);
                 int frontierIndex = 0;
-                //arbitrary value which is always higher than any distance found
-                double minDist = 40000.0;
[…]

### 2013-01-21 trunk/wfd_exploration/src/wfd.cpp
Some refactoring.

@@ -12,226 +12,220 @@ const int MAP_close_list = 2;
 const int FRONTIER_open_list = 3;
 const int FRONTIER_close_list = 4;
 
-// WFD implementation
-	std::vector<std::vector<int> > frontierDetection(const nav_msgs::OccupancyGrid& map, int x, int y) {
-//		ROS_INFO("FRONTIER DETECTION START....");
-		std::vector<std::vector<int> > frontierList; 
-		std::vector<int> pose;
-		std::vector<std::vector<int> > marker;
-		for(int i = 0 ; i < map.info.width ; i++) {
-			std::vector<int> row;
-			for(int j = 0 ; j < map.info.height ; j++) {
-				row.push_back(-1);
-			}
-			marker.push_back(row);
-		}
-		pose.push_back(x); // x-coordinate robot
-		pose.push_back(y); // y-coordinate robot
-		//create neccessary lists
-		std::vector<std::vector<int> > queue_m;
-//		ROS_INFO("queue_m first push pose[0] = %i , pose[1] = %i", pose[0], pose[1]);
-		queue_m.push_back(pose);
-		int index1 = pose[0];
-		int index2 = pose[1];
-		marker[index1][index2] = MAP_open_list; 
-		while (!queue_m.empty()) {
-			//ROS_INFO("Loop starting");
-			// dequeue
-			std::vector<int> p = queue_m.front();
-			queue_m.erase(queue_m.begin());
-			//ROS_INFO("if map close list contains p, continue");
-			//if map close list contains p, continue
-			index1 = p[0];
-			index2 = p[1];
-			if(marker[index1][index2] == MAP_close_list) {	
-				//ROS_INFO("CONTINUE!!!!! NO1");
-				continue;
-			}//*/
-			//ROS_INFO("if(isFrontier(p[0], p[1])){");
-			if (isFrontier(map, p[0], p[1])) {
-				std::vector<std::vector<int> > queue_f;
-				std::vector<int> newFrontier;
-				queue_f.push_back(p);
-				marker[index1][index2] = FRONTIER_open_list;	
+// Algorithm for Wavefront Frontier Detection
+std::vector<std::vector<int> > frontierDetection(const nav_msgs::OccupancyGrid& map, int x, int y) {
 
-		
-				while (!queue_f.empty()) {
-					//ROS_INFO("Second queue starting");
-					std::vector<int> q = queue_f.front();
-					queue_f.erase(queue_f.begin());
[…]

### 2013-01-21 trunk/wfd_exploration/src/wfd.cpp
Finished refactoring.

@@ -154,9 +154,9 @@ bool isFrontier(const nav_msgs::OccupancyGrid& grid, int x, int y) {
     int numExplored = 0;
     int numUnknown = 0;
 
-    // checking neighbors
-    // if there are known and unknown neighbors
-    // the robot is at a frontiers
+    // Checking neighbors
+    // If there are known and unknown neighbors
+    // the robot is at a frontier
     bool obstacle = false;
     for (unsigned int i = 0; i < neighbors.size(); i += 2) {
         std::vector<int> current;
@@ -228,7 +228,7 @@ std::vector<int> getNeighbors(const nav_msgs::OccupancyGrid& grid, int x, int y)
 }
 
 
-// returns the value of position [x,y] from the occupancy grid
+// Returns the value of position [x,y] from the occupancy grid
 // the value represents the probability of an obstacle or frontier
 // at position [x,y]. Values are:
 // -1 = unknown

### 2013-01-21 wfd_exploration/src/exploration.cpp
Finished refactoring.

@@ -50,7 +50,7 @@ public:
         MoveBaseClient ac("move_base", true);
 
         while(!ac.waitForServer(ros::Duration(5.0))){
-            ROS_INFO("Waiting for the move_base action server to come up");
+            ROS_INFO("Waiting for the move_base action server...");
         }
 
         move_base_msgs::MoveBaseGoal goal;
@@ -62,15 +62,12 @@ public:
         goal.target_pose.pose.position.y = y;
         goal.target_pose.pose.orientation.w = 1;
 
-        ROS_INFO_STREAM("Moving to position x:"<<goal.target_pose.pose.position.x<<" y:"<<goal.target_pose.pose.position.y);
         ac.sendGoal(goal);
 
+        // Wait for either a result or for 3 seconds
+        // This way the robot is constantly heading for a new goal, avoiding unreachable points
         ac.waitForResult(ros::Duration(3));
 
-        if(ac.getState() == actionlib::SimpleClientGoalState::SUCCEEDED)
-            ROS_INFO_STREAM("Robot moved to position x:"<< goal.target_pose.pose.position.x<<" y:"<<goal.target_pose.pose.position.y);
-        else
-            ROS_INFO_STREAM("The base failed to move to position x:"<< goal.target_pose.pose.position.x<<" y:"<<goal.target_pose.pose.position.y);
         onTheMove = false;
     }
 
@@ -79,10 +76,8 @@ public:
         // Get the current position of the robot
         robotX = msg.pose.pose.position.x;
         robotY = msg.pose.pose.position.y;
-        //        ROS_INFO_STREAM("Robot position : ("<< robotX <<", "<<robotY<<").");
     }
 
-    // get the grid from gmapping by listener
     double prevGoalX;
     double prevGoalY;
     std::vector<std::vector<double> > blocked;
@@ -104,11 +99,11 @@ public:
             grid = occupancyGrid;
             float resolution = occupancyGrid.info.resolution;
             tf::TransformListener listener(ros::Duration(10));
-            //transform object storing our robot's position
+
+            // Transform object storing the robots position
             tf::StampedTransform transform;
             try {
                 ros::Time now = ros::Time::now();
-                geometry_msgs::PointStamped base_point;
                 listener.waitForTransform("/world", "/base_link", now, ros::Duration(0.5));
                 listener.lookupTransform("/world", "/base_link", ros::Time(0), transform);
                 double x = transform.getOrigin().x();
[…]

### 2013-01-21 turtlebot_graph_slam/src/exploration.cpp
Launching wfd_exploration package. Deleted wfd and exploration stuff.

@@ -1,218 +1,0 @@
-#include "ros/ros.h"
-#include "ros/console.h"
-#include "geometry_msgs/Twist.h"
-#include "sensor_msgs/LaserScan.h"
-#include <cstdlib> // Needed for rand()
-#include <ctime> // Needed to seed random number generator with a time value
-#include <geometry_msgs/PointStamped.h>
-#include <tf/transform_listener.h>
-#include <std_msgs/String.h>
-#include <tf/transform_listener.h>
-#include <tf/transform_broadcaster.h>
-#include <geometry_msgs/Vector3.h>
-#include <std_msgs/Float64.h>
-#include <nav_msgs/OccupancyGrid.h>
-#include <std_msgs/Float64MultiArray.h>
-#include <list>
-#include <vector>
-#include <wfd.hpp>
-#include <math.h>
-// Include for Ocupancy Grid
-#include <nav_msgs/OccupancyGrid.h>
-
-//includes for navigation stack
-#include <move_base_msgs/MoveBaseAction.h>
-#include <actionlib/client/simple_action_client.h>
-#include <tf/transform_broadcaster.h>
-typedef actionlib::SimpleActionClient<move_base_msgs::MoveBaseAction> MoveBaseClient;
-
-class Exploration {
-
-public:
-	Exploration(ros::NodeHandle& nh) {
-		// Initialize random time generator
-		srand(time(NULL));
-
-		commandPub = nh.advertise < geometry_msgs::Twist > ("cmd_vel", 1);
-
-		mapSub = nh.subscribe("/world", 1, &Exploration::occupancyGridCallback,
-				this);
-
-		frontier_publisher = nh.advertise < sensor_msgs::PointCloud> ("frontiers", 1);
-		next_frontier_publisher = nh.advertise < sensor_msgs::PointCloud> ("next_frontier", 1);
-        next_frontier.header.frame_id = "/world";
-		frontier_cloud.header.frame_id = "/world";
-	}
-
-//Tell the navigation stack where to run to
-void run(double x, double y){
-    ROS_ERROR("Run is called in turtlebot_graph_slam");
-	//@TODO make a turn
-
-          //tell the action client that we want to spin a thread by default
-  MoveBaseClient ac("move_base", true);
-
-  //wait for the action server to come up
-  while(!ac.waitForServer(ros::Duration(5.0))){
-  	  ROS_INFO("Waiting for the move_base action server to come up");
-  }
-
-  move_base_msgs::MoveBaseGoal goal;
-
[…]

### 2013-01-21 turtlebot_graph_slam/src/test.cpp

@@ -19,7 +19,10 @@
 #include <lasertrans.hpp>
 #include <geometry_msgs/PoseArray.h>
 #include <visualization_msgs/Marker.h>
-
+#include <iostream>
+#include <ctime>
+#include <sys/time.h>
+using namespace std;
 using namespace Eigen;
 
 /**
@@ -63,7 +66,6 @@ bool loopDetected;
 
 // Distance after which a laser scan should be matched again
 double MATCH_DISTANCE = 0.5;
-
 void callback(const sensor_msgs::LaserScan::ConstPtr& msg) {
     /**
       Simply save the current scan.
@@ -210,6 +212,10 @@ void vel_callback(const nav_msgs::Odometry& msg) {
 
     // If the robot has walked far enough to make the second scan
     if(distance(prevX,newX,prevY,newY)>= MATCH_DISTANCE /*|| rotation(prevZ, newZ)*/) {
+        std::cout << "Nodes = " << nodes.size();
+        std::cout << ", Constraints = " << constraints.size() << std::endl;
+        timespec beginTimeStep;
+           clock_gettime(CLOCK_REALTIME, &beginTimeStep);
 
         // Create new node for current robot pose
         Vector3d newNode;
@@ -254,8 +260,12 @@ void vel_callback(const nav_msgs::Odometry& msg) {
                 prevPointCloud = transformCloud(prevPointCloud, xX, xY, xYaw);
 
                 // Use scanmatch for both clouds
-                Vector3d correction = scanmatch(prevPointCloud, currentPointCloud);
-
+                timespec beginScanMatch;
+                   clock_gettime(CLOCK_REALTIME, &beginScanMatch);
+                   Vector3d correction = scanmatch(prevPointCloud, currentPointCloud);
+                timespec endScanMatch;
+                   clock_gettime(CLOCK_REALTIME, &endScanMatch);
+                std::cout << "ScanMatch duration: " << (endScanMatch.tv_nsec-beginScanMatch.tv_nsec)/1000 << std::endl;
                 // In order to test Graph SLAM with the actual odometry, with random noise
                 Vector3d zTransformation;
                 zTransformation(0) = xX+correction(0);
@@ -267,18 +277,25 @@ void vel_callback(const nav_msgs::Odometry& msg) {
                 c.i = i;
                 c.j = j;
                 c.z = zTransformation;
[…]

### 2013-01-21 trunk/turtlebot_graph_slam/include/wfd.hpp
Launching wfd_exploration package. Deleted wfd and exploration stuff.

@@ -1,249 +1,0 @@
-#include "ros/ros.h"
-#include <cstdlib>
-#include <ctime>
-#include <queue>
-#include <map>
-#include "nav_msgs/OccupancyGrid.h"
-#include <wfd.hpp>
-#include <vector>
-
-const int MAP_open_list = 1;
-const int MAP_close_list = 2;
-const int FRONTIER_open_list = 3;
-const int FRONTIER_close_list = 4;
-
-// WFD implementation
-	std::vector<std::vector<int> > frontierDetection(const nav_msgs::OccupancyGrid& map, int x, int y) {
-		//ROS_INFO("FRONTIER DETECTION START....");
-		std::vector<std::vector<int> > frontierList; 
-		std::vector<int> pose;
-		std::vector<std::vector<int> > marker;
-		for(int i = 0 ; i < map.info.width ; i++) {
-			std::vector<int> row;
-			for(int j = 0 ; j < map.info.height ; j++) {
-				row.push_back(-1);
-			}
-			marker.push_back(row);
-		}
-		pose.push_back(x); // x-coordinate robot
-		pose.push_back(y); // y-coordinate robot
-		//create neccessary lists
-		std::vector<std::vector<int> > queue_m;
-		//ROS_INFO("queue_m first push pose[0] = %i , pose[1] = %i", pose[0], pose[1]);
-		queue_m.push_back(pose);
-		int index1 = pose[0];
-		int index2 = pose[1];
-		marker[index1][index2] = MAP_open_list; 
-		while (!queue_m.empty()) {
-			//ROS_INFO("Loop starting");
-			// dequeue
-			std::vector<int> p = queue_m.front();
-			queue_m.erase(queue_m.begin());
-			//ROS_INFO("if map close list contains p, continue");
-			//if map close list contains p, continue
-			index1 = p[0];
-			index2 = p[1];
-			if(marker[index1][index2] == MAP_close_list) {	
-				//ROS_INFO("CONTINUE!!!!! NO1");
-				continue;
-			}//*/
-			//ROS_INFO("if(isFrontier(p[0], p[1])){");
-			if (isFrontier(map, p[0], p[1])) {
-				std::vector<std::vector<int> > queue_f;
-				std::vector<int> newFrontier;
-				queue_f.push_back(p);
-				marker[index1][index2] = FRONTIER_open_list;	
-
-		
-				while (!queue_f.empty()) {
-					//ROS_INFO("Second queue starting");
[…]

### 2013-01-21 trunk/turtlebot_graph_slam/src/wfd.cpp

@@ -1,249 +1,0 @@
-#include "ros/ros.h"
-#include <cstdlib>
-#include <ctime>
-#include <queue>
-#include <map>
-#include "nav_msgs/OccupancyGrid.h"
-#include <wfd.hpp>
-#include <vector>
-
-const int MAP_open_list = 1;
-const int MAP_close_list = 2;
-const int FRONTIER_open_list = 3;
-const int FRONTIER_close_list = 4;
-
-// WFD implementation
-	std::vector<std::vector<int> > frontierDetection(const nav_msgs::OccupancyGrid& map, int x, int y) {
-		ROS_INFO("FRONTIER DETECTION START....");
-		std::vector<std::vector<int> > frontierList; 
-		std::vector<int> pose;
-		std::vector<std::vector<int> > marker;
-		for(int i = 0 ; i < map.info.width ; i++) {
-			std::vector<int> row;
-			for(int j = 0 ; j < map.info.height ; j++) {
-				row.push_back(-1);
-			}
-			marker.push_back(row);
-		}
-		pose.push_back(x); // x-coordinate robot
-		pose.push_back(y); // y-coordinate robot
-		//create neccessary lists
-		std::vector<std::vector<int> > queue_m;
-		ROS_INFO("queue_m first push pose[0] = %i , pose[1] = %i", pose[0], pose[1]);
-		queue_m.push_back(pose);
-		int index1 = pose[0];
-		int index2 = pose[1];
-		marker[index1][index2] = MAP_open_list; 
-		while (!queue_m.empty()) {
-			//ROS_INFO("Loop starting");
-			// dequeue
-			std::vector<int> p = queue_m.front();
-			queue_m.erase(queue_m.begin());
-			//ROS_INFO("if map close list contains p, continue");
-			//if map close list contains p, continue
-			index1 = p[0];
-			index2 = p[1];
-			if(marker[index1][index2] == MAP_close_list) {	
-				//ROS_INFO("CONTINUE!!!!! NO1");
-				continue;
-			}//*/
-			//ROS_INFO("if(isFrontier(p[0], p[1])){");
-			if (isFrontier(map, p[0], p[1])) {
-				std::vector<std::vector<int> > queue_f;
-				std::vector<int> newFrontier;
-				queue_f.push_back(p);
-				marker[index1][index2] = FRONTIER_open_list;	
-
-		
-				while (!queue_f.empty()) {
-					//ROS_INFO("Second queue starting");
[…]

### 2013-01-21 Article/Article.tex
Updated the first section to include the needed references

@@
 Some even have devices similar to the human arm or hand which allows them to interact with the environment.
-The vaccuum cleaning robot presented in Figure \ref{fig:vacuum_cleaner} is an example of a robot which has whells to move around.
+The vacuum cleaning robot presented in Figure \ref{fig:vacuum_cleaner} is an example of a robot which has wheels to move around.
 Furthermore it is equipped with a set of sensors.
@@
 
-The acquisition of these maps has been a major research focus in the robotics community \cite{Grisetti} [\textbf{MORE SOURCES}].
+The acquisition of these maps has been a major research focus in the robotics community \cite{Thrun,Grisetti,Montemerlo02,Montemerlo}.
 Building or learning these maps under position uncertainty provided only sensor measurements is often referred to as the Simultaneous Localization And Mapping (SLAM) problem.
-A large variety of solutions is available to the SLAM problem, these approaches can be classified as either filtering or smoothing [\textbf{SOURCE}].
+A large variety of solutions is available to the SLAM problem, these approaches can be classified as either filtering or smoothing.
 
@@
 The other type of approach estimates the full trajectory from the full set of measurements.
-These approaches address the so called full SLAM problem and they typically rely on least-square error minimization techniques [\textbf{SOURCE}].
+These approaches address the so called full SLAM problem and they typically rely on least-square error minimization techniques \cite{Leastsquares}.
 
@@
 The edges encode sensor measurements that constraint the connected poses.
-Once such a graph is constructed the graph-based SLAM method tries to find a configuration of the nodes that is maximally consistent with the contraints contained in the edges.
-(\textbf{ABSTRACT GRAPH PICTURE HERE})
+Once such a graph is constructed the graph-based SLAM method tries to find a configuration of the nodes that is maximally consistent with the constraints contained in the edges.
 
@@
 	\caption{An example of a simple graph (based on \citep{Grisetti}).}
-	\label{fig:turtlebot}
+	\label{fig:example_graph}
 \end{figure}

### 2013-01-21 trunk/turtlebot_graph_slam/src/GraphBasedAlgo1.cpp
Some minor refactoring.

@@ -8,119 +8,91 @@
 using namespace std;
 using namespace Eigen;
 
+/**
+  Graph SLAM algorithm for correcting the pose vector
+  */
 std::vector<Vector3d> algorithm1(std::vector<Vector3d> x, std::vector<Constraint> constraints) {
 
-	// find the maximum likelihood solution
-	bool converged = false;
+    VectorXd b = VectorXd::Zero(3*x.size(),1);
+    MatrixXd H = MatrixXd::Zero(3*x.size(),3*x.size());
+    for(int constraintIndex = 0; constraintIndex < constraints.size(); constraintIndex++)
+    {
+        Constraint c = constraints[constraintIndex];
 
-//	VectorXd b(x.rows(),1);
-//	MatrixXd H(x.rows(),x.rows());
+        // Get the indices of the nodes this constraint refers to
+        int i = c.i;
+        int j = c.j;
 
-//	while(!converged)
-//	{
-        VectorXd b = VectorXd::Zero(3*x.size(),1);
-        MatrixXd H = MatrixXd::Zero(3*x.size(),3*x.size());
-        for(int constraintIndex = 0; constraintIndex < constraints.size(); constraintIndex++)
-		{
-//            std::cout << "constraintIndex = \n" << constraintIndex << std::endl;
-            Constraint c = constraints[constraintIndex];
+        Vector3d x_i = x[i];
+        Vector3d x_j = x[j];
+        Vector3d z_i_j = c.z;
 
-			// Get the indices of the nodes this constraint refers to
-            int i = c.i;
-            int j = c.j;
+        Vector2d t_i = x_i.block(0, 0, 2, 1);		// Position vector of node i
+        Vector2d t_j = x_j.block(0, 0, 2, 1);		// Position vector of node j
+        Vector2d t_i_j = z_i_j.block(0, 0, 2, 1);	// Measured position difference vector between nodes i and j
 
-//            std::cout << "i = \n" << i << std::endl;
-//            std::cout << "j = \n" << j << std::endl;
+        double phi_i = x_i(2,0);					// Angle of node i
[…]

### 2013-01-21 trunk/turtlebot_graph_slam/src/OccupancyGrid.cpp

@@ -23,18 +23,19 @@ nav_msgs::OccupancyGrid initializeOccupancyGrid(int length, double resolution) {
     og.header.stamp = ros::Time::now();
     og.info.width = SIZE;
     og.info.height = SIZE;
-    ROS_INFO_STREAM("MAP HEIGHT : " << og.info.height);
     og.data.resize(SIZE * SIZE);
 
-    for(int i = 0; i < og.data.size();i++){
-       og.data[i] = -1;
+    for(int i = 0; i < og.data.size();i++)
+    {
+        og.data[i] = -1;
     }
 
     return og;
 }
-//publish Occupacy grid
-void publishOccupancyGrid(nav_msgs::OccupancyGrid og,ros::Publisher occupub){
-	occupub.publish(og);
+// Publish Occupacy grid
+void publishOccupancyGrid(nav_msgs::OccupancyGrid og,ros::Publisher occupub)
+{
+    occupub.publish(og);
 }
 
 // Used for calculating the free space
@@ -51,111 +52,33 @@ int reformat(int i)
 {
     return (int)(i/RESOLUTION + SIZE/2);
 }
-    /*		Populate a Occupancy GridFF
-    --------------------------------------------------------------------------------------*/
+// Populate a Occupancy Grid
 nav_msgs::OccupancyGrid updateOccupancyGrid(nav_msgs::OccupancyGrid og, std::vector < sensor_msgs::LaserScan::ConstPtr > laserscansaver, std::vector<Vector3d> poses){
-    // ROS_INFO_STREAM("Occupancy Grid has height: "<<og.info.height<<" and width: "<<og.info.width);
 
-//	for(int i = 0 ; i < laserscansaver.size() ; i++){
-//        for(int i = laserscansaver.size() ; i < laserscansaver.size() ; i++){
-            int i = laserscansaver.size()-1;
-		sensor_msgs::PointCloud temp = lasertrans(laserscansaver[i]);
-//		if(i>= poses.size())
-//			break;
-        double xPose = poses[i](0,0);
-        double yPose = poses[i](1,0);
-        double yaw = poses[i](2,0);
+    int i = laserscansaver.size()-1;
+    sensor_msgs::PointCloud temp = lasertrans(laserscansaver[i]);
 
-        // Put the max range measurements in the occupancy grid
-            sensor_msgs::LaserScan::ConstPtr scan = laserscansaver[i];
-            int num_ranges = scan->ranges.size();
-            float incr = scan->angle_increment;
[…]

### 2013-01-21 trunk/turtlebot_graph_slam/src/lasertrans.cpp

@@ -57,13 +57,11 @@ sensor_msgs::PointCloud lasertrans (const sensor_msgs::LaserScan::ConstPtr& msg)
     try
     {
         projector_.projectLaser(*msg,cloud);   //Transform laser cloud to world frame
-
     }
     catch (...)
     {
         sensor_msgs::PointCloud error;
         return error;
-        // std::cout << e.what();                       //TODO Fix errors, they still occur decomment to see whats happening
     }
     return cloud;
 }

### 2013-01-21 trunk/turtlebot_graph_slam/src/robotpos.cpp

@@ -7,47 +7,45 @@ double robotyy = 0;
 double robotyyaw = 0;
 
 void robotpospub::robotpos(double x, double y, double roll, double pitch, double yaw) {
-	std::cout << "Robotpublisher created" << std::endl;
-	robotxx = x;
-	robotyy = y;
-	robotyyaw = yaw;
+
+    robotxx = x;
+    robotyy = y;
+    robotyyaw = yaw;
     int32_t publish_rate_ = 75;
-	tf::TransformBroadcaster tf_br_;
-	tf::StampedTransform tf_map_to_baselink_;
-	tf::TransformBroadcaster tf_br_1;
-	tf::StampedTransform tf_map_to_odom_;
-// set up parent and child frames
-	tf_map_to_baselink_.frame_id_ = std::string("world");
-	tf_map_to_baselink_.child_frame_id_ = std::string("base_link");
-	tf_map_to_odom_.frame_id_ = std::string("world");
-	tf_map_to_odom_.child_frame_id_ = std::string("odom");
+    tf::TransformBroadcaster tf_br_;
+    tf::StampedTransform tf_map_to_baselink_;
+    tf::TransformBroadcaster tf_br_1;
+    tf::StampedTransform tf_map_to_odom_;
 
-// set up publish rate
-	ros::Rate loop_rate(publish_rate_);
-//grid_x = (unsigned int)((map_x - map.info.origin.position.x) / map.info.resolution)
-//grid_y = (unsigned int)((map_y - map.info.origin.position.y) / map.info.resolution)
-// main loop
+    // Set up parent and child frames
+    tf_map_to_baselink_.frame_id_ = std::string("world");
+    tf_map_to_baselink_.child_frame_id_ = std::string("base_link");
+    tf_map_to_odom_.frame_id_ = std::string("world");
+    tf_map_to_odom_.child_frame_id_ = std::string("odom");
 
-	while (ros::ok()) {
-		// time stamp
-		tf_map_to_baselink_.stamp_ = ros::Time::now();
-		tf_map_to_odom_.stamp_ = ros::Time::now();
+    // Set up publish rate
+    ros::Rate loop_rate(publish_rate_);
 
-		// specify actual transformation vectors from odometry
-		// NOTE: zeros have to be substituted with actual variable data
-		tf_map_to_baselink_.setOrigin(tf::Vector3(robotxx, robotyy, 0.0f));
[…]

### 2013-01-21 trunk/turtlebot_graph_slam/src/test.cpp

@@ -184,7 +184,7 @@ void vel_callback(const nav_msgs::Odometry& msg) {
 
     // If this is the first call create the first node
     if(!initialised && currentScan){
-	
+
         // Save the current position as previous
         prevX = newX;
         prevY = newY;
@@ -212,10 +212,8 @@ void vel_callback(const nav_msgs::Odometry& msg) {
 
     // If the robot has walked far enough to make the second scan
     if(distance(prevX,newX,prevY,newY)>= MATCH_DISTANCE /*|| rotation(prevZ, newZ)*/) {
-        std::cout << "Nodes = " << nodes.size();
-        std::cout << ", Constraints = " << constraints.size() << std::endl;
-        timespec beginTimeStep;
-           clock_gettime(CLOCK_REALTIME, &beginTimeStep);
+        std::cout << "Nodes: " << nodes.size() << std::endl;
+        std::cout << "Constraints: " << constraints.size() << std::endl;
 
         // Create new node for current robot pose
         Vector3d newNode;
@@ -260,12 +258,8 @@ void vel_callback(const nav_msgs::Odometry& msg) {
                 prevPointCloud = transformCloud(prevPointCloud, xX, xY, xYaw);
 
                 // Use scanmatch for both clouds
-                timespec beginScanMatch;
-                   clock_gettime(CLOCK_REALTIME, &beginScanMatch);
-                   Vector3d correction = scanmatch(prevPointCloud, currentPointCloud);
-                timespec endScanMatch;
-                   clock_gettime(CLOCK_REALTIME, &endScanMatch);
-                std::cout << "ScanMatch duration: " << (endScanMatch.tv_nsec-beginScanMatch.tv_nsec)/1000 << std::endl;
+                Vector3d correction = scanmatch(prevPointCloud, currentPointCloud);
+
                 // In order to test Graph SLAM with the actual odometry, with random noise
                 Vector3d zTransformation;
                 zTransformation(0) = xX+correction(0);
@@ -290,12 +284,7 @@ void vel_callback(const nav_msgs::Odometry& msg) {
         }
 
         // Perform Graph SLAM optimization
-        timespec beginGraphSLAM;
-           clock_gettime(CLOCK_REALTIME, &beginGraphSLAM);
[…]

### 2013-01-21 turtlebot_graph_slam/src/feature_extractor.cpp
Some minor refactoring.

@@ -10,8 +10,7 @@
 #include <laser_geometry/laser_geometry.h>
 #include "nav_msgs/OccupancyGrid.h"
 using namespace Eigen;
-//https://gt-ros-pkg.googlecode.com/svn-history/r2368/trunk/hrl/simple_occupancy_grid/src/occupancy_grid.cpp
-//http://cs225turtle.googlecode.com/svn/trunk/project2/local_obstacles/src/local_obstacles.cpp
+
 int laserscancount = 0;
 double PI = 3.14;
 const static double MIN_SCAN_ANGLE_RAD = -2.0 / 180 * M_PI; //@TODO: Set the range
@@ -40,7 +39,6 @@ MatrixXd feature_extractor(const sensor_msgs::LaserScan::ConstPtr& msg,
 			(MIN_SCAN_ANGLE_RAD - msg->angle_min) / msg->angle_increment);
 	unsigned int maxIndex = ceil(
 			(MAX_SCAN_ANGLE_RAD - msg->angle_min) / msg->angle_increment);
-	float closestRange = msg->ranges[minIndex];
 	std::vector<int> laserpose;
 	int lasercount = maxIndex - minIndex;
 

### 2013-01-21 trunk/turtlebot_graph_slam/src/test.cpp
Deleted comment.

@@ -211,7 +211,7 @@ void vel_callback(const nav_msgs::Odometry& msg) {
     }
 
     // If the robot has walked far enough to make the second scan
-    if(distance(prevX,newX,prevY,newY)>= MATCH_DISTANCE /*|| rotation(prevZ, newZ)*/) {
+    if(distance(prevX,newX,prevY,newY)>= MATCH_DISTANCE) {
         std::cout << "Nodes: " << nodes.size() << std::endl;
         std::cout << "Constraints: " << constraints.size() << std::endl;
 

### 2013-01-21 trunk/Article/Article.tex
article update

@@
 
-The acquisition of these maps has been a major research focus in the robotics community \cite{Grisetti} [\textbf{MORE SOURCES}].
+The acquisition of these maps has been a major research focus in the robotics community \cite{Thrun, Grisetti, Montemerlo02, Montemerlo}.
 Building or learning these maps under position uncertainty provided only sensor measurements is often referred to as the Simultaneous Localization And Mapping (SLAM) problem.
-A large variety of solutions is available to the SLAM problem, these approaches can be classified as either filtering or smoothing [\textbf{SOURCE}].
+A large variety of solutions is available to the SLAM problem.
 
-Filtering approaches model the problem as a so called ``on-line" state estimation problem.
+Some approaches model the problem as a so called ``on-line" state estimation problem.
 The state of such a system consists of the current robot position and the map.
-New measurements augment and refine the estimate as they become available.
+New measurements incrementally refine the estimate as they become available.
 Due to their incremental nature these filtering approaches are commonly referred to as on-line SLAM methods.
-The other type of approach estimates the full trajectory from the full set of measurements.
-These approaches address the so called full SLAM problem and they typically rely on least-square error minimization techniques [\textbf{SOURCE}].
+
+Others estimate the full trajectory from the full set of measurements.
+These approaches address the so called full SLAM problem and they typically rely on least-square error minimization techniques \cite{Leastsquares}.
 
@@
 
-SLAM is a problem because, generally, localization requires a map and mapping requires the loacation of a device or a robot that wants to create a map.
+SLAM is a problem because, generally, localization requires a map and mapping requires the location of the device or robot that wants to create a map.
 There are quite a few approaches that successfully solve the SLAM problem.
-Mostly, filtering techniques like Kalman filters Particle filters and Information filters are used to estimate the state of the system.
-FastSlam \citep{Montemerlo02} is an example of a SLAM implementation using a particle filter to approximate the robots trajecotry and positions of landmarks on the map.
+In general there are two interpretations of SLAM.
+On-line SLAM methods estimate the current robot position along with the map.
+The core of a system solving the on-line SLAM problem is a filtering technique.
+FastSlam \citep{Montemerlo02} is an example of an on-line SLAM implementation using a Particle filter to approximate the robots position and the landmark positions on the map.
+Other methods have the goal to find the trajectory of the robot and the map.
+They solve the full SLAM problem.
+An example of such a an approach is graph-based SLAM of graphSLAM.
 
-Conversely, the smoothing approaches keep using all measurements to estimate the entire trajectory the robot has traveled.
-\textbf{TODO: CONNECTION}
-\subsubsection{GraphSLAM}
+\subsubsection{Graph SLAM}
 An intuitive way to formulate the SLAM problem is to view it as a pose graph optimization problem.
-A graph will be constructed where the nodes represent robot poses or landmarks and where edges correspond to measurements that constraint the connecting nodes.
-Note that constraints can be contradictory because of the noise on the measurements.
-Once a graph is constructed the algorithm will try to find a configuration of the nodes that is maximally consistent with the measurements.
+Each position along the trajectory of the robot is represented by one node.
+Edges are labelled with relative spatial information between the connecting nodes.
+This graph is build incrementally.
+While the robot is in motion, new nodes are added to the graph.
+While the graph building is a straight forward procedure the essence of the graphSLAM method is the optimization step.
+In this step the algorithm looks for a configuration of the trajectory which is maximally consistent with a system of constraints obtained from the edges.
+This problem is commonly solved using least-square error minimization techniques \citep{Grisetti}.
+
+The next section presents the algorithms implemented for this research project.
 
@@
 
-The implementation used for this paper is written on a Linux operating system.
-Programming has been done using ROS and the object oriented programming language C++.
-The installation process for ROS, as done in this paper, can be reviewed on Maastricht's swarmlab homepage \cite{swarmlab}.
+The algorithms explained in this sections are implemented in the object-oriented programming language C++.
+The code is specifically designed for the ROS environment (see Subsection \ref{subsec:ros}).
+The installation process for ROS, as performed for this paper, can be reviewed on \cite{swarmlab}.\\
+
+(\textbf{TODO: Rewrite + Eigen})
+
 For the coding part there exist two main packages, one of which includes the exploration and the other one the mapping part of the SLAM problem.
@@
 \end{enumerate}
-To understand the need of these four structures it is necessary to know that WFD is using a breadth first search (BFS), starting at the current robot position that has to be passed into the algorithm.
-WFD also requires a current map representation, which is acquired by the mapping package in this program.
-Upon launching the algorithm it first searches for unoccupied space around the current robot location, which is then saved into the Map-Open-List. 
+To understand the need of these four structures it is necessary to know that WFD is using a breadth first search (BFS), starting at the current robot position.
+Moreover frontier detection requires the robot to have an internal representation of the current environment.
+In this implementation an Occupancy grid (Subsection \ref{sec:mapping}) is used.
+Upon launching the algorithm it first searches for unoccupied space around the current robot location. This grid cells are then saved in the Map-Open-List. 
 Frontiers are found by making use of the occupancy grid's structure.
-A criterion for a point to be a frontier is that it has to have unknown space and known unoccupied space around it.
-If a frontier point is found a new BFS is performed, to extract all adjacent frontier points and connect them to one consistent frontier into the Frontier-Open-List.
-The benefit of using WFD is that by maintaining these lists, rescans only need to incorporate points that have not been scanned yet, as the algorithm saves detected open space and frontiers into its adjacent close lists, if a point is rescanned.
-For up following scans these points are no longer considered and by that the algorithm does not need to be rerun all over again.
+The criterion for a map-point to be a frontier is the existence of both, unknown space and known unoccupied space, in adjacent map-points.
+When a frontier point is found a new BFS is performed.
+Starting at the current point it extracts all adjacent frontier points and connects them to one frontier.
+It is saved in the Frontier-Open-List.
+The benefit of using WFD is that by maintaining these lists, rescans only need to incorporate points that have not been iterated yet.
+In fact, the algorithm keeps track of detected open space and frontiers.
+Following iterations make use of this information because they do not need to consider previously known regions any more.
 
- Detecting frontiers and navigating towards them directly is not possible due to ROS's navigation stack.
-Instead of sending the robot the goal to a frontier point, an algorithm has to be performed to find a point next to the frontier cell with as much free space around it as possible.
-For detection purposes an evaluation function for the frontiers has been written, which incorporates three features that are used to send a goal to the robot's navigation stack.
+Exploration does not only consist of finding frontiers.
+Based on the frontiers a target location for the robot needs to be selected.
+Navigating to frontier points directly is the simplest frontier-based exploration strategy.
+For this project, however, a more sophisticated implementation is chosen.
+Instead of sending the robot directly to a frontier point, a small search is performed to find suitable destinations.
+Generally, a small region around the frontier cell is searched.
+The goal is to have a cell with as much free space around it as possible.
+For this purposes an evaluation function has been created.
+It incorporates three features that are used to send a goal to the robot's navigation stack.
 The three features are :
 \begin{enumerate}
-\item{distance of frontier to the robot location}
-\item{number of obstacles around the designated} \item{distance to the previous goal}
+\item{Distance of frontier to the robot location}
+\item{Number of obstacles around the designated} \item{Distance to the previous goal}
 \end{enumerate}
-  The features are added up and the sum is used for the final goal for the navigation stack.
-By applying these three features the robot can always access the next point and is by that able to drive throughout the whole map and perform the algorithm and exploration perfectly.
+ The first two features have equal weights, whereas the third gets a weight of 5.
+This configuration turned out to create the best results (see Section \ref{sec:exp}) Their weighted sum is used to find the final goal to be sent to the navigation stack.
+This evaluation function enables the algorithm to select a reachable destination for the robot in almost every case.
+Additionally the information gain is still next to optimal since the selected goal is preferred to be close to frontier points.
+At last the third feature makes sure that the robots path is smooth.
+Selecting a goal near the previous one makes the robot avoid to many loops in the emerging graph.
 \subsection{GraphSLAM}
+\label{sec:implSLAM}
 Two different approaches of graph-based SLAM are implemented for this project. They differ in complexity and thereby applicability to the problem. Both algorithms create an occupancy grid map. 
-
+\begin{figure}[h]
+	\centering
+		\includegraphics[width=0.50\textwidth]{figures/graph.png}
+	\caption{An example of a simple graph (based on \citep{Grisetti}).}
+	\label{fig:turtlebotexample_graph}
+\end{figure}
 First, an algorithm as suggested by \citep{Grisetti} is implemented.
@@
 Each occupancy grid cell containing an obstacle is considered a map feature.
+This is because the map features are all that is added to the occupancy grid in the end.
 Edges are labelled with spatial constraints of the two nodes they are connecting.
 The output of this implementation is a vector containing the positions of the land marks as well as the trajectory of the robot.
-Here the constraints labelling the edges ar created using Jacobians similar to the previous approach.
+Here the constraints labelling the edges are created using Jacobians similar to the previous approach.
 The major difference lies in the node creation.
-Here nodes are added for landmarks while the previous implementation considered the robots' trajectory only.
-Since no obstacle in the final occupancy grid must be missed all obstacle points are treated as landmarks for this implementation.
-\subsubsection{Scan matching}
+Here nodes are added for landmarks while the previous implementation considered the robot's trajectory only.
+An additional difference lies in the fact that this implementation uses features to characterize the map.
+The algorithm needs to align each new observed feature with the set of features already seen.
+This matching step has to be performed and repeated along with the complete optimization procedure until convergence.
 
-The scan matcher, used by this program, is necessary to acquire an optimized movement vector between two robot positions, which is needed by the GraphSLAM algorithm.
-The programming of the scan matcher used requires two steps.
-For the first step the two scans are matched together by making use of raw odometry data, retrieved by the robot's motion sensor.
-The improvement done by the scan matcher is then to run an iteration of the iterative closest point algorithm (ICP), to incorporate its perception with the movement.
+\subsection{Scan matching}
+\label{sec:scan}
+The scan matcher, used by this program, is necessary to acquire an optimized movement vector between two robot positions.
+This is needed by the GraphSLAM algorithm (see Subsection \ref{sec:implSLAM}).
+The scan matcher realized for this project can be divided into two steps.
+First, the two scans are matched making use of raw odometry data, retrieved by the robot's motion sensor.
+This is improved upon by running an iteration of the iterative closest point algorithm (ICP).
+It tries to align the two scans minimizing the total distance of each point in the first scan to its nearest neighbour in the second.
 
-The general procedure of the algorithm is to first find corresponding neighbors.
-These are found by using the nearest neighbor search, through which adjacent points are found by calculating two points from different sets according to their euclidean distance to one another.
-Points in this implementation can have multiple correspondences to neighbors.
-To avoid incorporating outlying points, a fixed distance has been set.
-Outliers would otherwise impact further calculations negatively.
-The next step in the process is to calculate center points for both clouds, which are needed to normalize them in respect ot each other.
-After performing the calculation the center points for each cloud are subtracted from every point belonging to that cloud.
-The calculated points are then added to two separate matrices, both representing one point cloud as a $2xn$ Matrix, where n is the number of neighbor pairs.
-The first cloud's matrix is then multiplied by the transposition of the second cloud matrix, so that a new matrix is created for which we calculate the Singular Value Decomposition (SVD) values to retrieve a rotational matrix, responsible for rotating the second point cloud to the first one.
-The next step is to translate the second point cloud to the first one with the use of a translation vector.
-Both rotation matrix and translation vector are then passed to a matrix containing both of them as well as a third line with the values $0 0 1$, for symmetry purposes.
-That matrix is then used to adjust the second point cloud to the first one.
-It is also used to correct the robot position.
+The general procedure of the algorithm is to first find the nearest neighbouring point in the second scan for all point in the first. To achieve this nearest neighbour search is used.
+In this implementation multiple points can have the same corresponding neighbour-point.
+If there is no point from the other scan within a specified radius of one point, it is excluded.
+These outliers would influence further calculations negatively.
+The next step in the process is to calculate the centroid of both clouds.
+These are subtracted from every point in the concerning cloud.
+In that way the clouds are centred at the origin.
+The calculated points are then added to two separate $2xn$ matrices, where $n$ is the number of neighbour pairs.
+Next the rotation matrix which transforms one cloud to the other is needed.
+The Singular Value Decomposition (SVD) of the product of the two previously mentioned matrices returns three new matrices $U$, $\Sigma$ and $V$.
+The required rotation matrix R can be obtained as described in formula \ref{eq:RVUT}.
+\begin{equation}
+\label{eq:RVUT}
+ R = V * U^T
+\end{equation}
+The translation vector $t$ is the difference between the two centroids.
+$R$ and $t$ together form a rigid transformation.
+This transformation matches the two scans.
+Moreover it is used to correct the robot's odometry data.
 
@@
 
-For Mapping purposes an Occupancy grid is created on which information about obstacles, free/- and unknown space is saved. Using ROS's visualization program, RVIZ, the map can be visualized entirely.
+The map is internally represented as an occupancy grid.
+It consists of cells which hold information about obstacles, free/- and unknown space.
+Using ROS' visualization program, RViz, the map can be visualized on the screen.
 
-The occupancy grid is a structure that can be represented by a single dimensional array.
-Its size is responsible for the size of the grid and is often problem specific, especially for the purpose of Slam.
-Making the grid too large can lead to performance problems and by that the grid's size gets to be a tunable parameter that has to be chosen carefully for each problem, although making the grid large prevents the robot from going ``out of space''.
- This program converts the array into a two dimensional representation, so that navigating to either side gets more intuitive.
-Each value of the array is an integer and represents the current status of the cell.
-The value ranges for ROS's occupancy grid lie between $-1$ and $100$ \cite{occupancy}.
-Each value has its own meaning :
-
+The occupancy grid is a structure that is represented by a one-dimensional array.
+The size of an occupancy grid has to be tuned problem specific.
+A too large grid can lead to performance problems.
+If the grid is too small, however, the map will be inaccurate.
+Fitting a map on a smaller grid results in a low resolution.
+Internally the program represents the grid as a two dimensional array.
+This makes navigating to either side more intuitive.
+Each grid cell contains an integer value.
+Different values represent the states a cell can be in.
+The value-range for ROS' occupancy grid lies between $-1$ and $100$ \cite{occupancy}.
+Each value has a specific meaning:
 \begin{description}
-\item{$-1$} : unknown space, not yet covered by robot's perception sensors
-\item{$ 0$} : known free space, robot can drive on it safely
-\item{$1-100$} : probability of being an obstacle, $1$ being a low probability and $100$ being a definite obstacle
+\item{$-1$} : Unknown space, not yet covered by robot's perception sensors
+\item{$0$} : Known free space, robot can drive on it safely
+\item{$1-100$} : Probability of being an obstacle, $1$ being a low probability and $100$ being a definite obstacle.
 \end{description}
-
 \begin{figure}[htbp]
 	\centering
-		\includegraphics[width=0.50\textwidth,height = 44mm]{figures/Occup.png}
-	\caption{Occupancy Grid visualized}
+		\includegraphics[width=0.45\textwidth]{figures/Occup.png}
+	\caption{A visualized occupancy grid.}
 	\label{fig:Occupancy}
 \end{figure}
+Figure \ref{fig:Occupancy} shows an example occupancy grid.
+The robot is standing in the middle.
+It senses the state of all cells within a distance of one.
+The border cells are marked with a $-1$, which means that they are still unexplored.
+The remaining cells are explored being either an obstacle or free space.
+The occupancy grid is created using laser scan data.
+Each laser scan contains information about whether or not an obstacle has been detected at a certain bearing.
+Maximum range readings occur when the robot senses only free space in a certain direction.
+All laser scans are transformed into point clouds.
+The absolute locations of the points are dependent on the position at which the laser scan was recorded.
+In the occupancy grid detected obstacle points are indicated by value 100.
+Free space is being incorporated into the map by filling regions between detected points and the concerning robot position. This is achieved with the help of a linear function.
 
-Fig.\ref{fig:Occupancy} shows these three occurrences visualized, with a robot in the middle performing a sample scan of its environment.
-The occupancy grid is using laser scan data for this implementation.
-Each laser scan has information about whether an obstacle has been detected, or whether the maximum range of the scan has been reached.
-The scan is then transformed into a point cloud to extract its coordinate information and by that, obstacles can be found.
-
-Free space is being incorporated into the map by filling the occupancy grid between a detected obstacle, or detected maximum range point and the current robot position with $0$ values.
-This is achieved with the help of a linear function.
-If a point is not the scanner's maximum range away, the point is saved as an obstacle with value $100$.
-
-As time progresses and the robot explores more of its environment the map is constantly updated.
-This is programmed by saving laser scans for every pose of the robot, which, when traveling for a distance of $0.5$ meters is added and by that the occupancy grid is updated.
-Upon a loop closure the occupancy grid is optimized with its information about obstacles and free spaces.
+As time progresses, the robot explores more of its environment.
+As explained above, the locations of the observed obstacles need to be extracted from the laser-scan data.
+The program keeps track of the scan-data for every robot position saved in the trajectory \(x = (x_1, ..., x_n)^T \).
+Since the laser-scan data is relative to the position of recording, new information is added to the map once the robot's trajectory is adjusted by the graphSLAM module.
+When a new scan is made, it is added to the occupancy grid after the trajectory has been updated.
+When a loop closure is detected, the complete occupancy grid is updated.
+All previous scan points are refined relative to their corresponding recording positions and the grid is regenerated.
 
@@
 \label{sec:exp}
-This section demonstrates the results of the Turtlebot SLAM implementation as described in section \ref{sec:impl}.
+This section presents the results of the Turtlebot SLAM implementation as described in Section \ref{sec:impl}.
 Several outputs of the visualization tool RViz are presented and discussed.
-Comparisons are drawn between different variations of the implementation and the performance of the program is evaluated.
+Comparisons are drawn between different variations of the implementation.
+Finally, the performance of the program is evaluated.
 
@@
 When running the program via the launch file, RViz and Stage are automatically started.
-Stage shows the robot in its actual position on an accurate test map (see figure \ref{fig:stage_and_rviz} left).
+Stage shows the robot in its actual position on an accurate test map (see Figure \ref{fig:stage_and_rviz} left).
 RViz shows the current belief of the robot about its environment, position and trajectory (see figure \ref{fig:stage_and_rviz} right).
@@
 \begin{description}
-\item[green dots: ]past positions of the robot, saved as nodes of the pose graph
-\item[red arrows: ]past orientations of the robot, included in the nodes of the pose graph
-\item[blue lines: ]measurement constraints between the connected poses
-\item[yellow dots: ]frontier points detected by the wavefront frontier detection algorithm
-\item[orange dot: ] next navigation goal
-\item[black area: ] detected obstacles
-\item[light gray area: ] detected free space
-\item[dark gray area: ] unexplored space
+\item[Green dots:]Past positions of the robot, saved as nodes of the pose graph
+\item[Red arrows:]Past orientations of the robot, included in the nodes of the pose graph
+\item[Blue lines:]Measurement constraints between the connected poses
+\item[Yellow dots:]Frontier points detected by the wavefront frontier detection algorithm
+\item[Orange dot:]Next navigation goal
+\item[Black area:]Detected obstacles
+\item[Light gray area:]Detected free space
+\item[Dark gray area:]Unexplored space
 \end{description}
@@
 		\includegraphics[width=0.50\textwidth]{figures/Stage_and_rviz.jpg}
-	\caption{Robot and environment visualization with Stage (left) and RViz (right)}
+	\caption{Robot and environment visualization with Stage (Left) and RViz (Right).}
 	\label{fig:stage_and_rviz}
@@
 One of the core properties of the Graph SLAM algorithm is the correction of the pose graph and the map whenever a loop closing is detected.
-This means that when the robot has been exploring an unknown area and then reenters an explored area, the transformation between the current pose and the past re-detected pose gives essential information on the complete pose graph, correcting the current belief about the environment and position.
+This means that when the robot has been exploring an unknown area and then re-enters an explored area, the transformation between the current pose and the past, re-detected, pose gives essential information about the complete pose graph.
+It corrects the current belief about the environment and position.
 
 The effect of loop closing detection is illustrated in figure \ref{fig:loop_closing_comparison}.
-The transparent red lines indicate the accurate map, while the underlying black areas indicate the belief of the robot about the obstacles of the environment.
+The transparent red lines indicate the real map data, while the underlying black areas indicate the belief of the robot about obstacles in the environment.
 
 The left part of figure \ref{fig:loop_closing_comparison} shows the explored map when walking an eight-shaped path through the map without applying loop detection.
-It can be seen that the belief about the map differs from the actual map in many areas, especially in the lower half of the map because this half has been explored last, exposing it to most of the measurement uncertainties.
+The belief about the map differs from the actual map in many areas.
+Especially in the lower half the map is inaccurate.
+This is because this half has been explored last, exposing it to most of the measurement uncertainties.
 
-The right part of figure \ref{fig:loop_closing_comparison} depicts the exploration when walking the same path as before but applying loop closing.
-This results in a pose graph with 7 more constrains, providing more information for correcting the poses and corresponding measurements.
-One can now see that the lower half of the belief about the map is much more accurate since the area in the center has been re-detected by the algorithm.
+The right part of figure \ref{fig:loop_closing_comparison} depicts the exploration when walking the same path as before.
+This time loop closing is enabled during the process of exploration.
+The outcome is a pose graph with 7 more constrains as in the previous experiment.
+More constraints provide more information for correcting the poses and corresponding measurements.
+One can now see that the lower half of the belief about the map is much more accurate, since the area in the centre has been re-detected by the algorithm.
 
@@
 \end{figure}
+The next section explains which experiments have been conducted in order to find an optimal exploration strategy. It presents results for different goal selection methods.
+\subsection{Navigation evaluation}
+An important aspect which has a high impact on the exploration behaviour of the robot is the strategy by which the next navigation goal is determined.
+When facing the decision for the next goal, the robot needs to rely on the information it has access to.
+Frontiers can be spread all over the current map.
+The most obvious first implementation idea was to select the frontier which lies closest to the robot as the next navigation goal.
+This is desirable because the time the robot needs to reach the goal is low.
+In addition the robot gathers new information when the frontier point is reached.
 
-\subsection{Navigation evaluation}
-An important aspect which has a high impact on the exploration behavior of the robot is the strategy by which the next navigation goal is determined.
-When facing the decision for the next goal, the given information are the detected frontiers, which can be spread all over the current map.
-The most obvious first implementation idea was to select the frontier which lies closest to the robot as the next navigation goal.
-This is definitely desirable because the robot can then explore this closest point very quickly and find out about the next frontiers.
+Figure \ref{fig:navigation_comparison} shows three resulting exploration paths of the robot as well as the explored maps after 300 seconds.
+In Figure \ref{fig:navigation_comparison}(a) the simple strategy, as explained above, was applied.
+10 nodes and 14 constraints have been created in the pose graph.
+Compared to other tests this is very little information.
+Also the nodes are limited to a small area of the map.
+Applying the simple strategy results in a low degree of exploration.
+This is due to the fact that when only considering the closest frontier point there is a high risk that points are selected which are very close to obstacles.
+When the navigation algorithm tries to find a path to this point, it inflates the obstacles first.
+This causes goals which are very close to any kind of obstacle to be unreachable.
+When a point has been discovered to be unreachable, the next closest point to the robot is chosen.
+This, however, is likely to be unreachable as well.
 
-Figure \ref{fig:navigation_comparison}(a) shows the resulting exploration path of the robot and the explored map after 300 seconds when applying this simple strategy.
-10 nodes and 14 constraints have been created in the pose graph during this time which is very little information.
-Also the nodes are limited a small area of the map.
-This little degree of exploration is due to the fact that when only considering the closest frontier point there is a high risk that points are selected which are very close to obstacles.
-When the navigation algorithm tries to find a way to this point, it is inflating the obstacles first, making very close points unreachable.
-After a point has been discovered to be unreachable, the next closest point to the robot is chosen, which will likely be unreachable as well.
-
-So it can be concluded that the number of close obstacles have to be taken into account in order to select reachable points as next goals.
-In the upcoming strategy, the sum of obstacles within a radius of 50 grid cells is added to the distance to the robot for every frontier point.
-The point with the smallest sum is then selected as the next goal.
+In order to overcome the above explained problems another exploration strategy is considered.
+Obstacles have to be taken into account to be able to select reachable points as the next goal.
+In this new strategy obstacles within a radius of 50 grid cells of the frontier point are counted.
+This number is added to the distance between the robot and the concerning frontier point.
+The point with the smallest sum is selected as the next goal.
 The application of this strategy is shown in figure \ref{fig:navigation_comparison}(b).
@@
 However, one can see that the robot did not take a very exploration-efficient route through the map.
-There are a lot of unreasonable turns in areas that are already explored, while unexplored areas get left out for the time being.
+There are a lot of unreasonable turns in areas that are already explored, while unexplored areas are omitted for the time being.
 
-Another aspect is to be added to the evaluation of every frontier point.
-That is the distance between the currently evaluated frontier point and the previous navigation goal.
-When this distance is minimized additionally to the distance to the robot and the number of close obstacles, the algorithm will select the next goals in a chain-like fashion, creating a rather intuitive exploration path through the map (figure \ref{fig:navigation_comparison}(c)).
+To avoid counter-intuitive routes another aspect is to be added to the evaluation of every frontier point.
+The distance between the currently evaluated point and the previous navigation goal.
+When this distance is minimized in addition to the previous evaluation features, the algorithm will select the next goals in a chain-like fashion, creating a rather intuitive exploration path through the map (Figure \ref{fig:navigation_comparison}(c)).
 The robot was able to explore the map within only 150 seconds, creating a much smaller pose graph of 64 nodes and 66 constraints.
@@
 \subsection{Runtime analysis}
-The implemented Graph SLAM algorithm has been tested on its feasibility.
+The implemented Graph SLAM algorithm has bee updateen tested on its feasibility.
 Navigating the robot manually, the complete test map has been explored, building up a pose graph of 123 nodes and 150 constraints.
-While doing that, the runtime of the complete update function was measured in microseconds.
-The update function is executed whenever a new node in the graph is created.
-It contains the algorithms for scan matching and Graph SLAM correction, which have been measured separately.
+While doing so, the runtime of the complete update- function was measured in microseconds.
+This function is executed whenever a new node is added to the graph.
+It contains the algorithms for scan matching and Graph SLAM correction, which have also been assessed separately.
 
@@
 In contrast the scan matching is executed outstandingly quickly since it is independent of the size of the pose graph.
-Most of the time is occupied by updating the occupancy grid because every single grid cell has to redrawn for every pose.
+Most of the time is occupied by updating the occupancy grid because every single grid cell has to be redrawn for every pose.
 In order to compromise this, the update of the complete occupancy grid is only executed every third second and only if a loop closing is detected.
+
 \section{Discussion}
 \label{sec:disc}
-The Wavefront Frontier detection algorithm \citep{Keidar} succeeds in finding frontiers in a fast and accurate manner.
+The Wavefront Frontier detection algorithm \citep{Keidar} succeeds in finding frontiers in a fast and accurate manner (see Section \ref{sec:exp}).
+It turned out, however, that finding the frontiers is only part of autonomous exploration.
+Sections \ref{sec:impl} and \ref{sec:exp} explain the various possibilities there are for selecting a feasible goal for the robot.
+Adding features to the evaluation of possible destinations showed to make selection more successful.
+During the process of implementation this evaluation function evolved to a prosperous and accurate goal chooser.
 
-Implementing the algorithm suggested in \citep{Grisetti} turned out to be well suited for attacking the problem statement of this project.
-The maps created by this algorithm look accurate.
-Specifically the margin between the output map and the real map is negligible (FIGURE???).
+Implementing the algorithm suggested in \citep{Grisetti} (Subsection \ref{sec:implSLAM}) turned out to be well suited for attacking the problem statement of this project.
+The maps created by this algorithm are accurate.
+The difference between the output map and the real map is negligible (see Figure \ref{fig:loop_closing_comparison}).
 Maps are created in real time and loop closures are added when needed.
-In fact, the algorithm keeps updating the map with no noticable delay while the robot drives around steering to new goals selected by the WFD algorithm and navigated by ROS' move\_base package.
+In fact, the algorithm is able to keep updating the map with no noticeable delay (see Section \ref{sec:exp}).
+Even when the robot is controlled by the exploration module and navigated by ROS' move\_base package the program runs smoothly.
 
-The second graph-based SLAM implementation \cite{Thrun} clearly is not suited for robots equipped with laser scanners only.
-Extensive feature extraction methods are required to create occupancy grid maps with the used TurtleBot.
-Furthermore the algorithm is expected to work best with a robot which is able to detect distinct landmarks as well as a simple environment.
-A number of unique landmarks can then serve as orientation points to build the map.
-Additionally the algorithm is able to assess map and trajectory more quickly if there are fewer nodes and constraints in the graph.
+The second graph-based SLAM implementation \cite{Thrun} (Subsection \ref{sec:implSLAM}) is not suited for the problem specified in Section \ref{sec:intro}.
+It turned out that the algorithm time complexity is too high. Even with a reduced number of laser beams the map update takes an infeasible amount of time after a few steps only.
 
-\subsubsection{Future research}
+\begin{table}[h]
+\begin{center}
+\scalebox{1}{
+\begin{tabular}{|l||c|c|c|l}
+\hline
+\# steps & 2 beams & 10 beams & 30 beams\\
+\hline \hline
+2	& 1 ms & 2 ms	&  22 ms	\\
+4	& 2	ms & 11 ms	&  269 ms	\\
+6	& 3 ms & 24 ms	&  1069 ms	\\
+8	& 4 ms & 56 ms	&  2682 ms	\\
+10	& 5 ms & 126 ms	&  5250 ms	\\
+12	& 6 ms & 273 ms	&  9170 ms	\\
+14	& 7 ms & 485 ms	&  15895 ms	\\
+\hline
+\end{tabular}
+}
+\end{center}
+\caption{The time the Graph SLAM implementation proposed by \citep{Thrun} needed using the specified number of laser beams after the specified number of steps.}
+\label{tab:graphResults}
+\end{table}
 
+Note that the number of beams used for the other implementation greatly exceeds the ones in Table \ref{tab:graphResults}. (\textbf{EXACT NUMBER})
+This, again, points out that this implementation is unable to create a map for the specified setting (see Section \ref{sec:intro}).
+
+Extensive feature extraction methods are required to transform the data into a form which the proposed algorithm is able to cope with.
+But then, if the features are extracted, how can they be displayed in an occupancy grid?
+Furthermore the algorithm is expected to work best with a robot that is able to detect distinct landmarks in the map.
+A number of unique ones could then serve as orientation points to build the map.
+Applying this procedure would result in fewer nodes to be iterated.
+The algorithm would be able to assess map and trajectory more quickly.
+
+The next subsection suggests a few ideas about how this research could be continued. In doing so proposes are made for each part of this project.
+
+\subsection{Future research}
+As mentioned in Section \ref{sec:disc} the exploration strategy evolved during the process of testing.
+This does not mean that it is near perfection.
+Other features like for instance loop closings can be taken into account still.
+This leads to more constraints to be added to the graph.
+More constraints could then result in an even more accurate assessment of the robot trajectory.
+This new feature would require to be tuned well, since more constraints also lead to higher space and time complexities.
+The Graph SLAM algorithm proposed by \citep{Grisetti} has some room for improvement as well.
+Considering the suggested change in the exploration algorithm, the method would need to cope with a larger number of edges and nodes.
+The current implementation uses sparse Cholesky factorization to find the most optimal spatial configuration of the trajectory.
+This method is inherited in the ``Eigen" library.
+Although sparse Cholesky factorization solves the linear system in an efficient manner, it is likely that better algorithms emerge in the future.
+
+Most room for improvement clearly is to be found in the second Graph SLAM implementation.
+The TurtleBot used for this research project has to rely on laser range and odometry data only.
+To extract features
+A robot equipped with for instance a colour sensor is much more suited to successfully perform this Graph SLAM method.
+Coloured objects could serve as landmarks and large parts of the feature detection can be left out.
 \section{Conclusion}

### 2013-01-21 trunk/Article/Article.tex
Ready except Conclusion

@@
 The TurtleBot used for this research project has to rely on laser range and odometry data only.
-To extract features
+Extracting features from laser range data is out the scope of this project.
+Doing so could dramatically improve the performance the method.
 A robot equipped with for instance a colour sensor is much more suited to successfully perform this Graph SLAM method.
-Coloured objects could serve as landmarks and large parts of the feature detection can be left out.
+Coloured objects could serve as landmarks and large parts of the feature detection could be left out.
+Implementing a powerful feature-extractor for laser-range data is definitely a possible solution.
+Applying the method to a robot that provides different kinds of data is the preferred one, though.
+
 \section{Conclusion}

### 2013-01-21 Article/Article.tex
Article 1.0

@@
 
-The acquisition of these maps has been a major research focus in the robotics community \cite{Thrun,Grisetti,Montemerlo02,Montemerlo}.
+The acquisition of these maps has been a major research focus in the robotics community \cite{Thrun, Grisetti, Montemerlo02, Montemerlo}.
 Building or learning these maps under position uncertainty provided only sensor measurements is often referred to as the Simultaneous Localization And Mapping (SLAM) problem.
-A large variety of solutions is available to the SLAM problem, these approaches can be classified as either filtering or smoothing.
+A large variety of solutions is available to the SLAM problem.
 
-Filtering approaches model the problem as a so called ``on-line" state estimation problem.
+Some approaches model the problem as a so called ``on-line" state estimation problem.
 The state of such a system consists of the current robot position and the map.
-New measurements augment and refine the estimate as they become available.
+New measurements incrementally refine the estimate as they become available.
 Due to their incremental nature these filtering approaches are commonly referred to as on-line SLAM methods.
-The other type of approach estimates the full trajectory from the full set of measurements.
+
+Others estimate the complete trajectory from the set of measurements.
 These approaches address the so called full SLAM problem and they typically rely on least-square error minimization techniques \cite{Leastsquares}.
@@
 
-\subsubsection{GraphSLAM}
+\subsubsection{Graph SLAM}
 An intuitive way to formulate the SLAM problem is to view it as a pose graph optimization problem.
@@
 The installation process for ROS, as performed for this paper, can be reviewed on \cite{swarmlab}.\\
-(\textbf{TODO:})
+
+(\textbf{TODO: Rewrite + Eigen})
+
 For the coding part there exist two main packages, one of which includes the exploration and the other one the mapping part of the SLAM problem.
@@
 \end{enumerate}
- The features have equal weight.
-Their sum is used to find the final goal to be sent to the navigation stack.
+ The first two features have equal weights, whereas the third gets a weight of 5.
+This configuration turned out to create the best results (see Section \ref{sec:exp}) Their weighted sum is used to find the final goal to be sent to the navigation stack.
 This evaluation function enables the algorithm to select a reachable destination for the robot in almost every case.
@@
 	\caption{An example of a simple graph (based on \citep{Grisetti}).}
-	\label{fig:example_graph}
+	\label{fig:turtlebotexample_graph}
 \end{figure}
@@
 The major difference lies in the node creation.
-Here nodes are added for landmarks while the previous implementation considered the robots' trajectory only.
+Here nodes are added for landmarks while the previous implementation considered the robot's trajectory only.
 An additional difference lies in the fact that this implementation uses features to characterize the map.
@@
 It consists of cells which hold information about obstacles, free/- and unknown space.
-Using ROS's visualization program, RViz, the map can be visualized on the screen.
+Using ROS' visualization program, RViz, the map can be visualized on the screen.
 
@@
 Different values represent the states a cell can be in.
-The value-range for ROS's occupancy grid lies between $-1$ and $100$ \cite{occupancy}.
+The value-range for ROS' occupancy grid lies between $-1$ and $100$ \cite{occupancy}.
 Each value has a specific meaning:
@@
 In contrast the scan matching is executed outstandingly quickly since it is independent of the size of the pose graph.
-Most of the time is occupied by updating the occupancy grid because every single grid cell has to redrawn for every pose.
+Most of the time is occupied by updating the occupancy grid because every single grid cell has to be redrawn for every pose.
 In order to compromise this, the update of the complete occupancy grid is only executed every third second and only if a loop closing is detected.
+
 \section{Discussion}
 \label{sec:disc}
-The Wavefront Frontier detection algorithm \citep{Keidar} succeeds in finding frontiers in a fast and accurate manner.
+The Wavefront Frontier detection algorithm \citep{Keidar} succeeds in finding frontiers in a fast and accurate manner (see Section \ref{sec:exp}).
+It turned out, however, that finding the frontiers is only part of autonomous exploration.
+Sections \ref{sec:impl} and \ref{sec:exp} explain the various possibilities there are for selecting a feasible goal for the robot.
+Adding features to the evaluation of possible destinations showed to make selection more successful.
+During the process of implementation this evaluation function evolved to a prosperous and accurate goal chooser.
 
-Implementing the algorithm suggested in \citep{Grisetti} turned out to be well suited for attacking the problem statement of this project.
-The maps created by this algorithm look accurate.
-Specifically the margin between the output map and the real map is negligible (FIGURE???).
+Implementing the algorithm suggested in \citep{Grisetti} (Subsection \ref{sec:implSLAM}) turned out to be well suited for attacking the problem statement of this project.
+The maps created by this algorithm are accurate.
+The difference between the output map and the real map is negligible (see Figure \ref{fig:loop_closing_comparison}).
 Maps are created in real time and loop closures are added when needed.
-In fact, the algorithm keeps updating the map with no noticable delay while the robot drives around steering to new goals selected by the WFD algorithm and navigated by ROS' move\_base package.
+In fact, the algorithm is able to keep updating the map with no noticeable delay (see Section \ref{sec:exp}).
+Even when the robot is controlled by the exploration module and navigated by ROS' move\_base package the program runs smoothly.
 
-The second graph-based SLAM implementation \cite{Thrun} clearly is not suited for robots equipped with laser scanners only.
-Extensive feature extraction methods are required to create occupancy grid maps with the used TurtleBot.
-Furthermore the algorithm is expected to work best with a robot which is able to detect distinct landmarks as well as a simple environment.
-A number of unique landmarks can then serve as orientation points to build the map.
-Additionally the algorithm is able to assess map and trajectory more quickly if there are fewer nodes and constraints in the graph.
+The second graph-based SLAM implementation \cite{Thrun} (Subsection \ref{sec:implSLAM}) is not suited for the problem specified in Section \ref{sec:intro}.
+It turned out that the algorithm time complexity is too high. Even with a reduced number of laser beams the map update takes an infeasible amount of time after a few steps only.
 
-\subsubsection{Future research}
+\begin{table}[h]
+\begin{center}
+\scalebox{1}{
+\begin{tabular}{|l||c|c|c|l}
+\hline
+\# steps & 2 beams & 10 beams & 30 beams\\
+\hline \hline
+2	& 1 ms & 2 ms	&  22 ms	\\
+4	& 2	ms & 11 ms	&  269 ms	\\
+6	& 3 ms & 24 ms	&  1069 ms	\\
+8	& 4 ms & 56 ms	&  2682 ms	\\
+10	& 5 ms & 126 ms	&  5250 ms	\\
+12	& 6 ms & 273 ms	&  9170 ms	\\
+14	& 7 ms & 485 ms	&  15895 ms	\\
+\hline
+\end{tabular}
+}
+\end{center}
+\caption{The time the Graph SLAM implementation proposed by \citep{Thrun} needed using the specified number of laser beams after the specified number of steps.}
+\label{tab:graphResults}
+\end{table}
 
+Note that the number of beams used for the other implementation greatly exceeds the ones in Table \ref{tab:graphResults}. (\textbf{EXACT NUMBER})
+This, again, points out that this implementation is unable to create a map for the specified setting (see Section \ref{sec:intro}).
+
+Extensive feature extraction methods are required to transform the data into a form which the proposed algorithm is able to cope with.
+But then, if the features are extracted, how can they be displayed in an occupancy grid?
+Furthermore the algorithm is expected to work best with a robot that is able to detect distinct landmarks in the map.
+A number of unique ones could then serve as orientation points to build the map.
+Applying this procedure would result in fewer nodes to be iterated.
+The algorithm would be able to assess map and trajectory more quickly.
+
+The next subsection suggests a few ideas about how this research could be continued. In doing so proposes are made for each part of this project.
+
+\subsection{Future research}
+As mentioned in Section \ref{sec:disc} the exploration strategy evolved during the process of testing.
+This does not mean that it is near perfection.
+Other features like for instance loop closings can be taken into account still.
+This leads to more constraints to be added to the graph.
+More constraints could then result in an even more accurate assessment of the robot trajectory.
+This new feature would require to be tuned well, since more constraints also lead to higher space and time complexities.
+The Graph SLAM algorithm proposed by \citep{Grisetti} has some room for improvement as well.
+Considering the suggested change in the exploration algorithm, the method would need to cope with a larger number of edges and nodes.
+The current implementation uses sparse Cholesky factorization to find the most optimal spatial configuration of the trajectory.
+This method is inherited in the ``Eigen" library.
+Although sparse Cholesky factorization solves the linear system in an efficient manner, it is likely that better algorithms emerge in the future.
+
+Most room for improvement clearly is to be found in the second Graph SLAM implementation.
+The TurtleBot used for this research project has to rely on laser range and odometry data only.
+Extracting features from laser range data is out of the scope of this project.
+Doing so could dramatically improve the performance the method, though.
+A robot equipped with for instance a colour sensor is much more suited to successfully perform this Graph SLAM method.
+Coloured objects could serve as landmarks and large parts of the feature detection could be left out.
+Implementing a powerful feature-extractor for laser-range data is definitely a possible solution.
+Applying the method to a robot that provides different kinds of data is the preferred one, nevertheless.
+
+The last section draws a conclusion based on the results (Section \ref{sec:exp}) and the discussion (Section \ref{sec:disc}).
 \section{Conclusion}
 \label{sec:conc}
+In conclusion the Graph SLAM method as described by Grisetti et al.
+\citep{Grisetti} is able to solve the full SLAM problem in real time.
+Furthermore WFD succeeds in quickly detecting frontiers (Section \ref{sec:exp}).
+It turned out that the exploration strategy, besides frontier detection, is a crucial task.
+The developed evaluation function accurately assesses suitable goal locations (Section \ref{sec:disc}).
+These components in combination with the presented mapping method (Section \ref{sec:impl}) solve the problem specified in Section \ref{sec:intro}.
+
 \bibliography{references}

### 2013-01-22 trunk/turtlebot_graph_slam/src/test.cpp
Using arbitrary uncertainty of 100.

@@ -271,8 +271,8 @@ void vel_callback(const nav_msgs::Odometry& msg) {
                 c.i = i;
                 c.j = j;
                 c.z = zTransformation;
-                double uncertainty = 0.1;
-                Matrix3d covariance = Matrix3d::Identity(3, 3); // @todo: Find correct matrix
+                double uncertainty = 100;
+                Matrix3d covariance = Matrix3d::Identity(3, 3);
                 covariance(0,0) = uncertainty;
                 covariance(1,1) = uncertainty;
                 covariance(2,2) = uncertainty;

### 2013-01-22 trunk/Article/Article.tex
Some small changes in the article on the first 4 pages

@@
 \label{sec:intro}
-Nowadays mobile robots are able to assist industry and society in a diverse set of problems.
-Transportation, search, rescue and automated vacuum cleaning are examples from different fields.
+Nowadays mobile robots succeed to assist industry and society in a diverse set of problems.
+Transportation, search, rescue and, more specifically, automated vacuum cleaning are examples from different fields.
 Typically, mobile robots can move through an environment using legs or wheels.
-In addition they are able to gather information about their environment using sensors.
+In addition they are able to gather information about their surrounding using sensors.
 Some even have devices similar to the human arm or hand which allows them to interact with the environment.
-The vacuum cleaning robot presented in Figure \ref{fig:vacuum_cleaner} is an example of a robot which has wheels to move around.
+A vacuum cleaning robot depicted in Figure \ref{fig:vacuum_cleaner} is an example of a robot which has wheels to move around.
 Furthermore it is equipped with a set of sensors.
@@
 For some tasks map data is available for the robot from the beginning.
-Provided a map the robot can locate itself using the incoming sensor data.
+Provided a map the robot can locate itself using the incoming sensor data or, if available, GPS data.
 For tasks in unknown environments, however, the robot must rely on internal sensors and make its own map of the environment.
 
-The acquisition of these maps has been a major research focus in the robotics community \cite{Thrun, Grisetti, Montemerlo02, Montemerlo}.
+The acquisition of these maps has been a major research focus in the robotics community \cite{ Grisetti, Montemerlo02, Montemerlo, Thrun}.
 Building or learning these maps under position uncertainty provided only sensor measurements is often referred to as the Simultaneous Localization And Mapping (SLAM) problem.
-A large variety of solutions is available to the SLAM problem.
+A large variety of solutions is available to it.
 
@@
 The state of such a system consists of the current robot position and the map.
-New measurements incrementally refine the estimate as they become available.
+New measurements augment and refine the estimate as they become available.
 Due to their incremental nature these filtering approaches are commonly referred to as on-line SLAM methods.
 
-Others estimate the full trajectory from the full set of measurements.
+Other approaches estimate the complete trajectory from the set of measurements.
 These approaches address the so called full SLAM problem and they typically rely on least-square error minimization techniques \cite{Leastsquares}.
@@
 The edges encode sensor measurements that constraint the connected poses.
-Once such a graph is constructed the graph-based SLAM method tries to find a configuration of the nodes that is maximally consistent with the constraints contained in the edges.
+Once such a graph is constructed the graph-based SLAM, or Graph SLAM, method tries to find a configuration of the nodes that is maximally consistent with the constraints contained in the edges.\\
 
-The rest of this paper is structured as follows.
-Section \ref{sec:sota} provides the reader with the necessary background information on the topic.
-Section \ref{sec:impl} is about the implementation details of the methods used.
-Section \ref{sec:exp} explains the experiments ran to validate and benchmark the implemented methods and also evaluate the obtained results.
-Section \ref{sec:disc} discusses the acquired results of the experiments and proposes further research based on those.
-And the last section wraps up the paper with the conclusions of this research.
-
-\begin{figure}[htp]
+\begin{figure}[h]
 	\centering
 		\includegraphics[width=0.50\textwidth]{figures/vacuum_cleaner.jpg}
-	\caption{Automated vacuum cleaner}
+	\caption{A vacuum cleaner robot \citep{VacuumRobot}.}
 	\label{fig:vacuum_cleaner}
@@
 
+The remainder of this paper is structured as follows.
+Section \ref{sec:sota} provides the state of the art of the discussed fields.
+Section \ref{sec:impl} is about the implementation details of the methods used.
+Section \ref{sec:exp} explains the experiments ran to validate and benchmark the implemented methods and also evaluate the obtained results.
+Section \ref{sec:disc} discusses the acquired results of the experiments and proposes future research based on those.
+And the last section wraps up the paper with the conclusions of this research.
+
+Subsection \ref{sec:problem} defines the problem statement for this project.
 \subsection{Problem Statement}
@@
 The hardware to be used is a mobile robot called TurtleBot.
-Thus the final product should allow the robot to autonomously navigate through an unknown environment and simultaneously build a map of it using a graph-based SLAM algorithm.
-Information about SLAM methods, autonomous exploration and the TurtleBot are given in Section \ref{sec:sota}.
+Thus the final product should allow the robot to autonomously navigate through an unknown environment and simultaneously build a map of it using a Graph SLAM.
+
+Information about the programming environment, the TurtleBot, autonomous exploration and SLAM methods are given in Section \ref{sec:sota}.
 
@@
 The Robot Operating System (ROS) \cite{Quigley} is the environment used throughout the development of all product components for this project.
-As described by Quigley et al., writing software for robots is difficult since there is a huge variety of possible hardware setups.
+As described by Quigley et al., writing software for robots is difficult since there is a large variety of possible hardware setups.
 This also makes code reuse non-trivial.
@@
 Many robotic software systems are currently used by researchers and industry \cite{Kramer}.
-Each of these frameworks is designed for one specific purpose.
+Each of these frameworks is designed for a specific purpose, though.
 Generally, none of these systems is directly applicable to a problem it was not designed for in the first place.
@@
 
-\subsubsection{Simulation}
+\subsubsection{ROS' simulator: ``Stage"}
 During the development process of a system running on a robot one curcial part is testing.
@@
 
+The following subsection describes the robot used; the TurtleBot.
 \subsection{TurtleBot}
 \label{subsec:turtle}
-
-The product of this project is designed to run on a robot called TurtleBot (Figure \ref{fig:turtlebot}).
-The TurtleBot is a low-cost, customizable, personal robot kit (\textbf{PARAPHRASE?}).The base of the TurtleBot is an iRobot Create (A).
+The product of this project is designed to run on a mobile robot called TurtleBot (Figure \ref{fig:turtlebot}).
+It is a low-cost, customizable, personal robot kit.
+The base of the TurtleBot is an iRobot Create (A).
 It holds a battery pack and 150 degrees/second Single Axis Gyro.
@@
 The robot runs on an open-source SDK based on ROS which integrates all the software needed to get the TurtleBot running.
-Already included are some advanced capabilities like mapping and navigation.
+Some capabilities are included in ROS and can be used in the TurtleBot after minor configurations.
+These include mapping and navigation, for example.
 \begin{figure}[h]
@@
 \end{figure}
+
+One feature which is not included in ROS yet is exploration. The field is introduced in the following subsection.
 
@@
 Intuitively, when a robot moves to a frontier it is likely that it makes measurements in previously unexplored space.
-Assuming that the new data is immediately added to its map, the explored region grows once the robot reaches a frontier.
+Given that new data can be added to its map.
+The explored region grows once the robot reaches a frontier.
 Moreover an extended map results in new frontiers.
 By constantly moving to frontiers the robot expands its knowledge about the environment.
+
+The process of building an internal map representation is introduced in Subsection \ref{subsec:sotaSlam}.
 
@@
 They solve the full SLAM problem.
-An example of such a an approach is graph-based SLAM of graphSLAM.
+An example of such a an approach is graph-based SLAM or Graph SLAM.
 
@@
 While the robot is in motion, new nodes are added to the graph.
-While the graph building is a straight forward procedure the essence of the graphSLAM method is the optimization step.
-In this step the algorithm looks for a configuration of the trajectory which is maximally consistent with a system of constraints obtained from the edges.
+Building up the graph, as explained above, is a straight forward procedure.
+The essence of the Graph SLAM method, however, is the optimization step.
+In this step the algorithm tries to find a spatial configuration of the trajectory which is maximally consistent with a system of constraints obtained from the edges.
 This problem is commonly solved using least-square error minimization techniques \citep{Grisetti}.
@@
 
-The algorithms explained in this sections are implemented in the object-oriented programming language C++.
+The methods explained in this sections are implemented in the object-oriented programming language C++.
 The code is specifically designed for the ROS environment (see Subsection \ref{subsec:ros}).
-The installation process for ROS, as performed for this paper, can be reviewed on \cite{swarmlab}.\\
+The installation process for ROS, as performed for this paper, can be reviewed on \cite{swarmlab}.
 
-(\textbf{TODO: Rewrite + Eigen})
+In the product there exist two main packages.
+One deals with the autonomous exploration of the robot.
+The other one is responsible for map creation using Graph SLAM.
+Part of the autonomous exploration is performed by ROS' navigation stack move\_base.
+It takes care of finding a path from the current location to a specified destination.
+It is also capable of controlling the robot along this path.
+In addition to the move\_base package another external source is used by the program.
+The ``eigen" library \citep{eigen} provides the user with extensive linear algebra operations.
+Included are simple matrix operations as well as advanced techniques.
+Two techniques which are used in this project are Singular Value Decomposition and sparse Cholesky factorization.
+Both are implemented in ``eigen".
 
-For the coding part there exist two main packages, one of which includes the exploration and the other one the mapping part of the SLAM problem.
-To enable the robot with exploration capabilities, the ROS internal navigation stack for the TurtleBot has been used, making movement possible, by sending a designated goal to the robot.
+This section is structured as follows.
+Subsection \ref{sec:wfd} describes the implemented exploration strategy.
+Next, Subsection \ref{sec:implSLAM} deals with the implemented SLAM methods.
+This is followed by a subsection about scan matching.
+The section is concluded with a part about the mapping process.
 
 \subsection{Wavefront frontier detection}
-
+\label{sec:wfd}
 One part of the assignment is autonomous exploration. The robot has to drive autonomously through an unknown world. This feature is implemented using Wavefront frontier detection (WFD) \citep{Keidar}.
@@
 In this implementation an Occupancy grid (Subsection \ref{sec:mapping}) is used.
-Upon launching the algorithm it first searches for unoccupied space around the current robot location. This grid cells are then saved in the Map-Open-List. 
+Upon launching the algorithm it first searches for unoccupied space around the current robot location. These grid cells are saved in the Map-Open-List. 
 Frontiers are found by making use of the occupancy grid's structure.
@@
 It is saved in the Frontier-Open-List.
-The benefit of using WFD is that by maintaining these lists, rescans only need to incorporate points that have not been iterated yet.
-In fact, the algorithm keeps track of detected open space and frontiers.
+The benefit of using WFD is that by maintaining these lists, the algorithm only needs to incorporate points that have not been iterated yet.
+In fact, the algorithm keeps track of detected open space and frontiers using the above mentioned list.
 Following iterations make use of this information because they do not need to consider previously known regions any more.
@@
 Generally, a small region around the frontier cell is searched.
-The goal is to have a cell with as much free space around it as possible.
-For this purposes an evaluation function has been created.
+The goal is to find a cell with as much free space around it as possible.
+For this purposes an evaluation function has been implemented.
 It incorporates three features that are used to send a goal to the robot's navigation stack.
@@
 \begin{enumerate}
-\item{Distance of frontier to the robot location}
-\item{Number of obstacles around the designated} \item{Distance to the previous goal}
+\item{Distance between frontier and robot location}
+\item{Number of obstacles around the designated point} \item{Distance to the previous goal}
 \end{enumerate}
@@
 \end{figure}
-First, an algorithm as suggested by \citep{Grisetti} is implemented.
+First, an algorithm as suggested in \citep{Grisetti} is implemented.
 Here the graph consists of nodes representing positions along the robot trajectory.
@@
 This information is obtained from a combination of odometry data and laser range sensor data.
-The odometry is corrected using a laser scan matcher (see Subsection \ref{sec:scan}).
+In fact, the odometry is corrected using a laser scan matcher (see Subsection \ref{sec:scan}).
 
@@
 This step is performed using sparse Cholesky factorization.
+
 A new node is added to the graph once the robot moves 0.5 meters. The algorithm optimizes the trajectory like explained once a new node is added.\\
@@
 The second implementation is based on \cite{Thrun}.
-It creates nodes for each odometry measurement of the robot as well as for each feature in the map spotted by the laser sensor.
+It creates nodes for each odometry measurement of the robot as well as for each feature in the map.
 Each occupancy grid cell containing an obstacle is considered a map feature.
-This is because the map features are all that is added to the occupancy grid in the end.
-Edges are labelled with spatial constraints of the two nodes they are connecting.
-The output of this implementation is a vector containing the positions of the land marks as well as the trajectory of the robot.
-Here the constraints labelling the edges are created using Jacobians similar to the previous approach.
+This is because all information for the occupancy-grid map is inherited in these features.
+Edges are labeled with spatial constraints of the two nodes they are connecting.
+The output of this implementation is a vector containing the positions of the landmarks as well as the trajectory of the robot.
+Here the constraints labeling the edges are created using Jacobians similar to the previous approach.
 The major difference lies in the node creation.
@@
 The scan matcher, used by this program, is necessary to acquire an optimized movement vector between two robot positions.
-This is needed by the GraphSLAM algorithm (see Subsection \ref{sec:implSLAM}).
-The scan matcher realized for this project can be divided into two steps.
+This is needed by the Graph SLAM algorithm (see Subsection \ref{sec:implSLAM}).
+The scan-matcher realized for this project can be divided into two steps.
 First, the two scans are matched making use of raw odometry data, retrieved by the robot's motion sensor.
 This is improved upon by running an iteration of the iterative closest point algorithm (ICP).
-It tries to align the two scans minimizing the total distance of each point in the first scan to its nearest neighbour in the second.
+It tries to align the two scans minimizing the total distance of each point in the first scan to its nearest neighbor in the second.
 
-The general procedure of the algorithm is to first find the nearest neighbouring point in the second scan for all point in the first. To achieve this nearest neighbour search is used.
-In this implementation multiple points can have the same corresponding neighbour-point.
+The general procedure of the algorithm is to first find the nearest neighboring point in the second scan for all point in the first. To achieve this nearest neighbor search is used.
+In this implementation multiple points can have the same corresponding neighbor-point.
 If there is no point from the other scan within a specified radius of one point, it is excluded.
 These outliers would influence further calculations negatively.
-The next step in the process is to calculate the centroid of both clouds.
+The next step in the process is to calculate the centroids of both clouds.
 These are subtracted from every point in the concerning cloud.
-In that way the clouds are centred at the origin.
-The calculated points are then added to two separate $2xn$ matrices, where $n$ is the number of neighbour pairs.
+In that way the clouds are centered at the origin.
+The calculated points are then added to two separate $2 \times n$ matrices, where $n$ is the number of neighbour pairs.
 Next the rotation matrix which transforms one cloud to the other is needed.
 The Singular Value Decomposition (SVD) of the product of the two previously mentioned matrices returns three new matrices $U$, $\Sigma$ and $V$.
-The required rotation matrix R can be obtained as described in formula \ref{eq:RVUT}.
+The required rotation matrix R can be obtained as described in Equation \ref{eq:RVUT}.
 \begin{equation}
@@
 Each grid cell contains an integer value.
-Different values represent the states a cell can be in.
+Different values represent the different states a cell can be in.
 The value-range for ROS' occupancy grid lies between $-1$ and $100$ \cite{occupancy}.
@@
 The TurtleBot used for this research project has to rely on laser range and odometry data only.
-Extracting features from laser range data is out the scope of this project.
-Doing so could dramatically improve the performance the method.
+Extracting features from laser range data is out of the scope of this project.
+Doing so could dramatically improve the performance the method, though.
 A robot equipped with for instance a colour sensor is much more suited to successfully perform this Graph SLAM method.
@@
 Implementing a powerful feature-extractor for laser-range data is definitely a possible solution.
-Applying the method to a robot that provides different kinds of data is the preferred one, though.
+Applying the method to a robot that provides different kinds of data is the preferred one, nevertheless.
 
+The last section draws a conclusion based on the results (Section \ref{sec:exp}) and the discussion (Section \ref{sec:disc}).
 \section{Conclusion}
 \label{sec:conc}
+In conclusion the Graph SLAM method as described by Grisetti et al.
+\citep{Grisetti} is able to solve the full SLAM problem in real time.
+Furthermore WFD succeeds in quickly detecting frontiers (Section \ref{sec:exp}).
+It turned out that the exploration strategy, besides frontier detection, is a crucial task.
+The developed evaluation function accurately assesses suitable goal locations (Section \ref{sec:disc}).
+These components in combination with the presented mapping method (Section \ref{sec:impl}) solve the problem specified in Section \ref{sec:intro}.
+
 \bibliography{references}

### 2013-01-22 trunk/Article/Article.tex
Corrected the small stuff. Did not include the reference for Scan mathcing yet.

@@
 Some even have devices similar to the human arm or hand which allows them to interact with the environment.
-A vacuum cleaning robot depicted in Figure \ref{fig:vacuum_cleaner} is an example of a robot which has wheels to move around.
+The vacuum cleaning robot depicted in Figure \ref{fig:vacuum_cleaner} is an example of a robot which has wheels to move around.
 Furthermore it is equipped with a set of sensors.
@@
 Provided a map the robot can locate itself using the incoming sensor data or, if available, GPS data.
-For tasks in unknown environments, however, the robot must rely on internal sensors and make its own map of the environment.
+For tasks in unknown environments, however, the robot must make its own map of the environment.
 
@@
 Some approaches model the problem as a so called ``on-line" state estimation problem.
-The state of such a system consists of the current robot position and the map.
+The state of such a system consists of the current robot pose and the map.
 New measurements augment and refine the estimate as they become available.
@@
 The Robot Operating System (ROS) \cite{Quigley} is the environment used throughout the development of all product components for this project.
-As described by Quigley et al., writing software for robots is difficult since there is a large variety of possible hardware setups.
+As described by Quigley et al., writing software for robots is difficult since there is a large variety of possible hardware set-ups.
 This also makes code reuse non-trivial.
-An additional problem is the amount of code needed to acttually ``run" something on a robot.
+An additional problem is the amount of code needed to actually ``run" something on a robot.
 A deep stack including driver-level software, perception modules and, for example, abstract reasoning systems is required before the actual problem can be attacked.
@@
 The emphasis of this framework lies on large-scale integrative robotics research, making it useful in a wide variety of situations.
-The difference between ROS and the above mentioned frameworks, is ROS' applicability to a wide range of hardware setups.
+The difference between ROS and the above mentioned frameworks, is ROS' applicability to a wide range of hardware set-ups.
 It provides several so called stacks.
@@
 To run a stack or your own code on any robot it is necessary to configure the robot and the ROS framework in a way that they can communicate with each other.
-For most robots the configuration is straight forward thanks to the online documentation of ROS \citep{Roswiki}.
+For most robots the configuration is straight forward thanks to the on-line documentation of ROS \citep{Roswiki}.
 An example for a stack is the multi-robot simulation environment ``Stage".
@@
 \subsubsection{ROS' simulator: ``Stage"}
-During the development process of a system running on a robot one curcial part is testing.
+During the development process of a system running on a robot one crucial part is testing.
 A ROS stack called Stage enables the developers to test code on a simulated robot.
@@
 
-The following subsection describes the robot used; the TurtleBot.
 \subsection{TurtleBot}
@@
 The base of the TurtleBot is an iRobot Create (A).
-It holds a battery pack and 150 degrees/second Single Axis Gyro.
-The model shown in Figure \ref{fig:turtlebot} uses a Microsoft Kinect sensor (B), the one used in this paper uses a laser range sensor, and a laptop (C) to run all the processes.
+It holds a battery pack and a 150 degrees/second Single Axis Gyro.
+The model shown in Figure \ref{fig:turtlebot} uses a Microsoft Kinect sensor (B), the one used in this paper uses a laser range sensor and a laptop (C) to run all the processes.
 The robot can also be customized using the mounting hardware (D).
@@
 
-One feature which is not included in ROS yet is exploration. The field is introduced in the following subsection.
-
 \subsection{Exploration}
@@
 In this case navigating from one location to another is an easy problem.
-If, however, the robot does not have access to a map, navigating through an environment is problematic.
+If the robot does not have access to a map, navigating through an environment is problematic.
 The problem treated in this paper (Subsection \ref{sec:problem}) assumes an unknown environment and, hence, no map.
@@
 Ideally the robot updates its internal map of the environment constantly while moving.
-One idea to efficiently explore the world is to move to locations which maximize the robots information gain.
+One idea to efficiently explore the world is to move to locations which maximize the robot's information gain.
 This target location needs to be extracted from the limited information the robot has collected with its sensors.
@@
 
-The process of building an internal map representation is introduced in Subsection \ref{subsec:sotaSlam}.
-
 \subsection{SLAM}
@@
 
-SLAM is a problem because, generally, localization requires a map and mapping requires the location of the device or robot that wants to create a map.
+SLAM is a problem because localization requires a map and mapping requires the location of the device or robot that wants to create a map.
 There are quite a few approaches that successfully solve the SLAM problem.
 In general there are two interpretations of SLAM.
-On-line SLAM methods estimate the current robot position along with the map.
+On-line SLAM methods estimate the current robot pose along with the map.
 The core of a system solving the on-line SLAM problem is a filtering technique.
-FastSlam \citep{Montemerlo02} is an example of an on-line SLAM implementation using a Particle filter to approximate the robots position and the landmark positions on the map.
+FastSlam \citep{Montemerlo02} is an example of an on-line SLAM implementation using a Particle Filter to approximate the robots pose and the landmark positions on the map.
 Other methods have the goal to find the trajectory of the robot and the map.
@@
 An intuitive way to formulate the SLAM problem is to view it as a pose graph optimization problem.
-Each position along the trajectory of the robot is represented by one node.
-Edges are labelled with relative spatial information between the connecting nodes.
-This graph is build incrementally.
+Each pose along the trajectory of the robot is represented by one node.
+Edges are labeled with relative spatial information between the connecting nodes.
+This graph is built incrementally.
 While the robot is in motion, new nodes are added to the graph.
@@
 This problem is commonly solved using least-square error minimization techniques \citep{Grisetti}.
-
-The next section presents the algorithms implemented for this research project.
 
@@
 
+%This is the new part
 In the product there exist two main packages.
@@
 In addition to the move\_base package another external source is used by the program.
-The ``eigen" library \citep{eigen} provides the user with extensive linear algebra operations.
+The ``Eigen" library \citep{eigen} provides the user with extensive linear algebra operations.
 Included are simple matrix operations as well as advanced techniques.
 Two techniques which are used in this project are Singular Value Decomposition and sparse Cholesky factorization.
-Both are implemented in ``eigen".
+Both are implemented in ``Eigen".
 
@@
 Subsection \ref{sec:wfd} describes the implemented exploration strategy.
-Next, Subsection \ref{sec:implSLAM} deals with the implemented SLAM methods.
+Next, Subsection \ref{sec:implSLAM} deals with the SLAM method.
 This is followed by a subsection about scan matching.
 The section is concluded with a part about the mapping process.
+%End of the new part
 
@@
 \label{sec:wfd}
-One part of the assignment is autonomous exploration. The robot has to drive autonomously through an unknown world. This feature is implemented using Wavefront frontier detection (WFD) \citep{Keidar}.
+One part of the assignment is autonomous exploration. The robot has to drive autonomously through an unknown world. This feature is implemented using Wavefront frontier detection (WFD).
 
-WFD is a frontier-based exploration algorithm. The WFD algorithm uses four different data structures. These store map and frontier information during the search process. They are :
+WFD is a frontier-based exploration algorithm \citep{Keidar}. It uses four different data structures. These store map and frontier information during the search process. They are:
 \begin{enumerate}
@@
 Moreover frontier detection requires the robot to have an internal representation of the current environment.
-In this implementation an Occupancy grid (Subsection \ref{sec:mapping}) is used.
+In this implementation an occupancy grid (Subsection \ref{sec:mapping}) is used.
 Upon launching the algorithm it first searches for unoccupied space around the current robot location. These grid cells are saved in the Map-Open-List. 
@@
 \begin{enumerate}
+\item{Number of obstacles in the area}
+\item{Distance to the previous goal}
 \item{Distance between frontier and robot location}
-\item{Number of obstacles around the designated point} \item{Distance to the previous goal}
 \end{enumerate}
 The first two features have equal weights, whereas the third gets a weight of 5.
-This configuration turned out to create the best results (see Section \ref{sec:exp}) Their weighted sum is used to find the final goal to be sent to the navigation stack.
+This configuration turned out to create the best results (see Subsection \ref{sec:navEval}) Their weighted sum is used to find the final goal to be sent to the navigation stack.
 This evaluation function enables the algorithm to select a reachable destination for the robot in almost every case.
@@
 Selecting a goal near the previous one makes the robot avoid to many loops in the emerging graph.
-\subsection{GraphSLAM}
+\subsection{Graph SLAM}
 \label{sec:implSLAM}
@@
 First, an algorithm as suggested in \citep{Grisetti} is implemented.
-Here the graph consists of nodes representing positions along the robot trajectory.
+Here the graph consists of nodes representing poses along the robot trajectory.
 Odometry and sensor measurements are incorporated into the constraints contained in the edges.
-Similar to the graphSLAM procedure explained in Section \ref{subsec:sotaSlam}, an edge connects two adjacent poses of the robot.
+Similar to the Graph SLAM procedure explained in Section \ref{subsec:sotaSlam}, an edge connects two adjacent poses of the robot.
 This implementation optimizes the pose graph according to the constraints.
 This ``optimized" trajectory then determines the positions of obstacles on the map by refining the laser scan data according to the poses they were obtained at.
-Each edge thus is labelled with the spatial difference between two poses \(x_i\) and \(x_j\).
+Each edge thus is labeled with the spatial difference between two poses \(x_i\) and \(x_j\).
 Let this error be denoted by \(e_{ij}(x_i, x_j)\).
@@
 This graph-based SLAM approach calculates the most likely path as follows:
-First it finds the Jacobians \(A\) and \(B\) of every two estimated positions connected by an edge.
+First it finds the Jacobians \(A\) and \(B\) of every two estimated poses connected by an edge.
 \begin{equation}
 \label{eq:jacobi_a}
-	A_{ij} = \frac{\partial e_{ij}(x)}{\partial x_i}\uplus
+	A_{ij} = \frac{\partial e_{ij}(x)}{\partial x_i}
 \end{equation}
@@
 \end{equation}
-Note that in Equations \ref{eq:jacobi_a} and \ref{eq:jacobi_b} the x represents the current guess of the concerning robot position.
+Note that in Equations \ref{eq:jacobi_a} and \ref{eq:jacobi_b} the x represents the current guess of the concerning robot pose.
 As explained above, \(e_{ij}(x_i, x_j)\) is the error between the two poses \(x_i\) and \(x_j\).
@@
 In one special case the algorithm adds an extra edge, connecting to non-adjacent nodes, to the graph.
-This happens when the distance between the two connected robot positions is smaller than a margin.
+This happens when the distance between the two connected robot poses is smaller than a margin.
 In fact, this special case enables the algorithm to create loops in the graph and more accurately estimate the robots trajectory and the map.
-This procedure is referred to as ``loop closing".
+This procedure is referred to as ``loop closing" (see Subsection \ref{sec:loopClosing}).
 \\
@@
 \label{sec:scan}
-The scan matcher, used by this program, is necessary to acquire an optimized movement vector between two robot positions.
+The scan matcher, used by this program, is necessary to acquire an optimized movement vector between two robot poses.
 This is needed by the Graph SLAM algorithm (see Subsection \ref{sec:implSLAM}).
@@
 In that way the clouds are centered at the origin.
-The calculated points are then added to two separate $2 \times n$ matrices, where $n$ is the number of neighbour pairs.
+The calculated points are then added to two separate $2 \times n$ matrices, where $n$ is the number of neighbor pairs.
 Next the rotation matrix which transforms one cloud to the other is needed.
@@
 
-The map is internally represented as an occupancy grid.
-It consists of cells which hold information about obstacles, free/- and unknown space.
-Using ROS' visualization program, RViz, the map can be visualized on the screen.
+The map is internally represented as an occupancy grid. It consists of cells which hold information about obstacles, free/- and unknown space.
 
@@
 Each grid cell contains an integer value.
-Different values represent the different states a cell can be in.
 The value-range for ROS' occupancy grid lies between $-1$ and $100$ \cite{occupancy}.
-Each value has a specific meaning:
+The values indicate the state as follows:
 \begin{description}
@@
 		\includegraphics[width=0.45\textwidth]{figures/Occup.png}
-	\caption{A visualized occupancy grid.}
+	\caption{A visualized occupancy grid}
 	\label{fig:Occupancy}
@@
 Figure \ref{fig:Occupancy} shows an example occupancy grid.
-The robot is standing in the middle.
+The robot stands in the center of the it.
 It senses the state of all cells within a distance of one.
 The border cells are marked with a $-1$, which means that they are still unexplored.
-The remaining cells are explored being either an obstacle or free space.
+The remaining cells are explored being either an obstacle (100) or free space (0).
 The occupancy grid is created using laser scan data.
@@
 All laser scans are transformed into point clouds.
-The absolute locations of the points are dependent on the position at which the laser scan was recorded.
-In the occupancy grid detected obstacle points are indicated by value 100.
-Free space is being incorporated into the map by filling regions between detected points and the concerning robot position. This is achieved with the help of a linear function.
+The absolute locations of the points are dependent on the pose at which the laser scan was recorded.
+In the occupancy grid detected obstacle points are indicated by a value of 100.
+Free space is being incorporated into the map by filling regions between detected points and the concerning robot position.
 
@@
 As explained above, the locations of the observed obstacles need to be extracted from the laser-scan data.
-The program keeps track of the scan-data for every robot position saved in the trajectory \(x = (x_1, ..., x_n)^T \).
-Since the laser-scan data is relative to the position of recording, new information is added to the map once the robot's trajectory is adjusted by the graphSLAM module.
+The program keeps track of the scan-data for every robot pose saved in the trajectory \(x = (x_1, ..., x_n)^T \).
+Since the laser-scan data is relative to the pose of recording, new information is added to the map once the robot's trajectory is adjusted by the Graph SLAM module.
 When a new scan is made, it is added to the occupancy grid after the trajectory has been updated.
 When a loop closure is detected, the complete occupancy grid is updated.
-All previous scan points are refined relative to their corresponding recording positions and the grid is regenerated.
+All previous scan points are refined relative to their corresponding recording poses and the grid is regenerated.
 
-\section{Experiments and Results}
+\section{Experiments and results}
 \label{sec:exp}
-This section presents the results of the Turtlebot SLAM implementation as described in Section \ref{sec:impl}.
+This section presents the results of the TurtleBot SLAM implementation as described in Section \ref{sec:impl}.
 Several outputs of the visualization tool RViz are presented and discussed.
@@
 \subsection{Visualization}
-When running the program via the launch file, RViz and Stage are automatically started.
+\label{sec:visual}
+When running the program, RViz and Stage are started.
 Stage shows the robot in its actual position on an accurate test map (see Figure \ref{fig:stage_and_rviz} left).
-RViz shows the current belief of the robot about its environment, position and trajectory (see figure \ref{fig:stage_and_rviz} right).
+RViz visualizes the current belief of the robot about its environment, position and trajectory (see figure \ref{fig:stage_and_rviz} right).
 
@@
 \begin{description}
-\item[Green dots:]Past positions of the robot, saved as nodes of the pose graph
+\item[Green dots:]Past poses of the robot, saved as nodes of the pose graph
 \item[Red arrows:]Past orientations of the robot, included in the nodes of the pose graph
 \item[Blue lines:]Measurement constraints between the connected poses
-\item[Yellow dots:]Frontier points detected by the wavefront frontier detection algorithm
+\item[Yellow dots:]Frontier points detected by the Wavefront frontier detection algorithm
 \item[Orange dot:]Next navigation goal
@@
 \subsection{Loop closing effect}
+\label{sec:loopClosing}
 One of the core properties of the Graph SLAM algorithm is the correction of the pose graph and the map whenever a loop closing is detected.
 This means that when the robot has been exploring an unknown area and then re-enters an explored area, the transformation between the current pose and the past, re-detected, pose gives essential information about the complete pose graph.
-It corrects the current belief about the environment and position.
+It corrects the current belief about environment and position.
 
-The effect of loop closing detection is illustrated in figure \ref{fig:loop_closing_comparison}.
+The effect of loop closing detection is illustrated in Figure \ref{fig:loop_closing_comparison}.
 The transparent red lines indicate the real map data, while the underlying black areas indicate the belief of the robot about obstacles in the environment.
 
-The left part of figure \ref{fig:loop_closing_comparison} shows the explored map when walking an eight-shaped path through the map without applying loop detection.
+The left part of Figure \ref{fig:loop_closing_comparison} shows the explored map when walking an eight-shaped path through the map without applying loop detection.
 The belief about the map differs from the actual map in many areas.
@@
 
-The right part of figure \ref{fig:loop_closing_comparison} depicts the exploration when walking the same path as before.
+The right part of Figure \ref{fig:loop_closing_comparison} depicts the exploration when following the same path as before.
 This time loop closing is enabled during the process of exploration.
-The outcome is a pose graph with 7 more constrains as in the previous experiment.
+The outcome is a pose graph with seven additional constrains in comparison to the previous experiment.
 More constraints provide more information for correcting the poses and corresponding measurements.
-One can now see that the lower half of the belief about the map is much more accurate, since the area in the centre has been re-detected by the algorithm.
+One can now see that the lower half of the belief about the map is much more accurate, since the area in the center has been re-detected by the algorithm.
 
@@
 The next section explains which experiments have been conducted in order to find an optimal exploration strategy. It presents results for different goal selection methods.
+
 \subsection{Navigation evaluation}
-An important aspect which has a high impact on the exploration behaviour of the robot is the strategy by which the next navigation goal is determined.
+\label{sec:navEval}
+An important aspect which has a high impact on the exploration behavior of the robot is the strategy by which the next navigation goal is determined.
 When facing the decision for the next goal, the robot needs to rely on the information it has access to.
 Frontiers can be spread all over the current map.
-The most obvious first implementation idea was to select the frontier which lies closest to the robot as the next navigation goal.
+The most obvious first implementation idea is to select the frontier which lies closest to the robot as the next navigation goal.
 This is desirable because the time the robot needs to reach the goal is low.
@@
 This is due to the fact that when only considering the closest frontier point there is a high risk that points are selected which are very close to obstacles.
-When the navigation algorithm tries to find a path to this point, it inflates the obstacles first.
+When the navigation stack tries to find a path to this point, it inflates the obstacles first.
 This causes goals which are very close to any kind of obstacle to be unreachable.
@@
 The application of this strategy is shown in figure \ref{fig:navigation_comparison}(b).
-The robot managed to explore the map within 300 seconds, creating a pose graph with 104 nodes and 150 constraints.
+The robot managed to explore the map within 300 seconds, creating a pose graph containing 104 nodes and 150 constraints.
 However, one can see that the robot did not take a very exploration-efficient route through the map.
@@
 The distance between the currently evaluated point and the previous navigation goal.
-When this distance is minimized in addition to the previous evaluation features, the algorithm will select the next goals in a chain-like fashion, creating a rather intuitive exploration path through the map (Figure \ref{fig:navigation_comparison}(c)).
+When this distance is minimized in addition to the previous evaluation features, the algorithm will select the next goals in a chain-like fashion.
+This creates an intuitive exploration path through the map (Figure \ref{fig:navigation_comparison}(c)).
 The robot was able to explore the map within only 150 seconds, creating a much smaller pose graph of 64 nodes and 66 constraints.
-This evaluation feature has been given a weight factor of 5 to increase the effect on the goal selection.
+This evaluation feature has been given a weight factor of $5$ to increase the effect on the goal selection.
 
@@
 \subsection{Runtime analysis}
-The implemented Graph SLAM algorithm has bee updateen tested on its feasibility.
+\label{sec:runtime}
+The implemented Graph SLAM algorithm has been tested on its feasibility.
 Navigating the robot manually, the complete test map has been explored, building up a pose graph of 123 nodes and 150 constraints.
-While doing so, the runtime of the complete update- function was measured in microseconds.
+While doing so, the runtime of the complete update-function was measured in microseconds.
 This function is executed whenever a new node is added to the graph.
@@
 
-Even with a graph size of over 100 nodes, the runtime of the function stays in a very feasible range between 60 and 80 milliseconds.
-The robot can be moved without any delay.
+Even with a graph size of over 100 nodes, the runtime of the function stays in a feasible range between 60 and 80 milliseconds.
+The robot can be moved without delay.
 Taking only the Graph SLAM update into account, the runtime even stays below 20 milliseconds.
@@
 Most of the time is occupied by updating the occupancy grid because every single grid cell has to be redrawn for every pose.
-In order to compromise this, the update of the complete occupancy grid is only executed every third second and only if a loop closing is detected.
+In order to compromise this, the update of the complete occupancy grid is only executed every three seconds and only if a loop closing is detected.
 
@@
 Implementing the algorithm suggested in \citep{Grisetti} (Subsection \ref{sec:implSLAM}) turned out to be well suited for attacking the problem statement of this project.
-The maps created by this algorithm are accurate.
+The maps created by the algorithm are accurate (see Subsection \ref{sec:visual}).
 The difference between the output map and the real map is negligible (see Figure \ref{fig:loop_closing_comparison}).
@@
 
-The second graph-based SLAM implementation \cite{Thrun} (Subsection \ref{sec:implSLAM}) is not suited for the problem specified in Section \ref{sec:intro}.
+The second Graph SLAM implementation \cite{Thrun} (Subsection \ref{sec:implSLAM}) is not suited for the problem specified in Section \ref{sec:intro}.
 It turned out that the algorithm time complexity is too high. Even with a reduced number of laser beams the map update takes an infeasible amount of time after a few steps only.
@@
 
-Note that the number of beams used for the other implementation greatly exceeds the ones in Table \ref{tab:graphResults}. (\textbf{EXACT NUMBER})
+Note that the number of beams used for the other implementation greatly exceeds the ones in Table \ref{tab:graphResults}.
+While the working Graph SLAM method deals with roughly 390 laser beams, this one has major difficulties with not more than 30.
 This, again, points out that this implementation is unable to create a map for the specified setting (see Section \ref{sec:intro}).
@@
 Other features like for instance loop closings can be taken into account still.
-This leads to more constraints to be added to the graph.
+This would lead to more constraints to be added to the graph.
 More constraints could then result in an even more accurate assessment of the robot trajectory.
@@
 Extracting features from laser range data is out of the scope of this project.
-Doing so could dramatically improve the performance the method, though.
-A robot equipped with for instance a colour sensor is much more suited to successfully perform this Graph SLAM method.
-Coloured objects could serve as landmarks and large parts of the feature detection could be left out.
+Doing so could dramatically improve the performance of the method, though.
+A robot equipped with for instance a color sensor is much more suited to successfully perform this Graph SLAM method.
+Colored objects could serve as landmarks and large parts of the feature detection could be left out.
 Implementing a powerful feature-extractor for laser-range data is definitely a possible solution.

### 2013-01-22 Article/Article.tex
Deleted double definition of graph slam

@@
 \label{sec:intro}
-Nowadays mobile robots are able to assist industry and society in a diverse set of problems.
-Transportation, search, rescue and automated vacuum cleaning are examples from different fields.
+Nowadays mobile robots succeed to assist industry and society in a diverse set of problems.
+Transportation, search, rescue and, more specifically, automated vacuum cleaning are examples from different fields.
 Typically, mobile robots can move through an environment using legs or wheels.
-In addition they are able to gather information about their environment using sensors.
+In addition they are able to gather information about their surrounding using sensors.
 Some even have devices similar to the human arm or hand which allows them to interact with the environment.
-The vacuum cleaning robot presented in Figure \ref{fig:vacuum_cleaner} is an example of a robot which has wheels to move around.
+The vacuum cleaning robot depicted in Figure \ref{fig:vacuum_cleaner} is an example of a robot which has wheels to move around.
 Furthermore it is equipped with a set of sensors.
@@
 For some tasks map data is available for the robot from the beginning.
-Provided a map the robot can locate itself using the incoming sensor data.
-For tasks in unknown environments, however, the robot must rely on internal sensors and make its own map of the environment.
+Provided a map the robot can locate itself using the incoming sensor data or, if available, GPS data.
+For tasks in unknown environments, however, the robot must make its own map of the environment.
 
-The acquisition of these maps has been a major research focus in the robotics community \cite{Thrun, Grisetti, Montemerlo02, Montemerlo}.
+The acquisition of these maps has been a major research focus in the robotics community \cite{ Grisetti, Montemerlo02, Montemerlo, Thrun}.
 Building or learning these maps under position uncertainty provided only sensor measurements is often referred to as the Simultaneous Localization And Mapping (SLAM) problem.
-A large variety of solutions is available to the SLAM problem.
-
-Some approaches model the problem as a so called ``on-line" state estimation problem.
-The state of such a system consists of the current robot position and the map.
-New measurements incrementally refine the estimate as they become available.
-Due to their incremental nature these filtering approaches are commonly referred to as on-line SLAM methods.
-
-Others estimate the complete trajectory from the set of measurements.
-These approaches address the so called full SLAM problem and they typically rely on least-square error minimization techniques \cite{Leastsquares}.
-
-An intuitive way to formulate the SLAM problem is to consider it as a pose graph optimization problem.
-The nodes of the graph represent robot poses or landmarks.
-The edges encode sensor measurements that constraint the connected poses.
-Once such a graph is constructed the graph-based SLAM method tries to find a configuration of the nodes that is maximally consistent with the constraints contained in the edges.
-
-The rest of this paper is structured as follows.
-Section \ref{sec:sota} provides the reader with the necessary background information on the topic.
-Section \ref{sec:impl} is about the implementation details of the methods used.
-Section \ref{sec:exp} explains the experiments ran to validate and benchmark the implemented methods and also evaluate the obtained results.
-Section \ref{sec:disc} discusses the acquired results of the experiments and proposes further research based on those.
-And the last section wraps up the paper with the conclusions of this research.
-
-\begin{figure}[htp]
+A large variety of solutions is available to it.
+\begin{figure}[h]
 	\centering
 		\includegraphics[width=0.50\textwidth]{figures/vacuum_cleaner.jpg}
-	\caption{Automated vacuum cleaner}
+	\caption{A vacuum cleaner robot \citep{VacuumRobot}.}
 	\label{fig:vacuum_cleaner}
@@
 The Robot Operating System (ROS) \cite{Quigley} is the environment used throughout the development of all product components for this project.
-As described by Quigley et al., writing software for robots is difficult since there is a huge variety of possible hardware setups.
+As described by Quigley et al., writing software for robots is difficult since there is a large variety of possible hardware set-ups.
 This also makes code reuse non-trivial.
-An additional problem is the amount of code needed to acttually ``run" something on a robot.
+An additional problem is the amount of code needed to actually ``run" something on a robot.
 A deep stack including driver-level software, perception modules and, for example, abstract reasoning systems is required before the actual problem can be attacked.
@@
 
-\subsubsection{Simulation}
-During the development process of a system running on a robot one curcial part is testing.
+\subsubsection{ROS' simulator: ``Stage"}
+During the development process of a system running on a robot one crucial part is testing.
 A ROS stack called Stage enables the developers to test code on a simulated robot.
@@
 \label{subsec:turtle}
-
-The product of this project is designed to run on a robot called TurtleBot (Figure \ref{fig:turtlebot}).
-The TurtleBot is a low-cost, customizable, personal robot kit (\textbf{PARAPHRASE?}).The base of the TurtleBot is an iRobot Create (A).
-It holds a battery pack and 150 degrees/second Single Axis Gyro.
-The model shown in Figure \ref{fig:turtlebot} uses a Microsoft Kinect sensor (B), the one used in this paper uses a laser range sensor, and a laptop (C) to run all the processes.
+The product of this project is designed to run on a mobile robot called TurtleBot (Figure \ref{fig:turtlebot}).
+It is a low-cost, customizable, personal robot kit.
+The base of the TurtleBot is an iRobot Create (A).
+It holds a battery pack and a 150 degrees/second Single Axis Gyro.
+The model shown in Figure \ref{fig:turtlebot} uses a Microsoft Kinect sensor (B), the one used in this paper uses a laser range sensor and a laptop (C) to run all the processes.
 The robot can also be customized using the mounting hardware (D).
 The robot runs on an open-source SDK based on ROS which integrates all the software needed to get the TurtleBot running.
-Already included are some advanced capabilities like mapping and navigation.
+Some capabilities are included in ROS and can be used in the TurtleBot after minor configurations.
+These include mapping and navigation, for example.
 \begin{figure}[h]
@@
 Intuitively, when a robot moves to a frontier it is likely that it makes measurements in previously unexplored space.
-Assuming that the new data is immediately added to its map, the explored region grows once the robot reaches a frontier.
+Given that new data can be added to its map.
+The explored region grows once the robot reaches a frontier.
 Moreover an extended map results in new frontiers.
@@
 
-SLAM is a problem because, generally, localization requires a map and mapping requires the location of the device or robot that wants to create a map.
+Other approaches estimate the complete trajectory from the set of measurements.
+These approaches address the so called full SLAM problem and they typically rely on least-square error minimization techniques \cite{Leastsquares}.
+
+An intuitive way to formulate the SLAM problem is to consider it as a pose graph optimization problem.
+The nodes of the graph represent robot poses or landmarks.
+The edges encode sensor measurements that constraint the connected poses.
+Once such a graph is constructed the graph-based SLAM, or Graph SLAM, method tries to find a configuration of the nodes that is maximally consistent with the constraints contained in the edges.\\
+
+SLAM is a problem because localization requires a map and mapping requires the location of the device or robot that wants to create a map.
 There are quite a few approaches that successfully solve the SLAM problem.
-In general there are two interpretations of SLAM.
-On-line SLAM methods estimate the current robot position along with the map.
-The core of a system solving the on-line SLAM problem is a filtering technique.
-FastSlam \citep{Montemerlo02} is an example of an on-line SLAM implementation using a Particle filter to approximate the robots position and the landmark positions on the map.
+In general there are two interpretations it.
+The first estimates the current robot pose along with the map.
+The core of such a system is a filtering technique.
+New measurements augment and refine the estimate as they become available.
+Due to their incremental nature these filtering approaches are commonly referred to as on-line SLAM methods.
+FastSlam \citep{Montemerlo02} is an example of an on-line SLAM implementation.
+It uses a Particle Filter to approximate the robots pose and the landmark positions on the map.
 Other methods have the goal to find the trajectory of the robot and the map.
-They solve the full SLAM problem.
-An example of such a an approach is graph-based SLAM of graphSLAM.
+These solve the full SLAM problem and typically rely on least-square error minimization techniques \cite{Leastsquares} to find adequate estimates.
+An example of such a an approach is graph-based SLAM or Graph SLAM.
 
@@
 An intuitive way to formulate the SLAM problem is to view it as a pose graph optimization problem.
-Each position along the trajectory of the robot is represented by one node.
-Edges are labelled with relative spatial information between the connecting nodes.
-This graph is build incrementally.
+Each pose along the trajectory of the robot is represented by one node.
+Edges are labeled with relative spatial information between the connecting nodes.
+This graph is built incrementally.
 While the robot is in motion, new nodes are added to the graph.
-While the graph building is a straight forward procedure the essence of the graphSLAM method is the optimization step.
-In this step the algorithm looks for a configuration of the trajectory which is maximally consistent with a system of constraints obtained from the edges.
+Building up the graph, as explained above, is a straight forward procedure.
+The essence of the Graph SLAM method, however, is the optimization step.
+In this step the algorithm tries to find a spatial configuration of the trajectory which is maximally consistent with a system of constraints obtained from the edges.
 This problem is commonly solved using least-square error minimization techniques \citep{Grisetti}.
-
-The next section presents the algorithms implemented for this research project.
 
@@
 
-The algorithms explained in this sections are implemented in the object-oriented programming language C++.
+The methods explained in this sections are implemented in the object-oriented programming language C++.
 The code is specifically designed for the ROS environment (see Subsection \ref{subsec:ros}).
-The installation process for ROS, as performed for this paper, can be reviewed on \cite{swarmlab}.\\
+The installation process for ROS, as performed for this paper, can be reviewed on \cite{swarmlab}.
 
-(\textbf{TODO: Rewrite + Eigen})
+%This is the new part
+In the product there exist two main packages.
+One deals with the autonomous exploration of the robot.
+The other one is responsible for map creation using Graph SLAM.
+Part of the autonomous exploration is performed by ROS' navigation stack move\_base.
+It takes care of finding a path from the current location to a specified destination.
+It is also capable of controlling the robot along this path.
+In addition to the move\_base package another external source is used by the program.
+The ``Eigen" library \citep{eigen} provides the user with extensive linear algebra operations.
+Included are simple matrix operations as well as advanced techniques.
+Two techniques which are used in this project are Singular Value Decomposition and sparse Cholesky factorization.
+Both are implemented in ``Eigen".
 
-For the coding part there exist two main packages, one of which includes the exploration and the other one the mapping part of the SLAM problem.
-To enable the robot with exploration capabilities, the ROS internal navigation stack for the TurtleBot has been used, making movement possible, by sending a designated goal to the robot.
+This section is structured as follows.
+Subsection \ref{sec:wfd} describes the implemented exploration strategy.
+Next, Subsection \ref{sec:implSLAM} deals with the SLAM method.
+This is followed by a subsection about scan matching.
+The section is concluded with a part about the mapping process.
+%End of the new part
 
 \subsection{Wavefront frontier detection}
+\label{sec:wfd}
+One part of the assignment is autonomous exploration. The robot has to drive autonomously through an unknown world. This feature is implemented using Wavefront frontier detection (WFD).
 
-One part of the assignment is autonomous exploration. The robot has to drive autonomously through an unknown world. This feature is implemented using Wavefront frontier detection (WFD) \citep{Keidar}.
-
-WFD is a frontier-based exploration algorithm. The WFD algorithm uses four different data structures. These store map and frontier information during the search process. They are :
+WFD is a frontier-based exploration algorithm \citep{Keidar}. It uses four different data structures. These store map and frontier information during the search process. They are:
 \begin{enumerate}
@@
 Moreover frontier detection requires the robot to have an internal representation of the current environment.
-In this implementation an Occupancy grid (Subsection \ref{sec:mapping}) is used.
-Upon launching the algorithm it first searches for unoccupied space around the current robot location. This grid cells are then saved in the Map-Open-List. 
+In this implementation an occupancy grid (Subsection \ref{sec:mapping}) is used.
+Upon launching the algorithm it first searches for unoccupied space around the current robot location. These grid cells are saved in the Map-Open-List. 
 Frontiers are found by making use of the occupancy grid's structure.
@@
 \begin{enumerate}
-\item{Distance of frontier to the robot location}
-\item{Number of obstacles around the designated} \item{Distance to the previous goal}
+\item{Number of obstacles in the area}
+\item{Distance to the previous goal}
+\item{Distance between frontier and robot location}
 \end{enumerate}
 The first two features have equal weights, whereas the third gets a weight of 5.
-This configuration turned out to create the best results (see Section \ref{sec:exp}) Their weighted sum is used to find the final goal to be sent to the navigation stack.
+This configuration turned out to create the best results (see Subsection \ref{sec:navEval}) Their weighted sum is used to find the final goal to be sent to the navigation stack.
 This evaluation function enables the algorithm to select a reachable destination for the robot in almost every case.
@@
 \end{figure}
-First, an algorithm as suggested by \citep{Grisetti} is implemented.
-Here the graph consists of nodes representing positions along the robot trajectory.
+First, an algorithm as suggested in \citep{Grisetti} is implemented.
+Here the graph consists of nodes representing poses along the robot trajectory.
 Odometry and sensor measurements are incorporated into the constraints contained in the edges.
-Similar to the graphSLAM procedure explained in Section \ref{subsec:sotaSlam}, an edge connects two adjacent poses of the robot.
+Similar to the Graph SLAM procedure explained in Section \ref{subsec:sotaSlam}, an edge connects two adjacent poses of the robot.
 This implementation optimizes the pose graph according to the constraints.
 This ``optimized" trajectory then determines the positions of obstacles on the map by refining the laser scan data according to the poses they were obtained at.
-Each edge thus is labelled with the spatial difference between two poses \(x_i\) and \(x_j\).
+Each edge thus is labeled with the spatial difference between two poses \(x_i\) and \(x_j\).
 Let this error be denoted by \(e_{ij}(x_i, x_j)\).
 This information is obtained from a combination of odometry data and laser range sensor data.
-The odometry is corrected using a laser scan matcher (see Subsection \ref{sec:scan}).
+In fact, the odometry is corrected using a laser scan matcher (see Subsection \ref{sec:scan}).
 
 This graph-based SLAM approach calculates the most likely path as follows:
-First it finds the Jacobians \(A\) and \(B\) of every two estimated positions connected by an edge.
+First it finds the Jacobians \(A\) and \(B\) of every two estimated poses connected by an edge.
 \begin{equation}
 \label{eq:jacobi_a}
-	A_{ij} = \frac{\partial e_{ij}(x)}{\partial x_i}\uplus
+	A_{ij} = \frac{\partial e_{ij}(x)}{\partial x_i}
 \end{equation}
@@
 This step is performed using sparse Cholesky factorization.
+
 A new node is added to the graph once the robot moves 0.5 meters. The algorithm optimizes the trajectory like explained once a new node is added.\\
 In one special case the algorithm adds an extra edge, connecting to non-adjacent nodes, to the graph.
-This happens when the distance between the two connected robot positions is smaller than a margin.
+This happens when the distance between the two connected robot poses is smaller than a margin.
 In fact, this special case enables the algorithm to create loops in the graph and more accurately estimate the robots trajectory and the map.
-This procedure is referred to as ``loop closing".
+This procedure is referred to as ``loop closing" (see Subsection \ref{sec:loopClosing}).
 \\
@@
 \label{sec:scan}
-The scan matcher, used by this program, is necessary to acquire an optimized movement vector between two robot positions.
-This is needed by the GraphSLAM algorithm (see Subsection \ref{sec:implSLAM}).
-The scan matcher realized for this project can be divided into two steps.
+The scan matcher, used by this program, is necessary to acquire an optimized movement vector between two robot poses.
+This is needed by the Graph SLAM algorithm (see Subsection \ref{sec:implSLAM}).
+The scan-matcher realized for this project can be divided into two steps.
 First, the two scans are matched making use of raw odometry data, retrieved by the robot's motion sensor.
 This is improved upon by running an iteration of the iterative closest point algorithm (ICP).
-It tries to align the two scans minimizing the total distance of each point in the first scan to its nearest neighbour in the second.
+It tries to align the two scans minimizing the total distance of each point in the first scan to its nearest neighbor in the second.
 
-The general procedure of the algorithm is to first find the nearest neighbouring point in the second scan for all point in the first. To achieve this nearest neighbour search is used.
-In this implementation multiple points can have the same corresponding neighbour-point.
+The general procedure of the algorithm is to first find the nearest neighboring point in the second scan for all point in the first. To achieve this nearest neighbor search is used.
+In this implementation multiple points can have the same corresponding neighbor-point.
 If there is no point from the other scan within a specified radius of one point, it is excluded.
 These outliers would influence further calculations negatively.
-The next step in the process is to calculate the centroid of both clouds.
+The next step in the process is to calculate the centroids of both clouds.
 These are subtracted from every point in the concerning cloud.
-In that way the clouds are centred at the origin.
-The calculated points are then added to two separate $2xn$ matrices, where $n$ is the number of neighbour pairs.
+In that way the clouds are centered at the origin.
+The calculated points are then added to two separate $2 \times n$ matrices, where $n$ is the number of neighbor pairs.
 Next the rotation matrix which transforms one cloud to the other is needed.
 The Singular Value Decomposition (SVD) of the product of the two previously mentioned matrices returns three new matrices $U$, $\Sigma$ and $V$.
-The required rotation matrix R can be obtained as described in formula \ref{eq:RVUT}.
+The required rotation matrix R can be obtained as described in Equation \ref{eq:RVUT}.
 \begin{equation}
@@
 Each grid cell contains an integer value.
-Different values represent the states a cell can be in.
 The value-range for ROS' occupancy grid lies between $-1$ and $100$ \cite{occupancy}.
-Each value has a specific meaning:
+The values indicate the state as follows:
 \begin{description}

### 2013-01-22 Article/Article.tex
Changed exploration. Evaluation is now explained correctly

@@
 Based on the frontiers a target location for the robot needs to be selected.
-Navigating to frontier points directly is the simplest frontier-based exploration strategy.
+Navigating to the closest frontier points is the simplest frontier-based exploration strategy.
 For this project, however, a more sophisticated implementation is chosen.
-Instead of sending the robot directly to a frontier point, a small search is performed to find suitable destinations.
-Generally, a small region around the frontier cell is searched.
-The goal is to find a cell with as much free space around it as possible.
+Instead of sending the robot to the closest frontier point, a small search is performed to find suitable destinations.
+Generally, a small region around each frontier cell is searched.
+The goal is to give each frontier point a rating on how well it is suited as a destination.
 For this purposes an evaluation function has been implemented.
-It incorporates three features that are used to send a goal to the robot's navigation stack.
-The three features are :
+It incorporates three features that are used to evaluate goals to be sent to the navigation stack.
+They are :
 \begin{enumerate}
@@
 \end{enumerate}
-The first two features have equal weights, whereas the third gets a weight of 5.
-This configuration turned out to create the best results (see Subsection \ref{sec:navEval}) Their weighted sum is used to find the final goal to be sent to the navigation stack.
+The first two features have a weight one, whereas the third is assigned a weight of five.
+This configuration turned out to create the best results (see Subsection \ref{sec:navEval}) The weighted sum of the features is used to find the goal to be sent to the navigation stack.
 This evaluation function enables the algorithm to select a reachable destination for the robot in almost every case.
-Additionally the information gain is still next to optimal since the selected goal is preferred to be close to frontier points.
-At last the third feature makes sure that the robots path is smooth.
-Selecting a goal near the previous one makes the robot avoid to many loops in the emerging graph.
+The frontier cells are sent to the move\_base in the order of the rating.
+When the navigation stack fails to steer the robot to a designated goal within a time constraint, the next best goal replaces the previous.
+The information gain is still next to optimal since the selected goal is preferred to be close to frontier points.
+At last the second feature makes sure that the robots path is smooth.
+Selecting a goal near the previous one makes the robot avoid too many loops in the emerging graph.
 \subsection{Graph SLAM}

### 2013-01-22 trunk/Article/Article.tex
Added figure for the program structure.

@@
 A large variety of solutions is available to it.
-
-Some approaches model the problem as a so called ``on-line" state estimation problem.
-The state of such a system consists of the current robot pose and the map.
-New measurements augment and refine the estimate as they become available.
-Due to their incremental nature these filtering approaches are commonly referred to as on-line SLAM methods.
-
-Other approaches estimate the complete trajectory from the set of measurements.
-These approaches address the so called full SLAM problem and they typically rely on least-square error minimization techniques \cite{Leastsquares}.
-
-An intuitive way to formulate the SLAM problem is to consider it as a pose graph optimization problem.
-The nodes of the graph represent robot poses or landmarks.
-The edges encode sensor measurements that constraint the connected poses.
-Once such a graph is constructed the graph-based SLAM, or Graph SLAM, method tries to find a configuration of the nodes that is maximally consistent with the constraints contained in the edges.\\
-
 \begin{figure}[h]
@@
 
+Other approaches estimate the complete trajectory from the set of measurements.
+These approaches address the so called full SLAM problem and they typically rely on least-square error minimization techniques \cite{Leastsquares}.
+
+An intuitive way to formulate the SLAM problem is to consider it as a pose graph optimization problem.
+The nodes of the graph represent robot poses or landmarks.
+The edges encode sensor measurements that constraint the connected poses.
+Once such a graph is constructed the graph-based SLAM, or Graph SLAM, method tries to find a configuration of the nodes that is maximally consistent with the constraints contained in the edges.\\
+
 SLAM is a problem because localization requires a map and mapping requires the location of the device or robot that wants to create a map.
 There are quite a few approaches that successfully solve the SLAM problem.
-In general there are two interpretations of SLAM.
-On-line SLAM methods estimate the current robot pose along with the map.
-The core of a system solving the on-line SLAM problem is a filtering technique.
-FastSlam \citep{Montemerlo02} is an example of an on-line SLAM implementation using a Particle Filter to approximate the robots pose and the landmark positions on the map.
+In general there are two interpretations it.
+The first estimates the current robot pose along with the map.
+The core of such a system is a filtering technique.
+New measurements augment and refine the estimate as they become available.
+Due to their incremental nature these filtering approaches are commonly referred to as on-line SLAM methods.
+FastSlam \citep{Montemerlo02} is an example of an on-line SLAM implementation.
+It uses a Particle Filter to approximate the robots pose and the landmark positions on the map.
 Other methods have the goal to find the trajectory of the robot and the map.
-They solve the full SLAM problem.
+These solve the full SLAM problem and typically rely on least-square error minimization techniques \cite{Leastsquares} to find adequate estimates.
 An example of such a an approach is graph-based SLAM or Graph SLAM.
@@
 
+\subsection{Program structure}
+\begin{figure}[htbp]
+	\centering
+		\includegraphics[width=0.45\textwidth]{figures/Program_structure.png}
+	\caption{Overview of the structure and functionalities of the relevant software components}
+	\label{fig:program_structure}
+\end{figure}
+
 \subsection{Wavefront frontier detection}

### 2013-01-22 trunk/Article/Article.tex
Small error corrected.
Occupancy grid corrected.

@@
 
-The methods explained in this sections are implemented in the object-oriented programming language C++.
+The methods explained in this section are implemented in the object-oriented programming language C++.
 The code is specifically designed for the ROS environment (see Subsection \ref{subsec:ros}).
@@
 Figure \ref{fig:Occupancy} shows an example occupancy grid.
-The robot stands in the center of the it.
-It senses the state of all cells within a distance of one.
-The border cells are marked with a $-1$, which means that they are still unexplored.
-The remaining cells are explored being either an obstacle (100) or free space (0).
-The occupancy grid is created using laser scan data.
-Each laser scan contains information about whether or not an obstacle has been detected at a certain bearing.
-Maximum range readings occur when the robot senses only free space in a certain direction.
-All laser scans are transformed into point clouds.
-The absolute locations of the points are dependent on the pose at which the laser scan was recorded.
-In the occupancy grid detected obstacle points are indicated by a value of 100.
-Free space is being incorporated into the map by filling regions between detected points and the concerning robot position.
+The robot stands in the center of it and makes one measurement consisting of several laser scans.
+Each scan returns the range it was able to reach.
+The returned range is either the maximally possible range of the laser or a lower value, indicating an obstacle (value of 100).
+The range is transformed into global coordinates with respect to the robot's location, orientation and the bearing of the laser beam.
+The resulting point is then transferred into the occupancy grid as an obstacle if the maximum range was not reached.
+The cells lying in a straight line between the transformed range and the robot are considered free space (value of 0), since the laser was able to pass through this area.
+All the cells that are not affected by any of the previous steps remain unexplored (value of -1).
+Naturally, this is the case for cells behind obstacles, out of range and directly behind the robot.
 

### 2013-01-22 Article/Article.tex

@@
 %End of the new part
+
+\subsection{Program structure}
+\begin{figure}[htbp]
+	\centering
+		\includegraphics[width=0.45\textwidth]{figures/Program_structure.png}
+	\caption{Overview of the structure and functionalities of the relevant software components}
+	\label{fig:program_structure}
+\end{figure}
 
@@
 \label{sec:implSLAM}
-Two different approaches of graph-based SLAM are implemented for this project. They differ in complexity and thereby applicability to the problem. Both algorithms create an occupancy grid map. 
+The implemented Graph SLAM algorithm is based on \citep{Grisetti}.
+The graph consists of nodes representing poses along the robot trajectory.
+Odometry and sensor measurements are incorporated into the constraints contained in the edges.
+Similar to the Graph SLAM procedure explained in Section \ref{subsec:sotaSlam}, an edge connects two adjacent poses of the robot.
 \begin{figure}[h]
@@
 \end{figure}
-First, an algorithm as suggested in \citep{Grisetti} is implemented.
-Here the graph consists of nodes representing poses along the robot trajectory.
-Odometry and sensor measurements are incorporated into the constraints contained in the edges.
-Similar to the Graph SLAM procedure explained in Section \ref{subsec:sotaSlam}, an edge connects two adjacent poses of the robot.
-This implementation optimizes the pose graph according to the constraints.
-This ``optimized" trajectory then determines the positions of obstacles on the map by refining the laser scan data according to the poses they were obtained at.
-Each edge thus is labeled with the spatial difference between two poses \(x_i\) and \(x_j\).
+When the graph is created the pose graph is optimized according to the constraints.
+This ``optimized" trajectory then determines the positions of obstacles on the map.
+The laser scan data are refined according to the poses they were obtained at.
+Each edge, thus, is labeled with the spatial difference between two poses \(x_i\) and \(x_j\).
 Let this error be denoted by \(e_{ij}(x_i, x_j)\).
 This information is obtained from a combination of odometry data and laser range sensor data.
-In fact, the odometry is corrected using a laser scan matcher (see Subsection \ref{sec:scan}).
+In fact, the raw odometry data is corrected using a laser scan matcher (see Subsection \ref{sec:scan}).
+Figure \ref{fig:turtlebotexample_graph} shows a sketch of a graph created by the algorithm.
+Two poses \(x_i\) and \(x_j\) are connected by an edge (Solid arrow).
+The ellipse at the top illustrates the label of the edge.
 
@@
 As explained above, \(e_{ij}(x_i, x_j)\) is the error between the two poses \(x_i\) and \(x_j\).
-The Jacobians \(A\) and \(B\) are then added to an information matrix \(H\) (Equation \ref{eq:H}).
+The Jacobians \(A\) and \(B\) are then added to a matrix \(H\) (Equation \ref{eq:H}).
 \begin{equation}
@@
 \end{equation}
-In this equation $\Omega_{ij}$ is the information matrix of a measurement between \(x_i\) and \(x_j\).
-\\
-Furthermore a coefficient vector \(b\) is always updated as shown in Equation \ref{eq:b}. This vector in combination with the \(H\) matrix form a set of linear constraints.
+In this equation $\Omega_{ij}$ is the information matrix of a measurement between \(x_i\) and \(x_j\). It is defined as shown in Equation \ref{eq:omega}.
+Sigma is the $3 \times 3$ covariance matrix of the measurements. In our implementation the diagonal of $\Sigma$ is filled with arbitrary values.
+\begin{equation}
+\label{eq:omega}
+\Omega = \Sigma^{-1}
+\end{equation}
+In addition to the matrix $H$ a coefficient vector \(b\) is kept up to date as shown in Equation \ref{eq:b}. This vector together with the \(H\) form a set of linear constraints.
 \begin{equation}
@@
 This procedure is referred to as ``loop closing" (see Subsection \ref{sec:loopClosing}).
-\\
-\\
-The second implementation is based on \cite{Thrun}.
-It creates nodes for each odometry measurement of the robot as well as for each feature in the map.
-Each occupancy grid cell containing an obstacle is considered a map feature.
-This is because all information for the occupancy-grid map is inherited in these features.
-Edges are labeled with spatial constraints of the two nodes they are connecting.
-The output of this implementation is a vector containing the positions of the landmarks as well as the trajectory of the robot.
-Here the constraints labeling the edges are created using Jacobians similar to the previous approach.
-The major difference lies in the node creation.
-Here nodes are added for landmarks while the previous implementation considered the robot's trajectory only.
-An additional difference lies in the fact that this implementation uses features to characterize the map.
-The algorithm needs to align each new observed feature with the set of features already seen.
-This matching step has to be performed and repeated along with the complete optimization procedure until convergence.
-
 \subsection{Scan matching}
@@
 The scan-matcher realized for this project can be divided into two steps.
-First, the two scans are matched making use of raw odometry data, retrieved by the robot's motion sensor.
+At the beginning, the two scans are matched making use of raw odometry data, retrieved by the robot's motion sensor.
 This is improved upon by running an iteration of the iterative closest point algorithm (ICP).
@@
 
-\subsection{Mapping}\label{sec:mapping}
-
+\subsection{Mapping}
+\label{sec:mapping}
 The map is internally represented as an occupancy grid. It consists of cells which hold information about obstacles, free/- and unknown space.
@@
 
-Even with a graph size of over 100 nodes, the runtime of the function stays in a feasible range between 60 and 80 milliseconds.
+Even with a graph size of more than 100 nodes, the runtime of the function stays in a feasible range between 60 and 80 milliseconds.
 The robot can be moved without delay.
@@
 \label{sec:disc}
+In this section the results of the implemented methods are evaluated and discussed. 
+Subsection [Exploration] deals with the exploration strategy. Subsection [Graph SLAM] presents the evolution of the Graph SLAM implementation. Different stages of it are evaluated.
+\subsection{Exploration strategy}
+\label{disc:expl}
 The Wavefront Frontier detection algorithm \citep{Keidar} succeeds in finding frontiers in a fast and accurate manner (see Section \ref{sec:exp}).
@@
 During the process of implementation this evaluation function evolved to a prosperous and accurate goal chooser.
+\\
+\subsection{Graph SLAM}
+\label{disc:grSlam}
+Developing a feasible Graph SLAM algorithm showed to be a challenging task. In earlier stages of the project another Graph SLAM algorithm was implemented.
+It is based on \cite{Thrun}. 
+Nodes are created for each odometry measurement as well as for every map feature detected by the robot.
+All occupancy grid cells containing an obstacle are considered map features.
+This is necessary because all information for the occupancy-grid map is inherited in these features.
+The output of this implementation is a vector containing the positions of the landmarks as well as the trajectory of the robot.
+Here the constraints labeling the edges are created using Jacobians similar to the final approach (see Subsection \ref{sec:implSLAM}).
+The major difference lies in the node creation.
+Here, additional nodes are added for landmarks while the previous implementation considered the robot's trajectory only.
+An additional difference lies in the fact that this implementation uses features to characterize the map.
+The algorithm needs to compare every feature from the new scan has to the ones detected earlier.
+This ``correspondence test" has to be performed and repeated along with the complete optimization procedure until convergence.
 
-Implementing the algorithm suggested in \citep{Grisetti} (Subsection \ref{sec:implSLAM}) turned out to be well suited for attacking the problem statement of this project.
-The maps created by the algorithm are accurate (see Subsection \ref{sec:visual}).
-The difference between the output map and the real map is negligible (see Figure \ref{fig:loop_closing_comparison}).
-Maps are created in real time and loop closures are added when needed.
-In fact, the algorithm is able to keep updating the map with no noticeable delay (see Section \ref{sec:exp}).
-Even when the robot is controlled by the exploration module and navigated by ROS' move\_base package the program runs smoothly.
-
-The second Graph SLAM implementation \cite{Thrun} (Subsection \ref{sec:implSLAM}) is not suited for the problem specified in Section \ref{sec:intro}.
-It turned out that the algorithm time complexity is too high. Even with a reduced number of laser beams the map update takes an infeasible amount of time after a few steps only.
-
+While testing it turned out that the algorithm's time complexity is too high for real time observations.
+Even with a reduced number of laser beams the map update takes an infeasible amount of time after a few steps only.
 \begin{table}[h]
@@
 \end{table}
-
 Note that the number of beams used for the other implementation greatly exceeds the ones in Table \ref{tab:graphResults}.
 While the working Graph SLAM method deals with roughly 390 laser beams, this one has major difficulties with not more than 30.
-This, again, points out that this implementation is unable to create a map for the specified setting (see Section \ref{sec:intro}).
+This points out that this implementation is unable to create a map in reasonable time.
 
-Extensive feature extraction methods are required to transform the data into a form which the proposed algorithm is able to cope with.
-But then, if the features are extracted, how can they be displayed in an occupancy grid?
-Furthermore the algorithm is expected to work best with a robot that is able to detect distinct landmarks in the map.
-A number of unique ones could then serve as orientation points to build the map.
-Applying this procedure would result in fewer nodes to be iterated.
-The algorithm would be able to assess map and trajectory more quickly.
+Implementing the algorithm suggested in \citep{Grisetti} (Subsection \ref{sec:implSLAM}) showed to be well suited for attacking the problem statement of this project.
+The maps created by the algorithm are accurate (see Subsection \ref{sec:visual}).
+The difference between the output map and the real map is negligible (see Figure \ref{fig:loop_closing_comparison}).
+Maps are created in real time and loop closures are added when needed.
+In fact, the algorithm is able to keep updating the map with no noticeable delay (see Section \ref{sec:exp}).
+Even when the robot is controlled by the exploration module and navigated by ROS' move\_base package the program runs smoothly.
 

### 2013-01-22 trunk/Article/Article.tex
Included the abstract

@@
 \begin{abstract}
+Robots are used for a lot of tasks nowadays, to perform an assignment a robot needs a map of the environment.
+Using external referencing methods, like GPS, they can locate them self.
+Sometimes the maps are provided, making the task easier to perform.
+It becomes far more difficult when the robot has no map or external reference available.
+The task of mapping an environment can be split into two main components, namely exploration and SLAM which stands for Simultaneous Localization And Mapping.
+The exploration determines the path the robot takes to explore the unknown space of the environment.
+Exploration is based on the idea of frontiers, these are cells which separate unexplored cells from free explored cells.
+The exploration strategy then determines the next location the robot should move to.
+The SLAM part incorporates the new data as is becomes available when the robot moves to unexplored areas and takes measurements of previously unrecorded space.
+One intuitive way is to represent all the data as a graph where the nodes represent a pose of the robot and the edges connecting two nodes the measured constraints.
+The algorithm then tries to find a configuration of nodes which is maximally consistent with the constraints.
+This article will describe an implementation of this so called GraphSLAM with frontier exploration.
+All the implemented code is first tested and validated in a simulator and later run on a real robot, a TurtleBot.
+
 This article accompanies the research project "TurtleBot SLAM" from the Master "Artificial Intelligence" of Maastricht University.

### 2013-01-22 trunk/Article/Article.tex
Discussion changed

@@
 \label{subsec:turtle}
-The product of this project is designed to run on a mobile robot called TurtleBot (Figure \ref{fig:turtlebot}).
+The product of this project is designed to run on a mobile robot called TurtleBot (Figure \ref{fig:turtlebot}) \citep{turtlebot}.
 It is a low-cost, customizable, personal robot kit.
@@
 In this section the results of the implemented methods are evaluated and discussed. 
-Subsection [Exploration] deals with the exploration strategy. Subsection [Graph SLAM] presents the evolution of the Graph SLAM implementation. Different stages of it are evaluated.
+Subsection \ref{disc:expl} deals with the exploration strategy. Subsection \ref{disc:grSlam} presents the evolution of the Graph SLAM implementation. Different stages of it are evaluated.
 \subsection{Exploration strategy}
 \label{disc:expl}
-The Wavefront Frontier detection algorithm \citep{Keidar} succeeds in finding frontiers in a fast and accurate manner (see Section \ref{sec:exp}).
+The Wavefront Frontier detection algorithm \citep{Keidar} succeeds in finding frontiers in a fast and accurate manner (see Subsection \ref{sec:visual}).
 It turned out, however, that finding the frontiers is only part of autonomous exploration.
-Sections \ref{sec:impl} and \ref{sec:exp} explain the various possibilities there are for selecting a feasible goal for the robot.
+Subsections \ref{sec:wfd} and \ref{sec:visual} explain the various possibilities there are for selecting a feasible goal for the robot.
 Adding features to the evaluation of possible destinations showed to make selection more successful.
@@
 \subsection{Future research}
-As mentioned in Section \ref{sec:disc} the exploration strategy evolved during the process of testing.
-This does not mean that it is near perfection.
-Other features like for instance loop closings can be taken into account still.
+As mentioned in Section \ref{disc:expl} the exploration strategy evolved during the process of testing.
+This does not mean that it cannot be improved upon.
+More features like, for instance, loop closings can be taken into account.
 This would lead to more constraints to be added to the graph.
@@
 This new feature would require to be tuned well, since more constraints also lead to higher space and time complexities.
-The Graph SLAM algorithm proposed by \citep{Grisetti} has some room for improvement as well.
+\\
+
+The Graph SLAM algorithm proposed by \citep{Grisetti} can be improved as well.
 Considering the suggested change in the exploration algorithm, the method would need to cope with a larger number of edges and nodes.
 The current implementation uses sparse Cholesky factorization to find the most optimal spatial configuration of the trajectory.
-This method is inherited in the ``Eigen" library.
-Although sparse Cholesky factorization solves the linear system in an efficient manner, it is likely that better algorithms emerge in the future.
+This method is inherited in the ``Eigen" library \ref{sec:impl}.
+Although sparse Cholesky factorization solves the linear system in an efficient manner, it is likely that even better algorithms emerge in the future.
+Since the time complexity of the Graph SLAM method depends almost entirely on the technique used, new methods would directly improve upon the running time of the algorithm.
 
-Most room for improvement clearly is to be found in the second Graph SLAM implementation.
-The TurtleBot used for this research project has to rely on laser range and odometry data only.
-Extracting features from laser range data is out of the scope of this project.
+The deprecated Graph SLAM implementation clearly has most room for improvement.
+The TurtleBot used for this research project has to rely on laser and odometry data only.
+Extracting features from laser data is out of scope of this project.
 Doing so could dramatically improve the performance of the method, though.
@@
 Colored objects could serve as landmarks and large parts of the feature detection could be left out.
-Implementing a powerful feature-extractor for laser-range data is definitely a possible solution.
+Implementing a powerful feature-extractor for laser data is definitely a possible solution.
 Applying the method to a robot that provides different kinds of data is the preferred one, nevertheless.
@@
 \label{sec:conc}
-In conclusion the Graph SLAM method as described by Grisetti et al.
-\citep{Grisetti} is able to solve the full SLAM problem in real time.
-Furthermore WFD succeeds in quickly detecting frontiers (Section \ref{sec:exp}).
+In conclusion Graph SLAM is an algorithm with a variety of implementations.
+Incorporating the map into the graph showed to lead to an infeasible time complexity (Section \ref{sec:disc}).
+The Graph SLAM method as described by Grisetti et al.
+\citep{Grisetti} is able to solve the full SLAM problem in real time (Subsection \ref{sec:runtime}).
+The created maps are accurate (Subsection \ref{sec:navEval}).
+Furthermore WFD succeeds in quickly detecting frontiers (Subsection \ref{sec:runtime} and Figure \ref{fig:runtime}).
 It turned out that the exploration strategy, besides frontier detection, is a crucial task.
-The developed evaluation function accurately assesses suitable goal locations (Section \ref{sec:disc}).
-These components in combination with the presented mapping method (Section \ref{sec:impl}) solve the problem specified in Section \ref{sec:intro}.
+The developed evaluation function accurately assesses suitable goal locations (Subsection \ref{disc:expl}).
+It allows the robot to drive smooth, intuitive paths (Figure \ref{fig:navigation_comparison}).
+The program containing these components solves the problem specified in Subsection \ref{sec:problem}.
 

### 2013-01-22 Article/Article.tex
Corrected structure, added references.

@@
 
-The methods explained in this section are implemented in the object-oriented programming language C++.
+The software explained in this section are implemented in the object-oriented programming language C++.
 The code is specifically designed for the ROS environment (see Subsection \ref{subsec:ros}).
@@
 
-%This is the new part
-In the product there exist two main packages.
-One deals with the autonomous exploration of the robot.
-The other one is responsible for map creation using Graph SLAM.
-Part of the autonomous exploration is performed by ROS' navigation stack move\_base.
-It takes care of finding a path from the current location to a specified destination.
-It is also capable of controlling the robot along this path.
-In addition to the move\_base package another external source is used by the program.
-The ``Eigen" library \citep{eigen} provides the user with extensive linear algebra operations.
-Included are simple matrix operations as well as advanced techniques.
-Two techniques which are used in this project are Singular Value Decomposition and sparse Cholesky factorization.
-Both are implemented in ``Eigen".
+\subsection{Program structure}
+An overview of the relevant components of the final software product is given in Figure \ref{fig:program_structure}.
+The top two boxes represent the implemented packages.
+The left one performs the Graph SLAM algorithm (Subsection \ref{sec:implSLAM}), including the initialization and optimization of the pose graph and the correction by scan matching (Subsection \ref{sec:scan}).
+It also creates a map of the explored environment by maintaining an occupancy grid (Subsection \ref{sec:mapping}).
+The right package is responsible for autonomous exploration (Subsection \ref{sec:wfd}).
+This includes an algorithm for efficiently navigating through unknown space to gather as much information about the environment as possible.
 
-This section is structured as follows.
-Subsection \ref{sec:wfd} describes the implemented exploration strategy.
-Next, Subsection \ref{sec:implSLAM} deals with the SLAM method.
-This is followed by a subsection about scan matching.
-The section is concluded with a part about the mapping process.
-%End of the new part
-
-\subsection{Program structure}
 \begin{figure}[htbp]
@@
 
-\subsection{Wavefront frontier detection}
+Both packages publish data to the shared environment of ROS. This data is then used by the other packages to create an interdependent but modular program structure.
+
+The implemented packages are making use of the external library ``Eigen" \citep{eigen}.
+It provides extensive linear algebra operations.
+Included are simple matrix operations as well as advanced techniques.
+Two techniques which are used in this project are Singular Value Decomposition and sparse Cholesky factorization.
+
+\subsection{Exploration}
 \label{sec:wfd}

### 2013-01-22 trunk/Article/Article.tex
Retitled Stage.

@@
 
-\subsubsection{ROS' simulator: ``Stage"}
+\subsubsection{Stage}
 During the development process of a system running on a robot one crucial part is testing.
@@
 
-\subsection{Wavefront frontier detection}
-\label{sec:wfd}
-One part of the assignment is autonomous exploration. The robot has to drive autonomously through an unknown world. This feature is implemented using Wavefront frontier detection (WFD).
+Both packages publish data to the shared environment of ROS. This data is then used by the other packages to create an interdependent but modular program structure.
 
-WFD is a frontier-based exploration algorithm \citep{Keidar}. It uses four different data structures. These store map and frontier information during the search process. They are:
-\begin{enumerate}
-\item{Map-Open-List}
-\item{Map-Close-List}
-\item{Frontier-Open-List}
-\item{Frontier-Close-List}
-\end{enumerate}
-To understand the need of these four structures it is necessary to know that WFD is using a breadth first search (BFS), starting at the current robot position.
-Moreover frontier detection requires the robot to have an internal representation of the current environment.
-In this implementation an occupancy grid (Subsection \ref{sec:mapping}) is used.
-Upon launching the algorithm it first searches for unoccupied space around the current robot location. These grid cells are saved in the Map-Open-List. 
-Frontiers are found by making use of the occupancy grid's structure.
-The criterion for a map-point to be a frontier is the existence of both, unknown space and known unoccupied space, in adjacent map-points.
-When a frontier point is found a new BFS is performed.
-Starting at the current point it extracts all adjacent frontier points and connects them to one frontier.
-It is saved in the Frontier-Open-List.
-The benefit of using WFD is that by maintaining these lists, the algorithm only needs to incorporate points that have not been iterated yet.
-In fact, the algorithm keeps track of detected open space and frontiers using the above mentioned list.
-Following iterations make use of this information because they do not need to consider previously known regions any more.
+The implemented packages are making use of the external library ``Eigen" \citep{eigen}.
+It provides extensive linear algebra operations.
+Included are simple matrix operations as well as advanced techniques.
+Two techniques which are used in this project are Singular Value Decomposition and sparse Cholesky factorization.
 
-Exploration does not only consist of finding frontiers.
-Based on the frontiers a target location for the robot needs to be selected.
-Navigating to the closest frontier points is the simplest frontier-based exploration strategy.
-For this project, however, a more sophisticated implementation is chosen.
-Instead of sending the robot to the closest frontier point, a small search is performed to find suitable destinations.
-Generally, a small region around each frontier cell is searched.
-The goal is to give each frontier point a rating on how well it is suited as a destination.
-For this purposes an evaluation function has been implemented.
-It incorporates three features that are used to evaluate goals to be sent to the navigation stack.
-They are :
-\begin{enumerate}
-\item{Number of obstacles in the area}
-\item{Distance to the previous goal}
-\item{Distance between frontier and robot location}
-\end{enumerate}
-The first two features have a weight one, whereas the third is assigned a weight of five.
-This configuration turned out to create the best results (see Subsection \ref{sec:navEval}) The weighted sum of the features is used to find the goal to be sent to the navigation stack.
-This evaluation function enables the algorithm to select a reachable destination for the robot in almost every case.
-The frontier cells are sent to the move\_base in the order of the rating.
-When the navigation stack fails to steer the robot to a designated goal within a time constraint, the next best goal replaces the previous.
-The information gain is still next to optimal since the selected goal is preferred to be close to frontier points.
-At last the second feature makes sure that the robots path is smooth.
-Selecting a goal near the previous one makes the robot avoid too many loops in the emerging graph.
 \subsection{Graph SLAM}
@@
 All previous scan points are refined relative to their corresponding recording poses and the grid is regenerated.
+
+\subsection{Exploration}
+\label{sec:wfd}
+One part of the assignment is autonomous exploration. The robot has to drive autonomously through an unknown world. This feature is implemented using Wavefront frontier detection (WFD).
+
+WFD is a frontier-based exploration algorithm \citep{Keidar}. It uses four different data structures. These store map and frontier information during the search process. They are:
+\begin{enumerate}
+\item{Map-Open-List}
+\item{Map-Close-List}
+\item{Frontier-Open-List}
+\item{Frontier-Close-List}
+\end{enumerate}
+To understand the need of these four structures it is necessary to know that WFD is using a breadth first search (BFS), starting at the current robot position.
+Moreover frontier detection requires the robot to have an internal representation of the current environment.
+In this implementation an occupancy grid (Subsection \ref{sec:mapping}) is used.
+Upon launching the algorithm it first searches for unoccupied space around the current robot location. These grid cells are saved in the Map-Open-List. 
+Frontiers are found by making use of the occupancy grid's structure.
+The criterion for a map-point to be a frontier is the existence of both, unknown space and known unoccupied space, in adjacent map-points.
+When a frontier point is found a new BFS is performed.
+Starting at the current point it extracts all adjacent frontier points and connects them to one frontier.
+It is saved in the Frontier-Open-List.
+The benefit of using WFD is that by maintaining these lists, the algorithm only needs to incorporate points that have not been iterated yet.
+In fact, the algorithm keeps track of detected open space and frontiers using the above mentioned list.
+Following iterations make use of this information because they do not need to consider previously known regions any more.
+
+Exploration does not only consist of finding frontiers.
+Based on the frontiers a target location for the robot needs to be selected.
+Navigating to the closest frontier points is the simplest frontier-based exploration strategy.
+For this project, however, a more sophisticated implementation is chosen.
+Instead of sending the robot to the closest frontier point, a small search is performed to find suitable destinations.
+Generally, a small region around each frontier cell is searched.
+The goal is to give each frontier point a rating on how well it is suited as a destination.
+For this purposes an evaluation function has been implemented.
+It incorporates three features that are used to evaluate goals to be sent to the navigation stack.
+They are :
+\begin{enumerate}
+\item{Number of obstacles in the area}
+\item{Distance to the previous goal}
+\item{Distance between frontier and robot location}
+\end{enumerate}
+The first two features have a weight one, whereas the third is assigned a weight of five.
+This configuration turned out to create the best results (see Subsection \ref{sec:navEval}) The weighted sum of the features is used to find the goal to be sent to the navigation stack.
+This evaluation function enables the algorithm to select a reachable destination for the robot in almost every case.
+The frontier cells are sent to the move\_base in the order of the rating.
+When the navigation stack fails to steer the robot to a designated goal within a time constraint, the next best goal replaces the previous.
+The information gain is still next to optimal since the selected goal is preferred to be close to frontier points.
+At last the second feature makes sure that the robots path is smooth.
+Selecting a goal near the previous one makes the robot avoid too many loops in the emerging graph.
 

### 2013-01-22 Article/Article.tex
Changed the order of the subsections.
discussion and conclusion updated. Scan match reference added
Updated abstract

@@
 \begin{abstract}
-Robots are used for a lot of tasks nowadays, to perform an assignment a robot needs a map of the environment.
-Using external referencing methods, like GPS, they can locate them self.
-Sometimes the maps are provided, making the task easier to perform.
-It becomes far more difficult when the robot has no map or external reference available.
+To perform an assignment a robot needs a map of the environment, an added difficulty could be that the robot needs to build this map on his own.
 The task of mapping an environment can be split into two main components, namely exploration and SLAM which stands for Simultaneous Localization And Mapping.
-The exploration determines the path the robot takes to explore the unknown space of the environment.
-Exploration is based on the idea of frontiers, these are cells which separate unexplored cells from free explored cells.
-The exploration strategy then determines the next location the robot should move to.
-The SLAM part incorporates the new data as is becomes available when the robot moves to unexplored areas and takes measurements of previously unrecorded space.
-One intuitive way is to represent all the data as a graph where the nodes represent a pose of the robot and the edges connecting two nodes the measured constraints.
-The algorithm then tries to find a configuration of nodes which is maximally consistent with the constraints.
+Exploration is based on the idea of frontiers, these are regions which separate unexplored space from free explored space, and a strategy determines to which frontier the robot moves next.
+SLAM incorporates the new data as it becomes available.
+One intuitive way is to represent all the data as a graph and then try to find a configuration of nodes which is maximally consistent with the constraints.
 This article will describe an implementation of this so called GraphSLAM with frontier exploration.
@@
 
-\subsection{Exploration}
-\label{sec:wfd}
-One part of the assignment is autonomous exploration. The robot has to drive autonomously through an unknown world. This feature is implemented using Wavefront frontier detection (WFD).
-
-WFD is a frontier-based exploration algorithm \citep{Keidar}. It uses four different data structures. These store map and frontier information during the search process. They are:
-\begin{enumerate}
-\item{Map-Open-List}
-\item{Map-Close-List}
-\item{Frontier-Open-List}
-\item{Frontier-Close-List}
-\end{enumerate}
-To understand the need of these four structures it is necessary to know that WFD is using a breadth first search (BFS), starting at the current robot position.
-Moreover frontier detection requires the robot to have an internal representation of the current environment.
-In this implementation an occupancy grid (Subsection \ref{sec:mapping}) is used.
-Upon launching the algorithm it first searches for unoccupied space around the current robot location. These grid cells are saved in the Map-Open-List. 
-Frontiers are found by making use of the occupancy grid's structure.
-The criterion for a map-point to be a frontier is the existence of both, unknown space and known unoccupied space, in adjacent map-points.
-When a frontier point is found a new BFS is performed.
-Starting at the current point it extracts all adjacent frontier points and connects them to one frontier.
-It is saved in the Frontier-Open-List.
-The benefit of using WFD is that by maintaining these lists, the algorithm only needs to incorporate points that have not been iterated yet.
-In fact, the algorithm keeps track of detected open space and frontiers using the above mentioned list.
-Following iterations make use of this information because they do not need to consider previously known regions any more.
-
-Exploration does not only consist of finding frontiers.
-Based on the frontiers a target location for the robot needs to be selected.
-Navigating to the closest frontier points is the simplest frontier-based exploration strategy.
-For this project, however, a more sophisticated implementation is chosen.
-Instead of sending the robot to the closest frontier point, a small search is performed to find suitable destinations.
-Generally, a small region around each frontier cell is searched.
-The goal is to give each frontier point a rating on how well it is suited as a destination.
-For this purposes an evaluation function has been implemented.
-It incorporates three features that are used to evaluate goals to be sent to the navigation stack.
-They are :
-\begin{enumerate}
-\item{Number of obstacles in the area}
-\item{Distance to the previous goal}
-\item{Distance between frontier and robot location}
-\end{enumerate}
-The first two features have a weight one, whereas the third is assigned a weight of five.
-This configuration turned out to create the best results (see Subsection \ref{sec:navEval}) The weighted sum of the features is used to find the goal to be sent to the navigation stack.
-This evaluation function enables the algorithm to select a reachable destination for the robot in almost every case.
-The frontier cells are sent to the move\_base in the order of the rating.
-When the navigation stack fails to steer the robot to a designated goal within a time constraint, the next best goal replaces the previous.
-The information gain is still next to optimal since the selected goal is preferred to be close to frontier points.
-At last the second feature makes sure that the robots path is smooth.
-Selecting a goal near the previous one makes the robot avoid too many loops in the emerging graph.
 \subsection{Graph SLAM}

### 2013-01-22 Article/Article.tex
Removed the line referring to the project
Added scanmatch figure and described.

@@
 All the implemented code is first tested and validated in a simulator and later run on a real robot, a TurtleBot.
-
-This article accompanies the research project "TurtleBot SLAM" from the Master "Artificial Intelligence" of Maastricht University.
 \end{abstract}
@@
 Moreover it is used to correct the robot's odometry data.
+
+Figure \ref{fig:scanmatcher} shows the robot in its current position (left) and the result of matching the previous scan to the current scan (right).
+The robot rotated slightly clockwise.
+The red point cloud represents the detected obstacles before the rotation when the robot was still standing parallel to the wall.
+The green point cloud represents the current laser scan.
+The scan matching algorithm then calculates the most likely transformation between these two point clouds and applies this transformation to the previous scan, resulting in the purple point cloud.
+
+\begin{figure}[htbp]
+	\centering
+		\includegraphics[width=0.45\textwidth]{figures/Scanmatch.jpg}
+	\caption{Previous scan matched with the current scan}
+	\label{fig:scanmatcher}
+\end{figure}
 

### 2013-01-22 trunk/Article/Article.tex
added reference for scan match

@@
 \begin{abstract}
-Robots are used for a lot of tasks nowadays, to perform an assignment a robot needs a map of the environment.
-Using external referencing methods, like GPS, they can locate them self.
-Sometimes the maps are provided, making the task easier to perform.
-It becomes far more difficult when the robot has no map or external reference available.
+To perform an assignment a robot needs a map of the environment, an added difficulty could be that the robot needs to build this map on his own.
 The task of mapping an environment can be split into two main components, namely exploration and SLAM which stands for Simultaneous Localization And Mapping.
-The exploration determines the path the robot takes to explore the unknown space of the environment.
-Exploration is based on the idea of frontiers, these are cells which separate unexplored cells from free explored cells.
-The exploration strategy then determines the next location the robot should move to.
-The SLAM part incorporates the new data as is becomes available when the robot moves to unexplored areas and takes measurements of previously unrecorded space.
-One intuitive way is to represent all the data as a graph where the nodes represent a pose of the robot and the edges connecting two nodes the measured constraints.
-The algorithm then tries to find a configuration of nodes which is maximally consistent with the constraints.
+Exploration is based on the idea of frontiers, these are regions which separate unexplored space from free explored space, and a strategy determines to which frontier the robot moves next.
+SLAM incorporates the new data as it becomes available.
+One intuitive way is to represent all the data as a graph and then try to find a configuration of nodes which is maximally consistent with the constraints.
 This article will describe an implementation of this so called GraphSLAM with frontier exploration.
 All the implemented code is first tested and validated in a simulator and later run on a real robot, a TurtleBot.
-
-This article accompanies the research project "TurtleBot SLAM" from the Master "Artificial Intelligence" of Maastricht University.
 \end{abstract}
@@
 At the beginning, the two scans are matched making use of raw odometry data, retrieved by the robot's motion sensor.
-This is improved upon by running an iteration of the iterative closest point algorithm (ICP).
+This is improved upon by running an iteration of the iterative closest point algorithm (ICP) \citep{scanmatch}.
 It tries to align the two scans minimizing the total distance of each point in the first scan to its nearest neighbor in the second.
 
-The general procedure of the algorithm is to first find the nearest neighboring point in the second scan for all point in the first. To achieve this nearest neighbor search is used.
+The general procedure of the algorithm \citep{scanmatch} is to first find the nearest neighboring point in the second scan for all point in the first. To achieve this nearest neighbor search is used.
 In this implementation multiple points can have the same corresponding neighbor-point.

### 2013-01-23 Article/Article.tex
changes

@@
 \label{sec:intro}
-Nowadays mobile robots succeed to assist industry and society in a diverse set of problems.
-Transportation, search, rescue and, more specifically, automated vacuum cleaning are examples from different fields.
+Nowadays mobile robots assist industry and society in a diverse set of problems.
+Transportation, search, rescue and, more specifically, automated vacuum cleaning are examples of such tasks.
 Typically, mobile robots can move through an environment using legs or wheels.
-In addition they are able to gather information about their surrounding using sensors.
-Some even have devices similar to the human arm or hand which allows them to interact with the environment.
+In addition, they are able to gather information about their surrounding using sensors.
+Some even have devices similar to the human arm or hand which allow them to interact with the environment.
 The vacuum cleaning robot depicted in Figure \ref{fig:vacuum_cleaner} is an example of a robot which has wheels to move around.

### 2013-01-23 Article/Article.tex
changes

@@
 These sensors provide data about the location of obstacles as well as, for example, gaps in the floor the robot is moving on.
-To efficiently perform this task robots needs a map of its environment.
-For some tasks map data is available for the robot from the beginning.
-Provided a map the robot can locate itself using the incoming sensor data or, if available, GPS data.
-For tasks in unknown environments, however, the robot must make its own map of the environment.
+To efficiently perform its task the robot needs a map of its environment.
+For some tasks map data is available to the robot from the beginning.
+Provided a map, the robot can locate itself using the incoming sensor data or, if available, external data like GPS.
+For tasks in unknown environments the robot must make its own map of the environment.
 
-The acquisition of these maps has been a major research focus in the robotics community \cite{ Grisetti, Montemerlo02, Montemerlo, Thrun}.
-Building or learning these maps under position uncertainty provided only sensor measurements is often referred to as the Simultaneous Localization And Mapping (SLAM) problem.
+The acquisition of these maps is a major research focus in the robotics community \cite{ Grisetti, Montemerlo02, Montemerlo, Thrun}.
+Building these maps relying only on noisy sensor measurements is often referred to as the Simultaneous Localization And Mapping (SLAM) problem (Subsection \ref{subsec:sotaSlam}).
 A large variety of solutions is available to it.
@@
 		\includegraphics[width=0.50\textwidth]{figures/vacuum_cleaner.jpg}
-	\caption{A vacuum cleaner robot \citep{VacuumRobot}.}
+	\caption{A vacuum cleaner robot \citep{VacuumRobot}}
 	\label{fig:vacuum_cleaner}
@@
 The remainder of this paper is structured as follows.
-Section \ref{sec:sota} provides the state of the art of the discussed fields.
+Section \ref{sec:sota} provides the state of the art of the relevant aspects of this project.
 Section \ref{sec:impl} is about the implementation details of the methods used.
-Section \ref{sec:exp} explains the experiments ran to validate and benchmark the implemented methods and also evaluate the obtained results.
+Section \ref{sec:exp} explains the conducted experiments to validate and benchmark the implemented methods and also evaluates the obtained results.
 Section \ref{sec:disc} discusses the acquired results of the experiments and proposes future research based on those.
-And the last section wraps up the paper with the conclusions of this research.
+And the last section wraps up the paper with the conclusion of this research.
 
@@
 \label{sec:problem}
-This paper accompanies a university group project.
-The project assignment is to implement both a graph-based SLAM algorithm and an autonomous exploration strategy.
-The hardware to be used is a mobile robot called TurtleBot.
-Thus the final product should allow the robot to autonomously navigate through an unknown environment and simultaneously build a map of it using a Graph SLAM.
-
-Information about the programming environment, the TurtleBot, autonomous exploration and SLAM methods are given in Section \ref{sec:sota}.
+This article accompanies a university group project.
+The project assignment is to implement both a Graph SLAM algorithm and an autonomous exploration strategy (Subsection \ref{subsec:sotaExplore}).
+The hardware to be used is a mobile robot called TurtleBot (Subsection \ref{subsec:turtle}).
+The final product should allow the robot to autonomously navigate through an unknown environment and simultaneously build a map of it using a Graph SLAM.
 
@@
 		\includegraphics[width=0.50\textwidth]{figures/graph.png}
-	\caption{An example of a simple graph (based on \citep{Grisetti}).}
+	\caption{An example of a simple graph (based on \citep{Grisetti})}
 	\label{fig:turtlebotexample_graph}
@@
 		\includegraphics[width=0.50\textwidth]{figures/Stage_and_rviz.jpg}
-	\caption{Robot and environment visualization with Stage (Left) and RViz (Right).}
+	\caption{Robot and environment visualization with Stage (Left) and RViz (Right)}
 	\label{fig:stage_and_rviz}
@@
 \end{center}
-\caption{The time the Graph SLAM implementation proposed by \citep{Thrun} needed using the specified number of laser beams after the specified number of steps.}
+\caption{The time the Graph SLAM implementation proposed by \citep{Thrun} needed using the specified number of laser beams after the specified number of steps}
 \label{tab:graphResults}

### 2013-01-23 trunk/Article/Article.tex
changes

@@
 The remainder of this paper is structured as follows.
-Section \ref{sec:sota} provides the state of the art of the discussed fields.
+Section \ref{sec:sota} provides the state of the art of the relevant aspects of this project.
 Section \ref{sec:impl} is about the implementation details of the methods used.
-Section \ref{sec:exp} explains the experiments ran to validate and benchmark the implemented methods and also evaluate the obtained results.
+Section \ref{sec:exp} explains the conducted experiments to validate and benchmark the implemented methods and also evaluates the obtained results.
 Section \ref{sec:disc} discusses the acquired results of the experiments and proposes future research based on those.
-And the last section wraps up the paper with the conclusions of this research.
+And the last section wraps up the paper with the conclusion of this research.
 
 Subsection \ref{sec:problem} defines the problem statement for this project.
-\subsection{Problem Statement}
+\subsection{Problem statement}
 \label{sec:problem}
-This paper accompanies a university group project.
-The project assignment is to implement both a graph-based SLAM algorithm and an autonomous exploration strategy.
-The hardware to be used is a mobile robot called TurtleBot.
-Thus the final product should allow the robot to autonomously navigate through an unknown environment and simultaneously build a map of it using a Graph SLAM.
-
-Information about the programming environment, the TurtleBot, autonomous exploration and SLAM methods are given in Section \ref{sec:sota}.
+This article accompanies a university group project.
+The project assignment is to implement both a Graph SLAM algorithm and an autonomous exploration strategy (Subsection \ref{subsec:sotaExplore}).
+The hardware to be used is a mobile robot called TurtleBot (Subsection \ref{subsec:turtle}).
+The final product should allow the robot to autonomously navigate through an unknown environment and simultaneously build a map using Graph SLAM.
 
@@
 \label{subsec:ros}
-The Robot Operating System (ROS) \cite{Quigley} is the environment used throughout the development of all product components for this project.
-As described by Quigley et al., writing software for robots is difficult since there is a large variety of possible hardware set-ups.
+The Robot Operating System (ROS) \cite{Quigley} is the programming environment used for the development of all product components of this project.
+As described by Quigley et al.
+\citep{Quigley}, writing software for robots is difficult since there is a large variety of possible hardware set-ups.
 This also makes code reuse non-trivial.
-An additional problem is the amount of code needed to actually ``run" something on a robot.
-A deep stack including driver-level software, perception modules and, for example, abstract reasoning systems is required before the actual problem can be attacked.
-Since the necessary expertise and effort is beyond capabilities of most single researchers, an environment that features this functionality is required to achieve anything similar to the problem specified in Subsection \ref{sec:problem}.
+An additional problem is the amount of code needed to run a program on a robot.
+A large framework including driver-level software, perception modules and, for example, abstract reasoning systems is required before the actual problem can be approached.
+Since the necessary expertise and effort is beyond capabilities of most single researchers, an environment that features this functionality is required to efficiently solve problems in robotics.
 
-To deal with the above mentioned problems robotics researchers have created a wide variety of frameworks.
-Many robotic software systems are currently used by researchers and industry \cite{Kramer}.
-Each of these frameworks is designed for a specific purpose, though.
-Generally, none of these systems is directly applicable to a problem it was not designed for in the first place.
+To deal with the problems mentioned above robotics researchers have created a wide variety of frameworks.
+Many robotics software systems are currently used by researchers and industry \cite{Kramer}.
+Each of these frameworks is designed for a specific purpose.
 
 The ROS framework used in this paper is the product of trade-offs and prioritizations.
-The emphasis of this framework lies on large-scale integrative robotics research, making it useful in a wide variety of situations.
-The difference between ROS and the above mentioned frameworks, is ROS' applicability to a wide range of hardware set-ups.
+Its emphasis lies on large-scale integrative robotics research, making it useful in a wide variety of situations.
+The main difference between ROS and the frameworks mentioned above, is its applicability to a wide range of hardware set-ups.
 It provides several so called stacks.
@@
 For most robots the configuration is straight forward thanks to the on-line documentation of ROS \citep{Roswiki}.
-An example for a stack is the multi-robot simulation environment ``Stage".
-It is explained in the following paragraph.
 
-\subsubsection{Stage}
+\subsubsection{Stage and RViz}
 During the development process of a system running on a robot one crucial part is testing.
-A ROS stack called Stage enables the developers to test code on a simulated robot.
-Stage allows the simulated robot to move around in virtual environments.
-It visualizes sensor data and other robot data when required.
-The user can load her own maps into stage and specify the kind of noise to be applied to measurements.
+A ROS stack called Stage enables developers to test code on a simulated robot allowing it to move around in virtual environments.
+Stage visualizes the robot, its movement and measurements.
+The user can load individually created maps into Stage and specify the robot's configuration.
 A second important stack needed especially for validation is RViz.
@@
 		\includegraphics[width=0.50\textwidth]{figures/graph.png}
-	\caption{An example of a simple graph (based on \citep{Grisetti}).}
+	\caption{An example of a simple graph \citep{Grisetti}}
 	\label{fig:turtlebotexample_graph}

### 2013-01-23 Article/Article.tex
changes

@@
 Subsection \ref{sec:problem} defines the problem statement for this project.
-\subsection{Problem Statement}
+\subsection{Problem statement}
 \label{sec:problem}
@@
 The hardware to be used is a mobile robot called TurtleBot (Subsection \ref{subsec:turtle}).
-The final product should allow the robot to autonomously navigate through an unknown environment and simultaneously build a map of it using a Graph SLAM.
+The final product should allow the robot to autonomously navigate through an unknown environment and simultaneously build a map using Graph SLAM.
 
@@
 In Figure \ref{fig:stage_and_rviz} Stage and RViz are presented next to each other.
-While Stage (Left) shows the real map, RViz (Right) presents the map data as seen by the robot.
+While Stage (Left) shows the real map, RViz (Right) presents the map data as built by the robot.
 
@@
 The base of the TurtleBot is an iRobot Create (A).
-It holds a battery pack and a 150 degrees/second Single Axis Gyro.
-The model shown in Figure \ref{fig:turtlebot} uses a Microsoft Kinect sensor (B), the one used in this paper uses a laser range sensor and a laptop (C) to run all the processes.
+It holds a battery pack and a 150 $degrees/second$ Single Axis Gyro.
+The model shown in Figure \ref{fig:turtlebot} uses a Microsoft Kinect sensor (B), the one used in this paper uses a laser range sensor, and a laptop (C) to run all the processes.
 The robot can also be customized using the mounting hardware (D).
 The robot runs on an open-source SDK based on ROS which integrates all the software needed to get the TurtleBot running.
-Some capabilities are included in ROS and can be used in the TurtleBot after minor configurations.
+Some capabilities are included in ROS and can be used on the TurtleBot after minor configurations.
 These include mapping and navigation, for example.
@@
 		\includegraphics[width=0.50\textwidth]{figures/turtlebot.png}
-	\caption{A TurtleBot}
+	\caption{A TurtleBot \citep{turtlebot}
 	\label{fig:turtlebot}
@@
 Usually robots localize themselves and navigate through environments using a map which is provided by an external source.
-In this case navigating from one location to another is an easy problem.
+In this case navigating from one location to another is an easy task.
 If the robot does not have access to a map, navigating through an environment is problematic.
@@
 In addition to the creation of the map and the localization of itself (the SLAM problem) the robot needs to explore the environment as efficient as possible.
-We define a good exploration strategy as one that generates an accurate map in the shortest amount of time.
+A good exploration strategy is defined as one that generates an accurate map in the shortest amount of time.
 
@@
 Frontiers are defined to be the boundaries between free and unexplored space.
-Intuitively, when a robot moves to a frontier it is likely that it makes measurements in previously unexplored space.
-Given that new data can be added to its map.
-The explored region grows once the robot reaches a frontier.
+Intuitively, when a robot moves to a frontier it is likely that it takes measurements of previously unexplored space.
+Given that, new data can be added to its map.
+The explored region expands once the robot reaches a frontier.
 Moreover an extended map results in new frontiers.
@@
 
-Other approaches estimate the complete trajectory from the set of measurements.
-These approaches address the so called full SLAM problem and they typically rely on least-square error minimization techniques \cite{Leastsquares}.
-
-An intuitive way to formulate the SLAM problem is to consider it as a pose graph optimization problem.
-The nodes of the graph represent robot poses or landmarks.
-The edges encode sensor measurements that constraint the connected poses.
-Once such a graph is constructed the graph-based SLAM, or Graph SLAM, method tries to find a configuration of the nodes that is maximally consistent with the constraints contained in the edges.\\
-
-SLAM is a problem because localization requires a map and mapping requires the location of the device or robot that wants to create a map.
+SLAM is a problem because localization requires a map and mapping requires the location of the robot that wants to create a map.
 There are quite a few approaches that successfully solve the SLAM problem.
@@
 Due to their incremental nature these filtering approaches are commonly referred to as on-line SLAM methods.
-FastSlam \citep{Montemerlo02} is an example of an on-line SLAM implementation.
-It uses a Particle Filter to approximate the robots pose and the landmark positions on the map.
+FastSLAM \citep{Montemerlo02} is an example of an on-line SLAM implementation.
+It uses a Particle Filter to approximate the robot's pose and the landmark positions on the map.
 Other methods have the goal to find the trajectory of the robot and the map.
 These solve the full SLAM problem and typically rely on least-square error minimization techniques \cite{Leastsquares} to find adequate estimates.
-An example of such a an approach is graph-based SLAM or Graph SLAM.
+An example of such a an approach is Graph SLAM.
 
@@
 		\includegraphics[width=0.50\textwidth]{figures/graph.png}
-	\caption{An example of a simple graph (based on \citep{Grisetti})}
+	\caption{An example of a simple graph \citep{Grisetti}}
 	\label{fig:turtlebotexample_graph}

### 2013-01-23 Article/Article.tex
works

@@
 
-The acquisition of these maps is a major research focus in the robotics community \cite{ Grisetti, Montemerlo02, Montemerlo, Thrun}.
+The acquisition of these maps is a major research focus in the robotics community \citep{Grisetti, Montemerlo02, Montemerlo, Thrun}.
 Building these maps relying only on noisy sensor measurements is often referred to as the Simultaneous Localization And Mapping (SLAM) problem (Subsection \ref{subsec:sotaSlam}).
@@
 		\includegraphics[width=0.50\textwidth]{figures/turtlebot.png}
-	\caption{A TurtleBot \citep{turtlebot}
+	\caption{A TurtleBot \citep{turtlebot}}
 	\label{fig:turtlebot}
@@
 An intuitive way to formulate the SLAM problem is to view it as a pose graph optimization problem.
-Each pose along the trajectory of the robot is represented by one node.
+Each pose along the trajectory of the robot is represented by a node.
 Edges are labeled with relative spatial information between the connecting nodes.

### 2013-01-23 Article/Article.tex
Corrected Graph SLAM and TurtleBot.

@@
 
-\title{Turtlebot SLAM }
+\title{TurtleBot SLAM }
 
@@
 One intuitive way is to represent all the data as a graph and then try to find a configuration of nodes which is maximally consistent with the constraints.
-This article will describe an implementation of this so called GraphSLAM with frontier exploration.
+This article will describe an implementation of this so called Graph SLAM with frontier exploration.
 All the implemented code is first tested and validated in a simulator and later run on a real robot, a TurtleBot.
@@
 
-This graph-based SLAM approach calculates the most likely path as follows:
+This Graph SLAM approach calculates the most likely path as follows:
 First it finds the Jacobians \(A\) and \(B\) of every two estimated poses connected by an edge.

### 2013-01-23 Article/Article.tex
changes

@@
 The product of this project is designed to run on a mobile robot called TurtleBot (Figure \ref{fig:turtlebot}) \citep{turtlebot}.
-It is a low-cost, customizable, personal robot kit.
+It is a low-cost, customizable, personal robot.
 The base of the TurtleBot is an iRobot Create (A).
@@
 \subsubsection{Graph SLAM}
-An intuitive way to formulate the SLAM problem is to view it as a pose graph optimization problem.
+An intuitive way to formulate the SLAM problem is to view it as a pose graph optimization problem \citep{Grisetti}.
 Each pose along the trajectory of the robot is represented by a node.
@@
 The essence of the Graph SLAM method, however, is the optimization step.
-In this step the algorithm tries to find a spatial configuration of the trajectory which is maximally consistent with a system of constraints obtained from the edges.
-This problem is commonly solved using least-square error minimization techniques \citep{Grisetti}.
+In this step the algorithm tries to find a spatial configuration of the trajectory which is maximally consistent with the system of constraints obtained from the edges.
+This problem is commonly solved using least-square error minimization techniques \citep{Leastsquares}.
 
@@
 \subsection{Program structure}
+\label{sec:struct}
 An overview of the relevant components of the final software product is given in Figure \ref{fig:program_structure}.
@@
 
-The implemented packages are making use of the external library ``Eigen" \citep{eigen}.
+The implemented packages make use of the external library ``Eigen" \citep{eigen}.
 It provides extensive linear algebra operations.
@@
 \label{sec:implSLAM}
-The implemented Graph SLAM algorithm is based on \citep{Grisetti}.
+The implemented Graph SLAM algorithm is based on Grisetti et al.\citep{Grisetti}.
 The graph consists of nodes representing poses along the robot trajectory.
@@
 Note that in Equations \ref{eq:jacobi_a} and \ref{eq:jacobi_b} the x represents the current guess of the concerning robot pose.
-As explained above, \(e_{ij}(x_i, x_j)\) is the error between the two poses \(x_i\) and \(x_j\).
+As explained above, \(e_{ij}(x)\) is the error between the two poses \(x_i\) and \(x_j\).
 The Jacobians \(A\) and \(B\) are then added to a matrix \(H\) (Equation \ref{eq:H}).
@@
 This happens when the distance between the two connected robot poses is smaller than a margin.
-In fact, this special case enables the algorithm to create loops in the graph and more accurately estimate the robots trajectory and the map.
+In fact, this special case enables the algorithm to create loops in the graph and more accurately estimate the robot's trajectory and the map.
 This procedure is referred to as ``loop closing" (see Subsection \ref{sec:loopClosing}).
@@
 In that way the clouds are centered at the origin.
-The calculated points are then added to two separate $2 \times n$ matrices, where $n$ is the number of neighbor pairs.
+The coordinates to the calculated points are then added to two separate $2 \times n$ matrices, where $n$ is the number of neighbor pairs.
 Next the rotation matrix which transforms one cloud to the other is needed.
@@
 \label{sec:mapping}
-The map is internally represented as an occupancy grid. It consists of cells which hold information about obstacles, free/- and unknown space.
+The map is internally represented as an occupancy grid. It consists of cells which hold information about obstacles, free and unknown space.
 
@@
 Fitting a map on a smaller grid results in a low resolution.
-Internally the program represents the grid as a two dimensional array.
+The program treats the grid as a two dimensional array.
 This makes navigating to either side more intuitive.
@@
 \label{sec:wfd}
-One part of the assignment is autonomous exploration. The robot has to drive autonomously through an unknown world. This feature is implemented using Wavefront frontier detection (WFD).
+One part of the assignment is autonomous exploration. The robot has to drive autonomously through an unknown world. This feature is implemented using Wavefront Frontier Detection (WFD \citep{Keidar}).
 
-WFD is a frontier-based exploration algorithm \citep{Keidar}. It uses four different data structures. These store map and frontier information during the search process. They are:
+WFD is a frontier-based exploration algorithm. It uses four different data structures. These store map and frontier information during the search process. They are:
 \begin{enumerate}
@@
 This evaluation function enables the algorithm to select a reachable destination for the robot in almost every case.
-The frontier cells are sent to the move\_base in the order of the rating.
+The frontier cells are sent to the navigation stack in the order of the rating.
 When the navigation stack fails to steer the robot to a designated goal within a time constraint, the next best goal replaces the previous.
 The information gain is still next to optimal since the selected goal is preferred to be close to frontier points.
-At last the second feature makes sure that the robots path is smooth.
+At last the second feature makes sure that the robot's path is smooth.
 Selecting a goal near the previous one makes the robot avoid too many loops in the emerging graph.
@@
 \end{figure}
-The next section explains which experiments have been conducted in order to find an optimal exploration strategy. It presents results for different goal selection methods.
 
@@
 In this section the results of the implemented methods are evaluated and discussed. 
-Subsection \ref{disc:expl} deals with the exploration strategy. Subsection \ref{disc:grSlam} presents the evolution of the Graph SLAM implementation. Different stages of it are evaluated.
+Subsection \ref{disc:expl} deals with the exploration strategy.
+Subsection \ref{disc:grSlam} presents the evolution of the Graph SLAM implementation.
+Different stages of it are evaluated.
+Finally, Subsection \ref{sec:future} suggests future research.
 \subsection{Exploration strategy}
@@
 The major difference lies in the node creation.
-Here, additional nodes are added for landmarks while the previous implementation considered the robot's trajectory only.
-An additional difference lies in the fact that this implementation uses features to characterize the map.
-The algorithm needs to compare every feature from the new scan has to the ones detected earlier.
+Here, additional nodes are added for features while the previous implementation considered the robot's trajectory only.
+An additional difference lies in the fact that this implementation uses only the features to characterize the map.
+The algorithm needs to compare every feature from the new scan to the ones detected earlier.
 This ``correspondence test" has to be performed and repeated along with the complete optimization procedure until convergence.
@@
 
-Implementing the algorithm suggested in \citep{Grisetti} (Subsection \ref{sec:implSLAM}) showed to be well suited for attacking the problem statement of this project.
+Implementing the algorithm suggested in \citep{Grisetti} (Subsection \ref{sec:implSLAM}) showed to be well suited for approaching the problem statement of this project.
 The maps created by the algorithm are accurate (see Subsection \ref{sec:visual}).
@@
 In fact, the algorithm is able to keep updating the map with no noticeable delay (see Section \ref{sec:exp}).
-Even when the robot is controlled by the exploration module and navigated by ROS' move\_base package the program runs smoothly.
+Even when the robot is controlled by the exploration module and navigated by ROS' navigation stack the program runs smoothly.
 
@@
 \subsection{Future research}
+\label{sec:future}
 As mentioned in Section \ref{disc:expl} the exploration strategy evolved during the process of testing.
@@
 The current implementation uses sparse Cholesky factorization to find the most optimal spatial configuration of the trajectory.
-This method is inherited in the ``Eigen" library \ref{sec:impl}.
+This method is inherited in the ``Eigen" library (Subsection \ref{sec:struct}).
 Although sparse Cholesky factorization solves the linear system in an efficient manner, it is likely that even better algorithms emerge in the future.

### 2013-01-23 trunk/Article/Article.tex
minor changes

@@
 The product of this project is designed to run on a mobile robot called TurtleBot (Figure \ref{fig:turtlebot}) \citep{turtlebot}.
-It is a low-cost, customizable, personal robot kit.
+It is a low-cost, customizable, personal robot.
 The base of the TurtleBot is an iRobot Create (A).
-It holds a battery pack and a 150 degrees/second Single Axis Gyro.
-The model shown in Figure \ref{fig:turtlebot} uses a Microsoft Kinect sensor (B), the one used in this paper uses a laser range sensor and a laptop (C) to run all the processes.
+It holds a battery pack and a 150 $degrees/second$ Single Axis Gyro.
+The model shown in Figure \ref{fig:turtlebot} uses a Microsoft Kinect sensor (B), the one used in this paper uses a laser range sensor, and a laptop (C) to run all the processes.
 The robot can also be customized using the mounting hardware (D).
 The robot runs on an open-source SDK based on ROS which integrates all the software needed to get the TurtleBot running.
-Some capabilities are included in ROS and can be used in the TurtleBot after minor configurations.
+Some capabilities are included in ROS and can be used on the TurtleBot after minor configurations.
 These include mapping and navigation, for example.
@@
 		\includegraphics[width=0.50\textwidth]{figures/turtlebot.png}
-	\caption{A TurtleBot}
+	\caption{A TurtleBot \citep{turtlebot}}
 	\label{fig:turtlebot}
@@
 Due to their incremental nature these filtering approaches are commonly referred to as on-line SLAM methods.
-FastSlam \citep{Montemerlo02} is an example of an on-line SLAM implementation.
-It uses a Particle Filter to approximate the robots pose and the landmark positions on the map.
+FastSLAM \citep{Montemerlo02} is an example of an on-line SLAM implementation.
+It uses a Particle Filter to approximate the robot's pose and the landmark positions on the map.
 Other methods have the goal to find the trajectory of the robot and the map.
 These solve the full SLAM problem and typically rely on least-square error minimization techniques \cite{Leastsquares} to find adequate estimates.
-An example of such a an approach is graph-based SLAM or Graph SLAM.
+An example of such a an approach is Graph SLAM.
 
 \subsubsection{Graph SLAM}
-An intuitive way to formulate the SLAM problem is to view it as a pose graph optimization problem.
-Each pose along the trajectory of the robot is represented by one node.
+An intuitive way to formulate the SLAM problem is to view it as a pose graph optimization problem \citep{Grisetti}.
+Each pose along the trajectory of the robot is represented by a node.
 Edges are labeled with relative spatial information between the connecting nodes.
@@
 This ``optimized" trajectory then determines the positions of obstacles on the map.
-The laser scan data are refined according to the poses they were obtained at.
+The laser scan data is refined according to the poses they were obtained at.
 Each edge, thus, is labeled with the spatial difference between two poses \(x_i\) and \(x_j\).
@@
 The size of an occupancy grid has to be tuned problem specific.
-A too large grid can lead to performance problems.
+A grid, which is too large can lead to performance problems.
 If the grid is too small, however, the map will be inaccurate.
 Fitting a map on a smaller grid results in a low resolution.
-Internally the program represents the grid as a two dimensional array.
+The program treats the grid as a two dimensional array.
 This makes navigating to either side more intuitive.
@@
 Frontiers are found by making use of the occupancy grid's structure.
-The criterion for a map-point to be a frontier is the existence of both, unknown space and known unoccupied space, in adjacent map-points.
 When a frontier point is found a new BFS is performed.
@@
 One of the core properties of the Graph SLAM algorithm is the correction of the pose graph and the map whenever a loop closing is detected.
-This means that when the robot has been exploring an unknown area and then re-enters an explored area, the transformation between the current pose and the past, re-detected, pose gives essential information about the complete pose graph.
+This means that when the robot is exploring an unknown area and then re-enters an explored area, the transformation between the current pose and the past, re-detected, pose gives essential information about the complete pose graph.
 It corrects the current belief about environment and position.
@@
 \subsection{Future research}
+\label{sec:future}
 As mentioned in Section \ref{disc:expl} the exploration strategy evolved during the process of testing.
-This does not mean that it cannot be improved upon.
+This does not mean that it cannot be improved.
 More features like, for instance, loop closings can be taken into account.
@@
 The current implementation uses sparse Cholesky factorization to find the most optimal spatial configuration of the trajectory.
-This method is inherited in the ``Eigen" library \ref{sec:impl}.
+This method is inherited in the ``Eigen" library (Subsection \ref{sec:struct}).
 Although sparse Cholesky factorization solves the linear system in an efficient manner, it is likely that even better algorithms emerge in the future.
-Since the time complexity of the Graph SLAM method depends almost entirely on the technique used, new methods would directly improve upon the running time of the algorithm.
+Since the time complexity of the Graph SLAM method depends almost entirely on the technique used, new methods would directly improve the running time of the algorithm.
 

### 2013-01-23 Article/Article.tex
abstract

@@
 \begin{abstract}
-To perform an assignment a robot needs a map of the environment, an added difficulty could be that the robot needs to build this map on his own.
-The task of mapping an environment can be split into two main components, namely exploration and SLAM which stands for Simultaneous Localization And Mapping.
-Exploration is based on the idea of frontiers, these are regions which separate unexplored space from free explored space, and a strategy determines to which frontier the robot moves next.
-SLAM incorporates the new data as it becomes available.
-One intuitive way is to represent all the data as a graph and then try to find a configuration of nodes which is maximally consistent with the constraints.
-This article will describe an implementation of this so called Graph SLAM with frontier exploration.
+ This article describes an implementation of Graph SLAM with frontier-based exploration.
+Simultaneous localization and mapping (SLAM) requires knowledge of the map and the robot's location at the same time.
+Achieving this is hard, as the robot's location needs knowledge of the map and vice-versa.
+To explore the map autonomously the robot has to use an algorithm to efficiently achieve knowledge of its region of operation.
+For the implementation in this article a Graph SLAM algorithm and a Wavefront frontier detection algorithm have been used to achieve both.
 All the implemented code is first tested and validated in a simulator and later run on a real robot, a TurtleBot.
+The result is a quickly built map of a sample region.
 \end{abstract}
@@
 The size of an occupancy grid has to be tuned problem specific.
-A too large grid can lead to performance problems.
+A grid, which is too large can lead to performance problems.
 If the grid is too small, however, the map will be inaccurate.
@@
 As mentioned in Section \ref{disc:expl} the exploration strategy evolved during the process of testing.
-This does not mean that it cannot be improved upon.
+This does not mean that it cannot be improved.
 More features like, for instance, loop closings can be taken into account.
@@
 Although sparse Cholesky factorization solves the linear system in an efficient manner, it is likely that even better algorithms emerge in the future.
-Since the time complexity of the Graph SLAM method depends almost entirely on the technique used, new methods would directly improve upon the running time of the algorithm.
+Since the time complexity of the Graph SLAM method depends almost entirely on the technique used, new methods would directly improve the running time of the algorithm.
 

### 2013-01-23 Article/Article.tex

@@
 \begin{abstract}
- This article describes an implementation of Graph SLAM with frontier-based exploration.
-Simultaneous localization and mapping (SLAM) requires knowledge of the map and the robot's location at the same time.
-Achieving this is hard, as the robot's location needs knowledge of the map and vice-versa.
+ This article describes an implementation of Graph-based Simultaneous localization and mapping (Graph SLAM) with frontier-based exploration.
+The challenge of the SLAM-problem is building the map of a robot's environment and localizing it using uncertain measurement data.
+Robot localization needs knowledge of the map and vice-versa.
 To explore the map autonomously the robot has to use an algorithm to efficiently achieve knowledge of its region of operation.

### 2013-01-23 trunk/Article/Article.tex

@@
 \begin{abstract}
-To perform an assignment a robot needs a map of the environment, an added difficulty could be that the robot needs to build this map on his own.
-The task of mapping an environment can be split into two main components, namely exploration and SLAM which stands for Simultaneous Localization And Mapping.
-Exploration is based on the idea of frontiers, these are regions which separate unexplored space from free explored space, and a strategy determines to which frontier the robot moves next.
-SLAM incorporates the new data as it becomes available.
-One intuitive way is to represent all the data as a graph and then try to find a configuration of nodes which is maximally consistent with the constraints.
-This article will describe an implementation of this so called Graph SLAM with frontier exploration.
+ This article describes an implementation of Graph-based Simultaneous Localization And Mapping (Graph SLAM) with frontier-based exploration.
+The challenge of the SLAM-problem is building a map of the robot's environment and localizing it using uncertain measurement data.
+Robot localization needs knowledge of the map and vice-versa.
+Another challenge is to find a strategy to efficiently explore the robot's environment.
+In the implementation for this article a Graph SLAM algorithm and a Wavefront frontier detection algorithm have been used to achieve both.
 All the implemented code is first tested and validated in a simulator and later run on a real robot, a TurtleBot.
+The result is a quickly built map of a sample region.
 \end{abstract}

### 2013-01-23 Article/Article.tex

@@
 \begin{abstract}
- This article describes an implementation of Graph-based Simultaneous localization and mapping (Graph SLAM) with frontier-based exploration.
-The challenge of the SLAM-problem is building the map of a robot's environment and localizing it using uncertain measurement data.
+ This article describes an implementation of Graph-based Simultaneous Localization And Mapping (Graph SLAM) with frontier-based exploration.
+The challenge of the SLAM-problem is building a map of the robot's environment and localizing it using uncertain measurement data.
 Robot localization needs knowledge of the map and vice-versa.
-To explore the map autonomously the robot has to use an algorithm to efficiently achieve knowledge of its region of operation.
-For the implementation in this article a Graph SLAM algorithm and a Wavefront frontier detection algorithm have been used to achieve both.
+Another challenge is finding a strategy to efficiently explore the robot's environment.
+A Graph SLAM algorithm and a Wavefront frontier detection algorithm have been implemented to solve these problems.
+This article gives an insight into the implementation process of these algorithms and on possible solutions.
 All the implemented code is first tested and validated in a simulator and later run on a real robot, a TurtleBot.

### 2013-01-23 trunk/Article/Article.tex

@@
 Robot localization needs knowledge of the map and vice-versa.
-Another challenge is to find a strategy to efficiently explore the robot's environment.
-In the implementation for this article a Graph SLAM algorithm and a Wavefront frontier detection algorithm have been used to achieve both.
-All the implemented code is first tested and validated in a simulator and later run on a real robot, a TurtleBot.
-The result is a quickly built map of a sample region.
+Another challenge is finding a strategy to efficiently explore the robot's environment.
+A Graph SLAM algorithm and a Wavefront frontier detection algorithm have been implemented to solve these problems.
+This article gives an insight into the implementation process of these algorithms.
+The implementation is tested and validated in a simulator.
+The result of the process is a fully explored sample environment.
 \end{abstract}

### 2013-01-23 Article/Article.tex

@@
 A Graph SLAM algorithm and a Wavefront frontier detection algorithm have been implemented to solve these problems.
-This article gives an insight into the implementation process of these algorithms and on possible solutions.
-All the implemented code is first tested and validated in a simulator and later run on a real robot, a TurtleBot.
-The result is a quickly built map of a sample region.
+This article gives an insight into the field of robotics and into the implementation process.
+The implementation is tested and validated in a simulator.
+The result of the process is a fully explored sample environment.
 \end{abstract}

### 2013-01-23 trunk/Article/Article.tex

@@
 A Graph SLAM algorithm and a Wavefront frontier detection algorithm have been implemented to solve these problems.
-This article gives an insight into the implementation process of these algorithms.
-The implementation is tested and validated in a simulator.
-The result of the process is a fully explored sample environment.
+This article gives an insight into the field of robotics and into the implementation process of the suggested algorithms.
+The resulting program is tested and validated in a simulator.
+The final product is able to fully explore an environment autonomously and to create a map during the exploration process.
 \end{abstract}

### 2013-01-24 trunk/turtlebot_graph_slam/src/test.cpp
Working demo.

@@ -65,7 +65,7 @@ bool initialised;
 bool loopDetected;
 
 // Distance after which a laser scan should be matched again
-double MATCH_DISTANCE = 0.5;
+double MATCH_DISTANCE = 0.3;
 void callback(const sensor_msgs::LaserScan::ConstPtr& msg) {
     /**
       Simply save the current scan.

### 2013-01-24 trunk/wfd_exploration/src/exploration.cpp

@@ -266,6 +266,7 @@ protected:
 };
 
 int main(int argc, char **argv) {
+    std::cout << "MAIN EXPLORATION";
     ros::init(argc, argv, "Exploration");
     ros::NodeHandle n;
     Exploration walker(n); // Create new Exploration object
